### 2024-11-18 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@TobiBu}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2024-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
-% Define document class
-\documentclass[twocolumn]{aastex631}
-\usepackage{showyourwork}
+%                                                                 aa.dem
+% AA vers. 9.1, LaTeX class for Astronomy & Astrophysics
+% demonstration file
+%                                                       (c) EDP Sciences
+%-----------------------------------------------------------------------
+%
+%\documentclass[referee]{aa} % for a referee version
+%\documentclass[onecolumn]{aa} % for a paper on 1 column  
+%\documentclass[longauth]{aa} % for the long lists of affiliations 
+%\documentclass[letter]{aa} % for the letters 
+%\documentclass[bibyear]{aa} % if the references are not structured 
+%                              according to the author-year natbib style
 
-% Begin!
+%
+\documentclass{aa}
+
+%
+\usepackage{graphicx}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{txfonts}
+\usepackage{siunitx}
+\usepackage{hyperref}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\usepackage[options]{hyperref}
+% To add links in your PDF file, use the package "hyperref"
+% with options according to your LaTeX or PDFLaTeX drivers.
+%
+
+\usepackage{tikz}
+
+\usepackage{xcolor}
+\newcommand{\todo}[1]{\textcolor{red}{#1}}
+\newcommand{\T}[1]{\textcolor{orange}{#1}}
+%\newcommand{\U}[1]{\textbf{\textit{\textcolor{blue}{#1}}}}
+\newcommand{\U}[1]{\textcolor{black}{#1}}
+\usepackage{subfigure}
 \begin{document}
 
-% Title
-\title{An open source scientific article}
+   \title{Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference}
 
-% Author list
-\author{@TobiBu}
+   %\subtitle{Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference}
 
-% Abstract with filler text
-\begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
-\end{abstract}
+   \author{Tobias Buck\inst{1,2}
+          \and
+          Berkay Günes\inst{1,2}
+          }
 
-% Main body with filler text
+   \institute{Interdisciplinary Center for Scientific Computing (IWR), University of Heidelberg,
+ Im Neuenheimer Feld 205, D-69120 Heidelberg
+ \and
+ Universität Heidelberg, Zentrum für Astronomie, Institut für Theoretische Astrophysik, Albert-Ueberle-Straße 2, D-69120 Heidelberg, Germany\\
+ \email{tobias.buck@iwr.uni-heidelberg.de},  \email{b.guenes@stud.uni-heidelberg.de}
+             }
+
+   \date{Received Month, XXXX; accepted Month Day, XXXX}
+
+% \abstract{}{}{}{}{} 
+% 5 {} token are mandatory
+ \abstract
+  % context heading (optional)
+  % {} leave it empty if necessary  
+   {Chemical abundances of stars are able to reveal important galactic parameters, such as the initial mass function high-mass slope and the frequency of type Ia supernovae.
+Constraining these parameters is of critical importance in order to create realistic hydrodynamical simulations and understand galactic physics.
+However, so far inference was significantly limited by the computational cost of traditional Bayesian algorithms.
+Here we present a new method using simulation-based inference (SBI) to circumvent previous limitations and enable inference from massive stellar surveys.
+    SBI enables a new approach to the inference of the posterior distribution of Galactic parameters under an intractable likelihood function by utilizing forward simulations of stellar abundance distributions.
+    Using SBI we are able to predict global galactic parameters of great accuracy and precision from chemical abundances of multiple stars, faster than conventional methods like Hamiltonian Monte Carlo inference.
+Our method easily scales to hundred-thousands of stars and beats previous approaches by orders of magnitude in compute time and accuracy.}
+  % aims heading (mandatory)
+   {}
+  % methods heading (mandatory)
+   {}
+  % results heading (mandatory)
+   {}
+  % conclusions heading (optional), leave it empty if necessary 
+   {}
+
+   \keywords{Galaxies: fundamental parameters --
+            Galaxies: stellar content --
+             Methods: data analysis --
+             Methods: statistical --
+             }
+\maketitle
+%-------------------------------------------------------------------
 \section{Introduction}
-\label{sec:intro}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+\section{Motivation}
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars.
+Two crucial unknowns are the shape of the initial mass function (IMF), setting the mass distribution of stars born from the interstellar medium (ISM), and the rate of Type Ia supernovae (SN\,Ia) explosions.
+%
+The exact values of parameters sensitively affect chemical evolution tracks \citep{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, yet their observational constraints remain weak.
+%
+For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,2015MNRAS.447.3880R,2014ApJ...796...75C}.
+In addition, the IMF slope may itself be not a constant but rather a function of metallicity, introducing further complexity \citep[e.g.][]{2019MNRAS.482..118G,2019A&A...626A.124M}.
+Similarly, the choice of SN\,Ia delay-time-distribution and normalization plays a crucial role in the enrichment of the ISM \citep[e.g.][]{Buck2021} and is heavily debated \citep{2010ApJ...722.1879M,2012MNRAS.426.3282M,2015ApJ...810..137J}.
 
-\bibliography{bib}
+The goal of this work is to demonstrate how we can use modern machine learning techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
+Our work focuses on two key parameters; the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF and the rate of SN\,Ia explosions per unit mass.
+We assume that both are constant across the galaxy and time.
+%
+Our method will make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
+For high-dimensional posterior functions, SBI gives tremendous computational improvements over more traditional methods such as Markov Chain Monte Carlo (MCMC) or even Hamilton Monte Carlo (HMC) methods.
 
+% \begin{figure*}
+%    %\sidecaption
+%    \subfigure[]{
+%        \includegraphics[width=0.49\hsize, trim={0 0 0 0cm},clip]{figures/data/sample_galaxies_with_colorbar.pdf}
+%    }
+%    \subfigure[]{
+%       \includegraphics[width=0.51\hsize, trim={0 0 2.3cm 0},clip]{figures/data/galaxy_grid_with_colorbar.pdf}
+%    }
+%    \caption{Galaxy images (upper panels) and image cubes (lower panels) in our dataset after the preprocessing steps as explained in Section \ref{sec:dataset_generation}.
+Each row contains one sample galaxy in three stellar maps - metallicity, age and mass from left to right in a) 2D and b) 3D.
+The images are normalised in the range $[0,1]$ and bright pixels correspond to high values.
+Note that in the stellar age maps, dark pixels correspond to young stars.
+The three dimensional plots have been created using the \emph{plotly} python package.}
+%    \label{fig: sample_galaxies_after_prepocessing}
+% \end{figure*}
+
+\textcolor{red}{To be changed...}
+This paper is structured as follows.
+In Section \ref{sec:methods} we detail our methods.
+Especially, Section \ref{sec:dataset_generation} discusses the methodology used for generating the dataset for our analysis by extracting galaxy data from simulations.
+Section \ref{sec: methods_PCA_on_images} outlines the application of PCA to the image data and the resulting lower-dimensional image space (Section \ref{sec: dimensionality reduction method using pca}).
+Our resulting model is presented in Section \ref{sec: Results} and its accuracy thoroughly tested in Section \ref{sec: model evaluation, reconstruction error}.
+In Section \ref{sec: application} we highlight potential applications of our PCA galaxy model and conclude in Section \ref{sec: conclusion} with a summary and outlook of our results.
+
+Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/ufuk-cakir/MEGS}}}and refer to Appendix \ref{sec:appendix_code_and_data} and Fig.
+\ref{fig: hdf5_file_struc} for an overview of our code and file structure.
+Our final dataset is publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}
+
+%--------------------------------------------------------------------
+\section{Methods}
+\label{sec:methods}
+
+\paragraph{Simulation-based inference}
+In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
+Once this mapping is trained, we can apply it to real observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
+\ref{fig:flowchart}).
+For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
+
+\begin{figure}[]
+    \centering
+    \vspace{-.25cm}
+    \includegraphics[width=1\linewidth]{figures/sbi.png}
+    \vspace{-.7cm}
+    \caption{SBI flow chart.
+From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} which we use to train a Neural Density Estimator.
+With this we infer the posterior distribution of the model parameters from a single star.
+Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
+    \label{fig:flowchart}
+\end{figure}
+
+\paragraph{Galactic chemical evolution models}
+Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}.
+\texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics.
+We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.
+%
+In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
+\vspace{-1em}
+%\setlist[1]{itemsep=-4pt}
+\begin{enumerate}
+    \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$.
+We treat these as star-independent and assume them to be constant across galactic environments and cosmic time.
+\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,2019A&A...626A.124M}.} We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).
+    \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
+As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, which controls the peak of the star formation rate (SFR), and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$.
+We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
+    \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
+\end{enumerate}
+
+%The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
+This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk (which offsets the metallicity for different galactocentric radii).
+Similarly \cite{2019ApJ...874..102W} find that ISM parameter variations are deprojected in the [X/Mg] vs [Mg/H] plane (their Fig.\,17) and that abundance tracks in that space are independent of the stellar sample's spatial position within the Galaxy (their Fig.\,3).
+
+Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step.
+This corresponds to the constraint $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)>0.294$.
+For this reason, a truncated Normal prior will be used for the SFR parameter.
+Furthermore, we constrain $T_i$ to the interval $[1,13.8]$\,Gyr (assuming an age of the Universe of 13.8\,Gyr), ignoring any stars formed before $1$\,Gyr, which is justified as these are expected to be rare.
+%
+\begin{tiny}
+\begin{table*}
+\begin{minipage}{\textwidth}
+\begin{center}
+\caption{Free \texttt{Chempy} parameters for each star, with their prior values and Gaussian widths.
+Stellar birth-times are set for each star individually from a Uniform prior, based on realistic age estimates.}
+\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|C{1.9cm}C{2.cm} }
+Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\
+
+ \hline
+  \multicolumn{4}{c}{$\vec{\Lambda}$: \textit{Global stellar (SSP) parameters}}\\
+\hline
+  $\alpha_\mathrm{IMF}$ & High-mass slope of the \citep{2003PASP..115..763C} IMF & $-2.3\pm0.3$ & \citep[Tab.\,1]{2003PASP..115..763C} \\
+
+  $\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.75\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\
+
+\hline
+  \multicolumn{4}{c}{$\vec{\Theta}_i$: \textit{Local ISM parameters}}\\
+
+\hline
+  $\log_{10}\left(\mathrm{SFE}\right)$ & Star formation efficiency governing gas infall & $-0.3\pm0.3$ & \citep{2008AJ....136.2846B}\\
+
+  $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ & SFR peak in Gyr (scale of $k=2$ $\Gamma$-distribution) & $0.55\pm0.1$ & \citep[fig.\,4b]{2013ApJ...771L..35V}\\
+
+  x$_\mathrm{out}$ & Stellar feedback fraction & $\phantom{-}0.5\pm0.1$ & \citep[Tab.\,1]{Rybizki_2017}\\
+
+\hline
+ \multicolumn{4}{c}{$T_i$: \textit{Timescale}}\\
+
+\hline
+$T_i$ & Time of stellar birth in Gyr & [$1$,$13.8$] & Observations
+
+\label{tab:priors}
+\end{tabularx}
+\end{center}
+\end{minipage}
+\end{table*}
+\end{tiny}
+%
+We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
+%To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
+The utilized yields are summarized in Tab.\,\ref{tab:chempy_TNG_yields}, matching \citet[Tab.\,2]{2018MNRAS.473.4077P}, and we note that the SN\,II yields are renormalized such that the IMF-weighted yield ratios at each metallicity are equal to those from the \citet{2006ApJ...653.1145K} mass range models alone.
+\texttt{Chempy} uses only net yields, such that they provide only newly synthesized material, with the remainder coming from the initial SSP composition.
+These tables may not well-represent true stellar chemistry, and the effects of this are examined in Sec.\,\ref{subsec: mocks_wrong_yield} by performing inference using an alternative set of yields.
+For the analysis of observational data, we would want to use the most up-to-date yields, such as \citet{2016ApJ...825...26K} AGB yields, and carefully chose elements which are known to be well reproduced by our current models (e.g.
+shown by \citet{2019ApJ...874..102W,2019arXiv190806113G}), though this is not appropriate in our context.
+To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{1998A&A...334..505P} and do not allow for any `hypernovae' (in contrary to \citetalias{2018ApJ...861...40P}).
+%
+% \begin{table}[]
+% \caption{Nucleosynthetic yield tables used in this analysis, matching those of the TNG simulation \citep[Tab.\,2]{2018MNRAS.473.4077P}.}
+%     \centering
+%     \begin{tabular}{c|c}
+%       Type & Yield Table \\
+%        \hline
+%         SN\,Ia & \citet{1997NuPhA.621..467N}\\
+%         SN\,II & \citet{2006ApJ...653.1145K,portinari}\\
+%         AGB & \citet{2010MNRAS.403.1413K,2014MNRAS.437..195D};\\
+%         & \citet{2014ApJ...797...44F}
+%     \end{tabular}
+% \label{tab:chempy_TNG_yields}
+% \end{table}
+%
+%TNG only tracks nine elements in their analysis: C, Fe, H, He, Mg, N, Ne, O and Si, reporting the mass-fractions of each \citep{2018MNRAS.473.4077P}. 
+In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si and principally compare the logarithmic abundances [X/Fe] and [Fe/H] (defined by 
+%\begin{equation}
+    $[\mathrm{X}/\mathrm{Y}] = \log_{10}(N_\mathrm X/N_\mathrm Y)_\mathrm{star} - \log_{10}(N_\mathrm X/N_\mathrm Y)_\odot$
+%\end{equation}
+for number fraction $N_\mathrm X$ of element X), where $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
+This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}.%\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}
+%In this paper, \texttt{Chempy} will be used as the principal GCE model, which, with the modifications described above, allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
+It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{Chempy} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
+This is a necessary test to prepare for an inference on real data.
+
+\paragraph{Neural network emulator for \texttt{CHEMPY}:}
+Despite the simplifications made by the GCE model \texttt{CHEMPY}, we still have difficulties sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ due to the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space.
+To alleviate this, we use a \textit{neural network} emulator of the \texttt{CHEMPY} simulations with a total of $n_\mathrm{neuron}=140$ fully-connected neurons in two hidden layers.
+%
+% In essence, instead of computing the full model for each input parameter set, we pass the parameters to the network which predicts the output abundances to high accuracy. This has two benefits;
+% \begin{enumerate}
+%     \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow posterior sampling.
+With the neural network, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{Chempy}.
+%     \item \textbf{Differentiablility:} The neural network has a simple closed-form analytic structure (described in appendix \ref{appen: neural_network}), unlike the complex \texttt{Chempy} model.
+This allows it to be differentiated, so we can sample via advanced methods (cf.\,Sec.\,\ref{sec: sampling}).
+% \end{enumerate}
+%
+Despite the additional complexity introduced by using multiple stellar data-points, our network simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
+The same network can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
+With the above choices, the network predicts abundances with an average error of $0.008_{-0.005}^{+0.007}$\,dex, which is far below typical observational errors and even smaller away from the extremes of parameter space.
+All our code to reproduce our results is publicly avaialble on \href{https://github.com/bGuenes/sbi_chemical_abundances}{github}.
+
+%The larger volume of data should be able to give tighter statistical constraints on those parameters that are held fixed across the galaxy, but complexity is added since we must allow each star to carry its own set of local ISM parameters.
+
+% Let's see how far we get with our work before the deadline...
+
+%We will test our analysis using mock observations drawn firstly from \texttt{Chempy} then from large-scale hydrodynamical simulations to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
+The methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
+
+\section{Results}
+\label{sec: Results}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/sbi_Nstar_comp.png}
+    \vspace{-.25cm}
+    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) \cite{Philcox_2019} .}
+    \label{fig:N_star_analysis}
+\end{figure}
+
+We train the NPE with a set of $n_{\rm stars}=10^6$ abundances for $n_{\rm elemets}=8$ with an assumed observational error of $0.05$ dex, generated with the \textit{neural network} emulator, with inputs sampled from the priors shown in Tab.
+\ref{tab:priors}.
+The sole purpose of the emulator is to highly reduce the computational time compared to full \texttt{CHEMPY} simulations.
+Training the NPE takes $~3$h on a single H100 GPU.
+
+\paragraph{Mock Observational Data}
+Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
+\ref{tab:priors}, additionally drawing $T_i$ uniformly in the range $[2,12.8]$ Gyr, to minimize overlap with the neural network training birth-time limits when observational uncertainties are included.
+Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
+The outcome of this mock data creation is a set of 1000 stars, emulating a real dataset.
+
+\paragraph{Inference}
+\begin{wrapfigure}[17]{r}{0.5\textwidth}
+\vspace{-1.3cm}
+    \centering
+    \includegraphics[width=0.49\textwidth]{figures/posterior_2.png}
+    \vspace{-.25cm}
+    \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
+    \label{fig:sbi}   
+\end{wrapfigure}
+\enlargethispage{0\baselineskip}
+For each set of mock abundances we infer the posterior distribution for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with the NPE by sampling 1000 times and fitting the posterior for each parameter.
+This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
+Our method takes around $5$ minutes to build a posterior for all six parameters for a dataset size of 1000 stars on a Macbook Air M2 chip, making it around $2400$x faster than current HMC methods \citep[cf.][who needs $40$h for 200 stars with HMC]{Philcox_2019}.
+Hence, shorter computing times make it feasible to use orders of magnitudes more observations.
+%
+Taking the means of multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$ (Fig.
+\ref{fig:N_star_analysis}).
+However, we also see that in the limit of very few stars (less than $\sim10$) SBI shows a larger bias than HMC results.
+But in the limit of large numbers of stars (few hundred) the accuracy and precision is superior compared to HMC.
+%
+To infer the global galactic parameters $\vec\Lambda$ of the stars as accurate as possible we take the mean of all inferred posteriors from $n_{\rm stars}=1000$ (Fig.
+\ref{fig:sbi}) resulting in $\alpha_{\rm IMF}=-2.292\pm0.003$ and $\log_{10}(\rm N_{Ia})=-2.883\pm0.005$ which deviates less than $0.4\%$ from the correct value.
+
+    % \textcolor{blue}{Number of stars is not limited.
+I used $1000$ stars with an inference time of around $5$ minutes, but accuracy doesn't improve much after $\sim200$ stars [see \ref{fig:N_star_analysis}].
+Philcox inferred $200$ stars in $\sim40$h}
+    % \textcolor{red}{awesome! that gives us a nice metric to quote speedup. can we specify the architecture we used? I.e. which GPU on the clyster did you use?}
+    % \textcolor{blue}{For training the sbi net i used compgpu11 and it took around 3h. The inference of the stars i did on my laptop (macbook air with a M2 chip) in 5 minutes}
+
+\section{Summary and conclusions}
+\label{sec: conclusion}
+
+\section{Conclusion}
+\vspace{-1em}
+Here we have explored innovative SBI methods to constrain important galactic parameters from the multi-dimensional observation of chemical abundances of stars.
+Our results show that SBI provides a super fast method, that beats previous inference algorithms by orders of magnitude, and delivers accurate and precise results.
+Hence, SBI enables faster and more accurate evaluation of observational data after a small initial computational investment of training a NPE ($\sim3$ h).
+Since evaluation of the NPE is only weakly dependent on the dimensionality of the input observations we expect almost perfect scaling to larger chemical abundances observations and even further improvements in the inference results with future high quality data from massive spectroscopic surveys.
+
+   \begin{enumerate}
+      \item bla bla
+      \end{enumerate}
+
+\begin{acknowledgements}
+      This project was made possible by funding from the Carl Zeiss Stiftung.
+      %Part of this work was supported by the German
+      %\emph{Deut\-sche For\-schungs\-ge\-mein\-schaft, DFG\/} project
+      %number Ts~17/2--1.
+\end{acknowledgements}
+
+% WARNING
+%-------------------------------------------------------------------
+% Please note that we have included the references to the file aa.dem in
+% order to compile it, but we ask you to:
+%
+% - use BibTeX with the regular commands:
+%   \bibliographystyle{aa} % style aa.bst
+%   \bibliography{Yourfile} % your references Yourfile.bib
+%
+% - join the .bib files when you upload your source files
+%-------------------------------------------------------------------
+
+\bibliographystyle{aa}
+\bibliography{bibtex/references}
+
+\begin{appendix}
+
+\section{Code and Data Availability}
+\label{sec:appendix_code_and_data}
+The source code for ...  is made available under an open source license and is publicly hosted on GitHub.\\
+To facilitate a wider community's usage and contributions, we have ensured that the repository is well-documented.
+The repository includes comprehensive documentation hosted on \emph{ReadTheDocs} that provides an overview of the project, installation instructions, and a guide on how to use the software.
+
+The ... repository, which contains code, additional documentation, and interactive dashboards, is available on GitHub.\footnote{URL: \url{https://github.com/bGuenes/sbi_chemical_abundances}}
+
+\textcolor{red}{To be finalized, do we want to upload data???}
+The cleaned data set with outliers removed is publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}
+
+   \end{appendix}
 \end{document}
+
+%%%% End of aa.dem

### 2024-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{txfonts}
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
 \usepackage{siunitx}
 \usepackage{hyperref}
+\usepackage{url}
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{tabularx}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Stellar birth-times are set for each star individually from a Uniform prior, based on realistic age estimates.}
-\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|C{1.9cm}C{2.cm} }
+\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c{1.9cm}c{2.cm} }
 Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\
@@
 \paragraph{Inference}
-\begin{wrapfigure}[17]{r}{0.5\textwidth}
+\begin{figure}
 \vspace{-1.3cm}
@@
     \label{fig:sbi}   
-\end{wrapfigure}
+\end{figure}
 \enlargethispage{0\baselineskip}

### 2024-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure}[]
-    \centering
-    \vspace{-.25cm}
-    \includegraphics[width=1\linewidth]{figures/sbi.png}
-    \vspace{-.7cm}
-    \caption{SBI flow chart.
+% \begin{figure}[]
+%     \centering
+%     \vspace{-.25cm}
+%     \includegraphics[width=1\linewidth]{figures/sbi.png}
+%     \vspace{-.7cm}
+%     \caption{SBI flow chart.
 From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} which we use to train a Neural Density Estimator.
@@
 Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
-    \label{fig:flowchart}
-\end{figure}
+%     \label{fig:flowchart}
+% \end{figure}
 
@@
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9\linewidth]{figures/sbi_Nstar_comp.png}
-    \vspace{-.25cm}
-    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) \cite{Philcox_2019} .}
-    \label{fig:N_star_analysis}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.9\linewidth]{figures/sbi_Nstar_comp.png}
+%     \vspace{-.25cm}
+%     \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) \cite{Philcox_2019} .}
+%     \label{fig:N_star_analysis}
+% \end{figure}
 
@@
 \paragraph{Inference}
-\begin{figure}
-\vspace{-1.3cm}
-    \centering
-    \includegraphics[width=0.49\textwidth]{figures/posterior_2.png}
-    \vspace{-.25cm}
-    \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
-    \label{fig:sbi}   
-\end{figure}
-\enlargethispage{0\baselineskip}
+
+% \begin{figure}
+% \vspace{-1.3cm}
+%     \centering
+%     \includegraphics[width=0.49\textwidth]{figures/posterior_2.png}
+%     \vspace{-.25cm}
+%     \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
+%     \label{fig:sbi}   
+% \end{figure}
+
 For each set of mock abundances we infer the posterior distribution for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with the NPE by sampling 1000 times and fitting the posterior for each parameter.

### 2024-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
+
 \begin{tiny}
@@
 Stellar birth-times are set for each star individually from a Uniform prior, based on realistic age estimates.}
-\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c{1.9cm}c{2.cm} }
+\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c c }
 Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\
@@
 \end{tiny}
+
 %

### 2024-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\section{Motivation}
-
-Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars.
+% Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars.
 Two crucial unknowns are the shape of the initial mass function (IMF), setting the mass distribution of stars born from the interstellar medium (ISM), and the rate of Type Ia supernovae (SN\,Ia) explosions.
-%
-The exact values of parameters sensitively affect chemical evolution tracks \citep{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, yet their observational constraints remain weak.
-%
-For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,2015MNRAS.447.3880R,2014ApJ...796...75C}.
+% %
+% The exact values of parameters sensitively affect chemical evolution tracks \citep{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, yet their observational constraints remain weak.
+% %
+% For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,2015MNRAS.447.3880R,2014ApJ...796...75C}.
 In addition, the IMF slope may itself be not a constant but rather a function of metallicity, introducing further complexity \citep[e.g.][]{2019MNRAS.482..118G,2019A&A...626A.124M}.
@@
 
-The goal of this work is to demonstrate how we can use modern machine learning techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
+% The goal of this work is to demonstrate how we can use modern machine learning techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
 Our work focuses on two key parameters; the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF and the rate of SN\,Ia explosions per unit mass.
 We assume that both are constant across the galaxy and time.
-%
-Our method will make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
+% %
+% Our method will make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
 For high-dimensional posterior functions, SBI gives tremendous computational improvements over more traditional methods such as Markov Chain Monte Carlo (MCMC) or even Hamilton Monte Carlo (HMC) methods.
 
-% \begin{figure*}
-%    %\sidecaption
-%    \subfigure[]{
-%        \includegraphics[width=0.49\hsize, trim={0 0 0 0cm},clip]{figures/data/sample_galaxies_with_colorbar.pdf}
-%    }
-%    \subfigure[]{
-%       \includegraphics[width=0.51\hsize, trim={0 0 2.3cm 0},clip]{figures/data/galaxy_grid_with_colorbar.pdf}
-%    }
-%    \caption{Galaxy images (upper panels) and image cubes (lower panels) in our dataset after the preprocessing steps as explained in Section \ref{sec:dataset_generation}.
+% % \begin{figure*}
+% %    %\sidecaption
+% %    \subfigure[]{
+% %        \includegraphics[width=0.49\hsize, trim={0 0 0 0cm},clip]{figures/data/sample_galaxies_with_colorbar.pdf}
+% %    }
+% %    \subfigure[]{
+% %       \includegraphics[width=0.51\hsize, trim={0 0 2.3cm 0},clip]{figures/data/galaxy_grid_with_colorbar.pdf}
+% %    }
+% %    \caption{Galaxy images (upper panels) and image cubes (lower panels) in our dataset after the preprocessing steps as explained in Section \ref{sec:dataset_generation}.
 Each row contains one sample galaxy in three stellar maps - metallicity, age and mass from left to right in a) 2D and b) 3D.
@@
 The three dimensional plots have been created using the \emph{plotly} python package.}
-%    \label{fig: sample_galaxies_after_prepocessing}
-% \end{figure*}
+% %    \label{fig: sample_galaxies_after_prepocessing}
+% % \end{figure*}
 
-\textcolor{red}{To be changed...}
-This paper is structured as follows.
+% \textcolor{red}{To be changed...}
+% This paper is structured as follows.
 In Section \ref{sec:methods} we detail our methods.
@@
 
-Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/ufuk-cakir/MEGS}}}and refer to Appendix \ref{sec:appendix_code_and_data} and Fig.
+% Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/ufuk-cakir/MEGS}}}and refer to Appendix \ref{sec:appendix_code_and_data} and Fig.
 \ref{fig: hdf5_file_struc} for an overview of our code and file structure.
@@
 
-%--------------------------------------------------------------------
-\section{Methods}
-\label{sec:methods}
+% %--------------------------------------------------------------------
+% \section{Methods}
+% \label{sec:methods}
 
-\paragraph{Simulation-based inference}
-In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
+% \paragraph{Simulation-based inference}
+% In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
 Once this mapping is trained, we can apply it to real observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
 \ref{fig:flowchart}).
-For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
+% For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
 
-% \begin{figure}[]
-%     \centering
-%     \vspace{-.25cm}
-%     \includegraphics[width=1\linewidth]{figures/sbi.png}
-%     \vspace{-.7cm}
-%     \caption{SBI flow chart.
+% % \begin{figure}[]
+% %     \centering
+% %     \vspace{-.25cm}
+% %     \includegraphics[width=1\linewidth]{figures/sbi.png}
+% %     \vspace{-.7cm}
+% %     \caption{SBI flow chart.
 From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} which we use to train a Neural Density Estimator.
@@
 Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
-%     \label{fig:flowchart}
-% \end{figure}
+% %     \label{fig:flowchart}
+% % \end{figure}
 
-\paragraph{Galactic chemical evolution models}
-Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}.
+% \paragraph{Galactic chemical evolution models}
+% Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}.
 \texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics.
 We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.
-%
-In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
-\vspace{-1em}
-%\setlist[1]{itemsep=-4pt}
-\begin{enumerate}
-    \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$.
+% %
+% In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
+% \vspace{-1em}
+% %\setlist[1]{itemsep=-4pt}
+% \begin{enumerate}
+%     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$.
 We treat these as star-independent and assume them to be constant across galactic environments and cosmic time.
 \footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,2019A&A...626A.124M}.} We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).
-    \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
+%     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
 As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, which controls the peak of the star formation rate (SFR), and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$.
 We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
-    \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
-\end{enumerate}
+%     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
+% \end{enumerate}
 
-%The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
+% %The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
 This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk (which offsets the metallicity for different galactocentric radii).
@@
 
-Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step.
+% Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step.
 This corresponds to the constraint $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)>0.294$.
@@
 Furthermore, we constrain $T_i$ to the interval $[1,13.8]$\,Gyr (assuming an age of the Universe of 13.8\,Gyr), ignoring any stars formed before $1$\,Gyr, which is justified as these are expected to be rare.
-%
+% %
 
-\begin{tiny}
-\begin{table*}
-\begin{minipage}{\textwidth}
-\begin{center}
-\caption{Free \texttt{Chempy} parameters for each star, with their prior values and Gaussian widths.
+% \begin{tiny}
+% \begin{table*}
+% \begin{minipage}{\textwidth}
+% \begin{center}
+% \caption{Free \texttt{Chempy} parameters for each star, with their prior values and Gaussian widths.
 Stellar birth-times are set for each star individually from a Uniform prior, based on realistic age estimates.}
-\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c c }
-Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\
+% \begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c c }
+% Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\
 
- \hline
-  \multicolumn{4}{c}{$\vec{\Lambda}$: \textit{Global stellar (SSP) parameters}}\\
-\hline
-  $\alpha_\mathrm{IMF}$ & High-mass slope of the \citep{2003PASP..115..763C} IMF & $-2.3\pm0.3$ & \citep[Tab.\,1]{2003PASP..115..763C} \\
+%  \hline
+%   \multicolumn{4}{c}{$\vec{\Lambda}$: \textit{Global stellar (SSP) parameters}}\\
+% \hline
+%   $\alpha_\mathrm{IMF}$ & High-mass slope of the \citep{2003PASP..115..763C} IMF & $-2.3\pm0.3$ & \citep[Tab.\,1]{2003PASP..115..763C} \\
 
-  $\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.75\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\
+%   $\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.75\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\
 
-\hline
-  \multicolumn{4}{c}{$\vec{\Theta}_i$: \textit{Local ISM parameters}}\\
+% \hline
+%   \multicolumn{4}{c}{$\vec{\Theta}_i$: \textit{Local ISM parameters}}\\
 
-\hline
-  $\log_{10}\left(\mathrm{SFE}\right)$ & Star formation efficiency governing gas infall & $-0.3\pm0.3$ & \citep{2008AJ....136.2846B}\\
+% \hline
+%   $\log_{10}\left(\mathrm{SFE}\right)$ & Star formation efficiency governing gas infall & $-0.3\pm0.3$ & \citep{2008AJ....136.2846B}\\
 
-  $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ & SFR peak in Gyr (scale of $k=2$ $\Gamma$-distribution) & $0.55\pm0.1$ & \citep[fig.\,4b]{2013ApJ...771L..35V}\\
+%   $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ & SFR peak in Gyr (scale of $k=2$ $\Gamma$-distribution) & $0.55\pm0.1$ & \citep[fig.\,4b]{2013ApJ...771L..35V}\\
 
-  x$_\mathrm{out}$ & Stellar feedback fraction & $\phantom{-}0.5\pm0.1$ & \citep[Tab.\,1]{Rybizki_2017}\\
+%   x$_\mathrm{out}$ & Stellar feedback fraction & $\phantom{-}0.5\pm0.1$ & \citep[Tab.\,1]{Rybizki_2017}\\
 
-\hline
- \multicolumn{4}{c}{$T_i$: \textit{Timescale}}\\
+% \hline
+%  \multicolumn{4}{c}{$T_i$: \textit{Timescale}}\\
 
-\hline
-$T_i$ & Time of stellar birth in Gyr & [$1$,$13.8$] & Observations
+% \hline
+% $T_i$ & Time of stellar birth in Gyr & [$1$,$13.8$] & Observations
 
-\label{tab:priors}
-\end{tabularx}
-\end{center}
-\end{minipage}
-\end{table*}
-\end{tiny}
+% \label{tab:priors}
+% \end{tabularx}
+% \end{center}
+% \end{minipage}
+% \end{table*}
+% \end{tiny}
 
-%
-We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
-%To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
+% %
+% We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
+% %To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
 The utilized yields are summarized in Tab.\,\ref{tab:chempy_TNG_yields}, matching \citet[Tab.\,2]{2018MNRAS.473.4077P}, and we note that the SN\,II yields are renormalized such that the IMF-weighted yield ratios at each metallicity are equal to those from the \citet{2006ApJ...653.1145K} mass range models alone.
@@
 To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{1998A&A...334..505P} and do not allow for any `hypernovae' (in contrary to \citetalias{2018ApJ...861...40P}).
-%
-% \begin{table}[]
-% \caption{Nucleosynthetic yield tables used in this analysis, matching those of the TNG simulation \citep[Tab.\,2]{2018MNRAS.473.4077P}.}
-%     \centering
-%     \begin{tabular}{c|c}
-%       Type & Yield Table \\
-%        \hline
-%         SN\,Ia & \citet{1997NuPhA.621..467N}\\
-%         SN\,II & \citet{2006ApJ...653.1145K,portinari}\\
-%         AGB & \citet{2010MNRAS.403.1413K,2014MNRAS.437..195D};\\
-%         & \citet{2014ApJ...797...44F}
-%     \end{tabular}
-% \label{tab:chempy_TNG_yields}
-% \end{table}
-%
-%TNG only tracks nine elements in their analysis: C, Fe, H, He, Mg, N, Ne, O and Si, reporting the mass-fractions of each \citep{2018MNRAS.473.4077P}. 
-In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si and principally compare the logarithmic abundances [X/Fe] and [Fe/H] (defined by 
-%\begin{equation}
-    $[\mathrm{X}/\mathrm{Y}] = \log_{10}(N_\mathrm X/N_\mathrm Y)_\mathrm{star} - \log_{10}(N_\mathrm X/N_\mathrm Y)_\odot$
-%\end{equation}
-for number fraction $N_\mathrm X$ of element X), where $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
+% %
+% % \begin{table}[]
+% % \caption{Nucleosynthetic yield tables used in this analysis, matching those of the TNG simulation \citep[Tab.\,2]{2018MNRAS.473.4077P}.}
+% %     \centering
+% %     \begin{tabular}{c|c}
+% %       Type & Yield Table \\
+% %        \hline
+% %         SN\,Ia & \citet{1997NuPhA.621..467N}\\
+% %         SN\,II & \citet{2006ApJ...653.1145K,portinari}\\
+% %         AGB & \citet{2010MNRAS.403.1413K,2014MNRAS.437..195D};\\
+% %         & \citet{2014ApJ...797...44F}
+% %     \end{tabular}
+% % \label{tab:chempy_TNG_yields}
+% % \end{table}
+% %
+% %TNG only tracks nine elements in their analysis: C, Fe, H, He, Mg, N, Ne, O and Si, reporting the mass-fractions of each \citep{2018MNRAS.473.4077P}. 
+% In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si and principally compare the logarithmic abundances [X/Fe] and [Fe/H] (defined by 
+% %\begin{equation}
+%     $[\mathrm{X}/\mathrm{Y}] = \log_{10}(N_\mathrm X/N_\mathrm Y)_\mathrm{star} - \log_{10}(N_\mathrm X/N_\mathrm Y)_\odot$
+% %\end{equation}
+% for number fraction $N_\mathrm X$ of element X), where $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
 This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}.%\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}
-%In this paper, \texttt{Chempy} will be used as the principal GCE model, which, with the modifications described above, allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
+% %In this paper, \texttt{Chempy} will be used as the principal GCE model, which, with the modifications described above, allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
 It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{Chempy} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
@@
 
-\paragraph{Neural network emulator for \texttt{CHEMPY}:}
-Despite the simplifications made by the GCE model \texttt{CHEMPY}, we still have difficulties sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ due to the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space.
+% \paragraph{Neural network emulator for \texttt{CHEMPY}:}
+% Despite the simplifications made by the GCE model \texttt{CHEMPY}, we still have difficulties sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ due to the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space.
 To alleviate this, we use a \textit{neural network} emulator of the \texttt{CHEMPY} simulations with a total of $n_\mathrm{neuron}=140$ fully-connected neurons in two hidden layers.
-%
-% In essence, instead of computing the full model for each input parameter set, we pass the parameters to the network which predicts the output abundances to high accuracy. This has two benefits;
-% \begin{enumerate}
-%     \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow posterior sampling.
+% %
+% % In essence, instead of computing the full model for each input parameter set, we pass the parameters to the network which predicts the output abundances to high accuracy. This has two benefits;
+% % \begin{enumerate}
+% %     \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow posterior sampling.
 With the neural network, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{Chempy}.
-%     \item \textbf{Differentiablility:} The neural network has a simple closed-form analytic structure (described in appendix \ref{appen: neural_network}), unlike the complex \texttt{Chempy} model.
+% %     \item \textbf{Differentiablility:} The neural network has a simple closed-form analytic structure (described in appendix \ref{appen: neural_network}), unlike the complex \texttt{Chempy} model.
 This allows it to be differentiated, so we can sample via advanced methods (cf.\,Sec.\,\ref{sec: sampling}).
-% \end{enumerate}
-%
-Despite the additional complexity introduced by using multiple stellar data-points, our network simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
+% % \end{enumerate}
+% %
+% Despite the additional complexity introduced by using multiple stellar data-points, our network simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
 The same network can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
 With the above choices, the network predicts abundances with an average error of $0.008_{-0.005}^{+0.007}$\,dex, which is far below typical observational errors and even smaller away from the extremes of parameter space.
-All our code to reproduce our results is publicly avaialble on \href{https://github.com/bGuenes/sbi_chemical_abundances}{github}.
+% All our code to reproduce our results is publicly avaialble on \href{https://github.com/bGuenes/sbi_chemical_abundances}{github}.
 
-%The larger volume of data should be able to give tighter statistical constraints on those parameters that are held fixed across the galaxy, but complexity is added since we must allow each star to carry its own set of local ISM parameters.
+% %The larger volume of data should be able to give tighter statistical constraints on those parameters that are held fixed across the galaxy, but complexity is added since we must allow each star to carry its own set of local ISM parameters.
 
-% Let's see how far we get with our work before the deadline...
+% % Let's see how far we get with our work before the deadline...
 
-%We will test our analysis using mock observations drawn firstly from \texttt{Chempy} then from large-scale hydrodynamical simulations to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
+% %We will test our analysis using mock observations drawn firstly from \texttt{Chempy} then from large-scale hydrodynamical simulations to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
 The methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
 
-\section{Results}
-\label{sec: Results}
+% \section{Results}
+% \label{sec: Results}
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=0.9\linewidth]{figures/sbi_Nstar_comp.png}
-%     \vspace{-.25cm}
-%     \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) \cite{Philcox_2019} .}
-%     \label{fig:N_star_analysis}
-% \end{figure}
+% % \begin{figure}
+% %     \centering
+% %     \includegraphics[width=0.9\linewidth]{figures/sbi_Nstar_comp.png}
+% %     \vspace{-.25cm}
+% %     \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) \cite{Philcox_2019} .}
+% %     \label{fig:N_star_analysis}
+% % \end{figure}
 
-We train the NPE with a set of $n_{\rm stars}=10^6$ abundances for $n_{\rm elemets}=8$ with an assumed observational error of $0.05$ dex, generated with the \textit{neural network} emulator, with inputs sampled from the priors shown in Tab.
+% We train the NPE with a set of $n_{\rm stars}=10^6$ abundances for $n_{\rm elemets}=8$ with an assumed observational error of $0.05$ dex, generated with the \textit{neural network} emulator, with inputs sampled from the priors shown in Tab.
 \ref{tab:priors}.
@@
 
-\paragraph{Mock Observational Data}
-Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
+% \paragraph{Mock Observational Data}
+% Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
 \ref{tab:priors}, additionally drawing $T_i$ uniformly in the range $[2,12.8]$ Gyr, to minimize overlap with the neural network training birth-time limits when observational uncertainties are included.
-Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
-The outcome of this mock data creation is a set of 1000 stars, emulating a real dataset.
+% Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
+% The outcome of this mock data creation is a set of 1000 stars, emulating a real dataset.
 
-\paragraph{Inference}
+% \paragraph{Inference}
 
-% \begin{figure}
-% \vspace{-1.3cm}
-%     \centering
-%     \includegraphics[width=0.49\textwidth]{figures/posterior_2.png}
-%     \vspace{-.25cm}
-%     \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
-%     \label{fig:sbi}   
-% \end{figure}
+% % \begin{figure}
+% % \vspace{-1.3cm}
+% %     \centering
+% %     \includegraphics[width=0.49\textwidth]{figures/posterior_2.png}
+% %     \vspace{-.25cm}
+% %     \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
+% %     \label{fig:sbi}   
+% % \end{figure}
 
-For each set of mock abundances we infer the posterior distribution for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with the NPE by sampling 1000 times and fitting the posterior for each parameter.
+% For each set of mock abundances we infer the posterior distribution for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with the NPE by sampling 1000 times and fitting the posterior for each parameter.
 This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
-Our method takes around $5$ minutes to build a posterior for all six parameters for a dataset size of 1000 stars on a Macbook Air M2 chip, making it around $2400$x faster than current HMC methods \citep[cf.][who needs $40$h for 200 stars with HMC]{Philcox_2019}.
+% Our method takes around $5$ minutes to build a posterior for all six parameters for a dataset size of 1000 stars on a Macbook Air M2 chip, making it around $2400$x faster than current HMC methods \citep[cf.][who needs $40$h for 200 stars with HMC]{Philcox_2019}.
 Hence, shorter computing times make it feasible to use orders of magnitudes more observations.
-%
-Taking the means of multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$ (Fig.
+% %
+% Taking the means of multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$ (Fig.
 \ref{fig:N_star_analysis}).
@@
 But in the limit of large numbers of stars (few hundred) the accuracy and precision is superior compared to HMC.
-%
-To infer the global galactic parameters $\vec\Lambda$ of the stars as accurate as possible we take the mean of all inferred posteriors from $n_{\rm stars}=1000$ (Fig.
+% %
+% To infer the global galactic parameters $\vec\Lambda$ of the stars as accurate as possible we take the mean of all inferred posteriors from $n_{\rm stars}=1000$ (Fig.
 \ref{fig:sbi}) resulting in $\alpha_{\rm IMF}=-2.292\pm0.003$ and $\log_{10}(\rm N_{Ia})=-2.883\pm0.005$ which deviates less than $0.4\%$ from the correct value.
 
-    % \textcolor{blue}{Number of stars is not limited.
+%     % \textcolor{blue}{Number of stars is not limited.
 I used $1000$ stars with an inference time of around $5$ minutes, but accuracy doesn't improve much after $\sim200$ stars [see \ref{fig:N_star_analysis}].
 Philcox inferred $200$ stars in $\sim40$h}
-    % \textcolor{red}{awesome! that gives us a nice metric to quote speedup. can we specify the architecture we used? I.e. which GPU on the clyster did you use?}
-    % \textcolor{blue}{For training the sbi net i used compgpu11 and it took around 3h. The inference of the stars i did on my laptop (macbook air with a M2 chip) in 5 minutes}
+%     % \textcolor{red}{awesome! that gives us a nice metric to quote speedup. can we specify the architecture we used? I.e. which GPU on the clyster did you use?}
+%     % \textcolor{blue}{For training the sbi net i used compgpu11 and it took around 3h. The inference of the stars i did on my laptop (macbook air with a M2 chip) in 5 minutes}
 
-\section{Summary and conclusions}
-\label{sec: conclusion}
+% \section{Summary and conclusions}
+% \label{sec: conclusion}
 
-\section{Conclusion}
-\vspace{-1em}
-Here we have explored innovative SBI methods to constrain important galactic parameters from the multi-dimensional observation of chemical abundances of stars.
+% Here we have explored innovative SBI methods to constrain important galactic parameters from the multi-dimensional observation of chemical abundances of stars.
 Our results show that SBI provides a super fast method, that beats previous inference algorithms by orders of magnitude, and delivers accurate and precise results.
@@
 \bibliographystyle{aa}
-\bibliography{bibtex/references}
+\bibliography{bib.bib}
 

### 2024-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-% Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars.
+ Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars.
 Two crucial unknowns are the shape of the initial mass function (IMF), setting the mass distribution of stars born from the interstellar medium (ISM), and the rate of Type Ia supernovae (SN\,Ia) explosions.

### 2024-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{xcolor}
+
 \newcommand{\todo}[1]{\textcolor{red}{#1}}
@@
 \usepackage{subfigure}
+
 \begin{document}
@@
 
-\begin{appendix}
+% \begin{appendix}
 
-\section{Code and Data Availability}
-\label{sec:appendix_code_and_data}
-The source code for ...  is made available under an open source license and is publicly hosted on GitHub.\\
-To facilitate a wider community's usage and contributions, we have ensured that the repository is well-documented.
+% \section{Code and Data Availability}
+% \label{sec:appendix_code_and_data}
+% The source code for ...  is made available under an open source license and is publicly hosted on GitHub.\\
+% To facilitate a wider community's usage and contributions, we have ensured that the repository is well-documented.
 The repository includes comprehensive documentation hosted on \emph{ReadTheDocs} that provides an overview of the project, installation instructions, and a guide on how to use the software.
 
-The ... repository, which contains code, additional documentation, and interactive dashboards, is available on GitHub.\footnote{URL: \url{https://github.com/bGuenes/sbi_chemical_abundances}}
+% The ... repository, which contains code, additional documentation, and interactive dashboards, is available on GitHub.%\footnote{URL: \url{https://github.com/bGuenes/sbi_chemical_abundances}}
 
-\textcolor{red}{To be finalized, do we want to upload data???}
-The cleaned data set with outliers removed is publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}
+% \textcolor{red}{To be finalized, do we want to upload data???}
+% The cleaned data set with outliers removed is publicly available on Zenodo.%\footnote{URL: \url{https://zenodo.org/record/8375344}}
 
-   \end{appendix}
+%   \end{appendix}
+
 \end{document}

### 2024-12-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
+\usepackage{showyourwork}
+
 \usepackage{graphicx}
@@
  Universität Heidelberg, Zentrum für Astronomie, Institut für Theoretische Astrophysik, Albert-Ueberle-Straße 2, D-69120 Heidelberg, Germany\\
- \email{tobias.buck@iwr.uni-heidelberg.de},  \email{b.guenes@stud.uni-heidelberg.de}
+ \email{tobias.buck@iwr.uni-heidelberg.de}
              }
@@
 
- Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars.
+Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars or the production of heavy elements inside stars \citep[e.g.][]{Buck2021}.
 Two crucial unknowns are the shape of the initial mass function (IMF), setting the mass distribution of stars born from the interstellar medium (ISM), and the rate of Type Ia supernovae (SN\,Ia) explosions.
-% %
-% The exact values of parameters sensitively affect chemical evolution tracks \citep{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, yet their observational constraints remain weak.
-% %
-% For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,2015MNRAS.447.3880R,2014ApJ...796...75C}.
+
+The exact values of parameters sensitively affect chemical evolution tracks \citep{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, yet their observational constraints remain weak.
+%
+For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,2015MNRAS.447.3880R,2014ApJ...796...75C}.
 In addition, the IMF slope may itself be not a constant but rather a function of metallicity, introducing further complexity \citep[e.g.][]{2019MNRAS.482..118G,2019A&A...626A.124M}.
@@
 
-% The goal of this work is to demonstrate how we can use modern machine learning techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
+The goal of this work is to demonstrate how we can use modern machine learning techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
 Our work focuses on two key parameters; the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF and the rate of SN\,Ia explosions per unit mass.
 We assume that both are constant across the galaxy and time.
-% %
-% Our method will make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
+
+Our method will make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
 For high-dimensional posterior functions, SBI gives tremendous computational improvements over more traditional methods such as Markov Chain Monte Carlo (MCMC) or even Hamilton Monte Carlo (HMC) methods.
 
-% % \begin{figure*}
-% %    %\sidecaption
-% %    \subfigure[]{
-% %        \includegraphics[width=0.49\hsize, trim={0 0 0 0cm},clip]{figures/data/sample_galaxies_with_colorbar.pdf}
-% %    }
-% %    \subfigure[]{
-% %       \includegraphics[width=0.51\hsize, trim={0 0 2.3cm 0},clip]{figures/data/galaxy_grid_with_colorbar.pdf}
-% %    }
-% %    \caption{Galaxy images (upper panels) and image cubes (lower panels) in our dataset after the preprocessing steps as explained in Section \ref{sec:dataset_generation}.
-Each row contains one sample galaxy in three stellar maps - metallicity, age and mass from left to right in a) 2D and b) 3D.
-The images are normalised in the range $[0,1]$ and bright pixels correspond to high values.
-Note that in the stellar age maps, dark pixels correspond to young stars.
-The three dimensional plots have been created using the \emph{plotly} python package.}
-% %    \label{fig: sample_galaxies_after_prepocessing}
-% % \end{figure*}
-
-% \textcolor{red}{To be changed...}
-% This paper is structured as follows.
+\textcolor{red}{to be adapted...\\
+This paper is structured as follows.
 In Section \ref{sec:methods} we detail our methods.
@@
 Our resulting model is presented in Section \ref{sec: Results} and its accuracy thoroughly tested in Section \ref{sec: model evaluation, reconstruction error}.
-In Section \ref{sec: application} we highlight potential applications of our PCA galaxy model and conclude in Section \ref{sec: conclusion} with a summary and outlook of our results.
+In Section \ref{sec: application} we highlight potential applications of our PCA galaxy model and conclude in Section \ref{sec: conclusion} with a summary and outlook of our results.}
 
-% Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/ufuk-cakir/MEGS}}}and refer to Appendix \ref{sec:appendix_code_and_data} and Fig.
+\textcolor{red}{to be adapted...\\
+Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/ufuk-cakir/MEGS}}}and refer to Appendix \ref{sec:appendix_code_and_data} and Fig.
 \ref{fig: hdf5_file_struc} for an overview of our code and file structure.
-Our final dataset is publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}
+Our final dataset is publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}}
 

### 2024-12-13 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % %--------------------------------------------------------------------
-% \section{Methods}
-% \label{sec:methods}
+\section{Methods}
+\label{sec:methods}
 
-% \paragraph{Simulation-based inference}
-% In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
+\paragraph{Simulation-based inference}
+In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
 Once this mapping is trained, we can apply it to real observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
 \ref{fig:flowchart}).
-% For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
+For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
 
-% % \begin{figure}[]
-% %     \centering
-% %     \vspace{-.25cm}
-% %     \includegraphics[width=1\linewidth]{figures/sbi.png}
-% %     \vspace{-.7cm}
-% %     \caption{SBI flow chart.
+\begin{figure}[]
+     \centering
+     \vspace{-.25cm}
+     \includegraphics[width=1\linewidth]{figures/sbi_overview.png}
+     \vspace{-.7cm}
+     \caption{SBI flow chart.
 From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} which we use to train a Neural Density Estimator.
@@
 Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
-% %     \label{fig:flowchart}
-% % \end{figure}
+     \label{fig:flowchart}
+\end{figure}
 
-% \paragraph{Galactic chemical evolution models}
-% Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}.
+\paragraph{Galactic chemical evolution models}
+Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}.
 \texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics.
@@
 % %
-% In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
-% \vspace{-1em}
+In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
+\vspace{-1em}
 % %\setlist[1]{itemsep=-4pt}
-% \begin{enumerate}
-%     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$.
+\begin{enumerate}
+     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$.
 We treat these as star-independent and assume them to be constant across galactic environments and cosmic time.
 \footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,2019A&A...626A.124M}.} We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).
-%     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
+     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
 As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, which controls the peak of the star formation rate (SFR), and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$.
 We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
-%     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
-% \end{enumerate}
+     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
+\end{enumerate}
 

### 2024-12-13 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We treat these as star-independent and assume them to be constant across galactic environments and cosmic time.
-\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,2019A&A...626A.124M}.} We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).
+\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G}}%,2019A&A...626A.124M}.}
+     We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).  
      \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.

### 2024-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2024-12-13 src/tex/ms.tex
@@
    \title{Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference}
+
+   \titlerunning{Simulation-Based Inference for Galactic Parameters}
+   \authorrunning{T. Buck and B. Günes}
 
@@
 
-\textcolor{red}{to be adapted...\\
-Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/ufuk-cakir/MEGS}}}and refer to Appendix \ref{sec:appendix_code_and_data} and Fig.
-\ref{fig: hdf5_file_struc} for an overview of our code and file structure.
-Our final dataset is publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}}
+\textcolor{red}{to be adapted... especially the zenodo link\\
+Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/TobiBu/sbi-chempy}}}and refer to Appendix \ref{sec:appendix_code_and_data} for an overview of our code and structure.
+All our datasets and network weights are publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}}
+
+\begin{figure*}[]
+     \centering
+     %\vspace{-.25cm}
+     \includegraphics[width=1\linewidth]{figures/sbi_overview.png}
+     \vspace{-.5cm}
+     \caption{SBI flow chart.
+From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} \citep{Rybizki_2017,Philcox_2019} which we use to train a \emph{neural network} emulator to speed up the data generation process.
+Using the \emph{neural network} emulator we produce training data to train the Neural Density Estimator.
+With this we infer the posterior distribution of the model parameters from a single star.
+Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
+     \label{fig:flowchart}
+\end{figure*}
 
@@
 For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
-
-\begin{figure}[]
-     \centering
-     \vspace{-.25cm}
-     \includegraphics[width=1\linewidth]{figures/sbi_overview.png}
-     \vspace{-.7cm}
-     \caption{SBI flow chart.
-From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} which we use to train a Neural Density Estimator.
-With this we infer the posterior distribution of the model parameters from a single star.
-Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
-     \label{fig:flowchart}
-\end{figure}
 
@@
 We treat these as star-independent and assume them to be constant across galactic environments and cosmic time.
-\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,2019A&A...626A.124M}.} We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).
+\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,Martin2019}.}
+     We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).  
      \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
@@
 
-% %The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
+The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
 This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk (which offsets the metallicity for different galactocentric radii).
@@
 
-% Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step.
+Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step.
 This corresponds to the constraint $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)>0.294$.
@@
 Furthermore, we constrain $T_i$ to the interval $[1,13.8]$\,Gyr (assuming an age of the Universe of 13.8\,Gyr), ignoring any stars formed before $1$\,Gyr, which is justified as these are expected to be rare.
-% %
 
-% \begin{tiny}
-% \begin{table*}
-% \begin{minipage}{\textwidth}
-% \begin{center}
-% \caption{Free \texttt{Chempy} parameters for each star, with their prior values and Gaussian widths.
+\begin{tiny}
+\begin{table*}
+\begin{minipage}{\textwidth}
+\begin{center}
+\caption{Free \texttt{Chempy} parameters for each star, with their prior values and Gaussian widths.
 Stellar birth-times are set for each star individually from a Uniform prior, based on realistic age estimates.}
-% \begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c c }
-% Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\
+\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c c }
+Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\
 
-%  \hline
-%   \multicolumn{4}{c}{$\vec{\Lambda}$: \textit{Global stellar (SSP) parameters}}\\
-% \hline
-%   $\alpha_\mathrm{IMF}$ & High-mass slope of the \citep{2003PASP..115..763C} IMF & $-2.3\pm0.3$ & \citep[Tab.\,1]{2003PASP..115..763C} \\
+\hline
+\multicolumn{4}{c}{$\vec{\Lambda}$: \textit{Global stellar (SSP) parameters}}\\
+\hline
+$\alpha_\mathrm{IMF}$ & High-mass slope of the \citep{2003PASP..115..763C} IMF & $-2.3\pm0.3$ & \citep[Tab.\,1]{2003PASP..115..763C} \\
 
-%   $\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.75\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\
+$\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.75\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\
 
-% \hline
-%   \multicolumn{4}{c}{$\vec{\Theta}_i$: \textit{Local ISM parameters}}\\
+\hline
+\multicolumn{4}{c}{$\vec{\Theta}_i$: \textit{Local ISM parameters}}\\
 
-% \hline
-%   $\log_{10}\left(\mathrm{SFE}\right)$ & Star formation efficiency governing gas infall & $-0.3\pm0.3$ & \citep{2008AJ....136.2846B}\\
+\hline
+$\log_{10}\left(\mathrm{SFE}\right)$ & Star formation efficiency governing gas infall & $-0.3\pm0.3$ & \citep{2008AJ....136.2846B}\\
 
-%   $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ & SFR peak in Gyr (scale of $k=2$ $\Gamma$-distribution) & $0.55\pm0.1$ & \citep[fig.\,4b]{2013ApJ...771L..35V}\\
+$\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ & SFR peak in Gyr (scale of $k=2$ $\Gamma$-distribution) & $0.55\pm0.1$ & \citep[fig.\,4b]{2013ApJ...771L..35V}\\
 
-%   x$_\mathrm{out}$ & Stellar feedback fraction & $\phantom{-}0.5\pm0.1$ & \citep[Tab.\,1]{Rybizki_2017}\\
+x$_\mathrm{out}$ & Stellar feedback fraction & $\phantom{-}0.5\pm0.1$ & \citep[Tab.\,1]{Rybizki_2017}\\
 
-% \hline
-%  \multicolumn{4}{c}{$T_i$: \textit{Timescale}}\\
+\hline
+\multicolumn{4}{c}{$T_i$: \textit{Timescale}}\\
 
-% \hline
-% $T_i$ & Time of stellar birth in Gyr & [$1$,$13.8$] & Observations
+\hline
+$T_i$ & Time of stellar birth in Gyr & [$1$,$13.8$] & Observations
 
-% \label{tab:priors}
-% \end{tabularx}
-% \end{center}
-% \end{minipage}
-% \end{table*}
-% \end{tiny}
+\label{tab:priors}
+\end{tabularx}
+\end{center}
+\end{minipage}
+\end{table*}
+\end{tiny}
 
-% %
-% We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
-% %To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
+We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
+To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
 The utilized yields are summarized in Tab.\,\ref{tab:chempy_TNG_yields}, matching \citet[Tab.\,2]{2018MNRAS.473.4077P}, and we note that the SN\,II yields are renormalized such that the IMF-weighted yield ratios at each metallicity are equal to those from the \citet{2006ApJ...653.1145K} mass range models alone.
@@
 To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{1998A&A...334..505P} and do not allow for any `hypernovae' (in contrary to \citetalias{2018ApJ...861...40P}).
-% %
-% % \begin{table}[]
-% % \caption{Nucleosynthetic yield tables used in this analysis, matching those of the TNG simulation \citep[Tab.\,2]{2018MNRAS.473.4077P}.}
-% %     \centering
-% %     \begin{tabular}{c|c}
-% %       Type & Yield Table \\
-% %        \hline
-% %         SN\,Ia & \citet{1997NuPhA.621..467N}\\
-% %         SN\,II & \citet{2006ApJ...653.1145K,portinari}\\
-% %         AGB & \citet{2010MNRAS.403.1413K,2014MNRAS.437..195D};\\
-% %         & \citet{2014ApJ...797...44F}
-% %     \end{tabular}
-% % \label{tab:chempy_TNG_yields}
-% % \end{table}
-% %
-% %TNG only tracks nine elements in their analysis: C, Fe, H, He, Mg, N, Ne, O and Si, reporting the mass-fractions of each \citep{2018MNRAS.473.4077P}. 
-% In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si and principally compare the logarithmic abundances [X/Fe] and [Fe/H] (defined by 
-% %\begin{equation}
-%     $[\mathrm{X}/\mathrm{Y}] = \log_{10}(N_\mathrm X/N_\mathrm Y)_\mathrm{star} - \log_{10}(N_\mathrm X/N_\mathrm Y)_\odot$
-% %\end{equation}
-% for number fraction $N_\mathrm X$ of element X), where $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
-This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}.%\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}
-% %In this paper, \texttt{Chempy} will be used as the principal GCE model, which, with the modifications described above, allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
+
+\begin{table}[]
+\caption{Nucleosynthetic yield tables used in this analysis, matching those of the TNG simulation \citep[Tab.\,2]{2018MNRAS.473.4077P}.}
+     \centering
+     \begin{tabular}{c|c}
+       Type & Yield Table \\
+        \hline
+         SN\,Ia & \citet{1997NuPhA.621..467N}\\
+         SN\,II & \citet{2006ApJ...653.1145K,portinari}\\
+         AGB & \citet{2010MNRAS.403.1413K,2014MNRAS.437..195D};\\
+         & \citet{2014ApJ...797...44F}
+     \end{tabular}
+ \label{tab:chempy_TNG_yields}
+ \end{table}
+
+%TNG only tracks nine elements in their analysis: C, Fe, H, He, Mg, N, Ne, O and Si, reporting the mass-fractions of each \citep{2018MNRAS.473.4077P}. 
+In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si and principally compare the logarithmic abundances [X/Fe] and [Fe/H] (defined by 
+\begin{equation}
+     [\mathrm{X}/\mathrm{Y}] = \log_{10}(N_\mathrm X/N_\mathrm Y)_\mathrm{star} - \log_{10}(N_\mathrm X/N_\mathrm Y)_\odot
+\end{equation}
+for number fraction $N_\mathrm X$ of element X), where $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
+This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}.\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}
+
+In this paper, \texttt{Chempy} will be used as the principal GCE model, which, with the modifications described above, allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
 It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{Chempy} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
@@
 
-% \paragraph{Neural network emulator for \texttt{CHEMPY}:}
-% Despite the simplifications made by the GCE model \texttt{CHEMPY}, we still have difficulties sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ due to the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space.
-To alleviate this, we use a \textit{neural network} emulator of the \texttt{CHEMPY} simulations with a total of $n_\mathrm{neuron}=140$ fully-connected neurons in two hidden layers.
-% %
-% % In essence, instead of computing the full model for each input parameter set, we pass the parameters to the network which predicts the output abundances to high accuracy. This has two benefits;
-% % \begin{enumerate}
-% %     \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow posterior sampling.
+\paragraph{Neural network emulator for \texttt{CHEMPY}:}
+Despite the simplifications made by the GCE model \texttt{CHEMPY}, we still have difficulties sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ due to the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space.
+To alleviate this, we follow \cite{Philcox_2019} and implement a \textit{neural network} (NN) emulator of the \texttt{CHEMPY} simulations.
+The NN is a simple feed-forward neural network with 2 hidden layers and 100 neurons in the first and 40 neurons in the second layer.
+
+In essence, instead of computing the full model for each input parameter set, we pass the parameters to the network which predicts the output abundances to high accuracy.
+As already argued in \cite{Philcox_2019} this has two benefits;
+\begin{enumerate}
+    \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow posterior sampling.
 With the neural network, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{Chempy}.
-% %     \item \textbf{Differentiablility:} The neural network has a simple closed-form analytic structure (described in appendix \ref{appen: neural_network}), unlike the complex \texttt{Chempy} model.
+    \item \textbf{Differentiablility:} The neural network has a simple closed-form analytic structure (described in appendix \ref{appen: neural_network}), unlike the complex \texttt{Chempy} model.
 This allows it to be differentiated, so we can sample via advanced methods (cf.\,Sec.\,\ref{sec: sampling}).
-% % \end{enumerate}
-% %
-% Despite the additional complexity introduced by using multiple stellar data-points, our network simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
+\end{enumerate}
+
+Despite the additional complexity introduced by using multiple stellar data-points, our network simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
 The same network can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
-With the above choices, the network predicts abundances with an average error of $0.008_{-0.005}^{+0.007}$\,dex, which is far below typical observational errors and even smaller away from the extremes of parameter space.
-% All our code to reproduce our results is publicly avaialble on \href{https://github.com/bGuenes/sbi_chemical_abundances}{github}.
+With the above choices, the network predicts abundances with an absolute percantage error of $1.6_{-0.9}^{+2.4}$\,\%, \textcolor{red}{is this still the right value?
+using showyourwork, we could do a little evaluation script to have always the right number :-)} \textcolor{blue}{Yep, also correct} which is far below typical observational errors and even smaller away from the extremes of parameter space.
+All our code to reproduce our results is publicly avaialble on \href{https://github.com/bGuenes/sbi_chemical_abundances}{github}.
 
@@
 
-% % Let's see how far we get with our work before the deadline...
+\section{Results}
+\label{sec: Results}
 
-% %We will test our analysis using mock observations drawn firstly from \texttt{Chempy} then from large-scale hydrodynamical simulations to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
+We test our analysis using mock observations drawn firstly from \texttt{Chempy} then from large-scale hydrodynamical simulations taken from the IlustrisTNG suite to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
 The methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
-
-% \section{Results}
-% \label{sec: Results}
 
@@
 
-% We train the NPE with a set of $n_{\rm stars}=10^6$ abundances for $n_{\rm elemets}=8$ with an assumed observational error of $0.05$ dex, generated with the \textit{neural network} emulator, with inputs sampled from the priors shown in Tab.
+We train the NPE with a set of $n_{\rm stars}=10^6$ abundances for $n_{\rm elemets}=8$ with an assumed observational error of $0.05$ dex, generated with the \textit{neural network} emulator, with inputs sampled from the priors shown in Tab.
 \ref{tab:priors}.
@@
 
-% \paragraph{Mock Observational Data}
-% Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
+\paragraph{Mock Observational Data}
+Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
 \ref{tab:priors}, additionally drawing $T_i$ uniformly in the range $[2,12.8]$ Gyr, to minimize overlap with the neural network training birth-time limits when observational uncertainties are included.
-% Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
-% The outcome of this mock data creation is a set of 1000 stars, emulating a real dataset.
+Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
+The outcome of this mock data creation is a set of 1000 stars \textcolor{red}{we vary this number, right?}, emulating a real dataset.
 
-% \paragraph{Inference}
+\paragraph{Inference}
 
@@
 
-% For each set of mock abundances we infer the posterior distribution for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with the NPE by sampling 1000 times and fitting the posterior for each parameter.
+For each set of mock abundances we infer the posterior distribution for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with the NPE by sampling 1000 times and fitting the posterior for each parameter.
 This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
-% Our method takes around $5$ minutes to build a posterior for all six parameters for a dataset size of 1000 stars on a Macbook Air M2 chip, making it around $2400$x faster than current HMC methods \citep[cf.][who needs $40$h for 200 stars with HMC]{Philcox_2019}.
+Our method takes around $5$ minutes to build a posterior for all six parameters for a dataset size of 1000 stars on a Macbook Air M2 chip, making it around $2400$x faster than current HMC methods \citep[cf.][who needs $40$h for 200 stars with HMC]{Philcox_2019}.
 Hence, shorter computing times make it feasible to use orders of magnitudes more observations.
-% %
-% Taking the means of multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$ (Fig.
+%
+Taking the means of multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$ (Fig.
 \ref{fig:N_star_analysis}).
@@
 But in the limit of large numbers of stars (few hundred) the accuracy and precision is superior compared to HMC.
-% %
-% To infer the global galactic parameters $\vec\Lambda$ of the stars as accurate as possible we take the mean of all inferred posteriors from $n_{\rm stars}=1000$ (Fig.
+%
+To infer the global galactic parameters $\vec\Lambda$ of the stars as accurate as possible we take the mean of all inferred posteriors from $n_{\rm stars}=1000$ (Fig.
 \ref{fig:sbi}) resulting in $\alpha_{\rm IMF}=-2.292\pm0.003$ and $\log_{10}(\rm N_{Ia})=-2.883\pm0.005$ which deviates less than $0.4\%$ from the correct value.
@@
 
-% \section{Summary and conclusions}
-% \label{sec: conclusion}
+\section{Summary and conclusions}
+\label{sec: conclusion}
 
-% Here we have explored innovative SBI methods to constrain important galactic parameters from the multi-dimensional observation of chemical abundances of stars.
+Here we have explored innovative SBI methods to constrain important galactic parameters from the multi-dimensional observation of chemical abundances of stars.
 Our results show that SBI provides a super fast method, that beats previous inference algorithms by orders of magnitude, and delivers accurate and precise results.
@@
 
-   \begin{enumerate}
-      \item bla bla
-      \end{enumerate}
-
 \begin{acknowledgements}
       This project was made possible by funding from the Carl Zeiss Stiftung.
-      %Part of this work was supported by the German
-      %\emph{Deut\-sche For\-schungs\-ge\-mein\-schaft, DFG\/} project
-      %number Ts~17/2--1.
 \end{acknowledgements}
@@
 
-% \begin{appendix}
+\begin{appendix}
 
-% \section{Code and Data Availability}
-% \label{sec:appendix_code_and_data}
-% The source code for ...  is made available under an open source license and is publicly hosted on GitHub.\\
-% To facilitate a wider community's usage and contributions, we have ensured that the repository is well-documented.
-The repository includes comprehensive documentation hosted on \emph{ReadTheDocs} that provides an overview of the project, installation instructions, and a guide on how to use the software.
+\section{Code and Data Availability}
+\label{sec:appendix_code_and_data}
+%The source code for this manuscript is made available under an open source license and is publicly hosted on GitHub under this url: \GitHubURL.\\
+% The github repository accompanying this article includes comprehensive documentation %hosted on \emph{ReadTheDocs} that provides an overview of the project, installation instructions, and a guide on how to use the software.
+
+To facilitate a wider community's usage and contributions, we make use of the reproducibility software
+\href{https://github.com/showyourwork/showyourwork}{\showyourwork}
+\citep{Luger2021}, which leverages continuous integration to
+programmatically download the data from
+\href{https://zenodo.org/}{zenodo.org}, create the figures, and
+compile the manuscript. Each figure caption contains two links: one
+to the dataset stored on zenodo used in the corresponding figure,
+and the other to the script used to make the figure (at the commit
+corresponding to the current build of the manuscript). The git
+repository associated to this study is publicly available at
+\url{https://github.com/TobiBu/sbi-chempy}, and the release
+v0.4.1 allows anyone to re-build the entire manuscript including rerunning all analysis.
+The datasets and neural network weights are stored at \url{https://doi.org/10.5281/zenodo.7343715} \textcolor{red}{change to final zenodo dataset!}.
 
@@
 
-%   \end{appendix}
+\end{appendix}
 

### 2024-12-17 src/scripts/train_torch_chempy.py
added train script and adapted the dependencies on zenodo data.

@@ -1,0 +1,189 @@
+import paths
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.stats import norm
+
+from Chempy.parameter import ModelParameters
+
+import torch
+from torch.distributions.normal import Normal
+from torch.distributions.uniform import Uniform
+
+import time as t
+import os
+
+# ---  You want to retrain? ---
+re_train = False
+
+# ----- Load the data ---------------------------------------------------------------------------------------------------------------------------------------------
+# --- Load in training data ---
+path_training = paths.data + '/chempy_data/chempy_TNG_train_data.npz'
+training_data = np.load(path_training, mmap_mode='r')
+
+elements = training_data['elements']
+train_x = training_data['params']
+train_y = training_data['abundances']
+
+
+# ---  Load in the validation data ---
+path_test = paths.data + '/chempy_data/chempy_TNG_val_data.npz'
+val_data = np.load(path_test, mmap_mode='r')
+
+val_x = val_data['params']
+val_y = val_data['abundances']
+
+
+# --- Clean the data ---
+# Chempy sometimes returns zeros or infinite values, which need to be removed
+def clean_data(x, y):
+    # Remove all zeros from the training data
+    index = np.where((y == 0).all(axis=1))[0]
+    x = np.delete(x, index, axis=0)
+    y = np.delete(y, index, axis=0)
+
+    # Remove all infinite values from the training data
+    index = np.where(np.isfinite(y).all(axis=1))[0]
+    x = x[index]
+    y = y[index]
+
+    return x, y
+
+
+train_x, train_y = clean_data(train_x, train_y)
+val_x, val_y     = clean_data(val_x, val_y)
+
+# convert to torch tensors
+train_x = torch.tensor(train_x, dtype=torch.float32)
+train_y = torch.tensor(train_y, dtype=torch.float32)
+val_x = torch.tensor(val_x, dtype=torch.float32)
+val_y = torch.tensor(val_y, dtype=torch.float32)
+
+
+# ----- Define the prior ------------------------------------------------------------------------------------------------------------------------------------------
+a = ModelParameters()
+labels = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
+priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
+
+
[…]

### 2024-12-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,2015MNRAS.447.3880R,2014ApJ...796...75C}.
-In addition, the IMF slope may itself be not a constant but rather a function of metallicity, introducing further complexity \citep[e.g.][]{2019MNRAS.482..118G,2019A&A...626A.124M}.
+For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,Rybizki2015,Chabrier2014}.
+In addition, the IMF slope may itself be not a constant but rather a function of metallicity, introducing further complexity \citep[e.g.][]{2019MNRAS.482..118G,Martin2019}.
 Similarly, the choice of SN\,Ia delay-time-distribution and normalization plays a crucial role in the enrichment of the ISM \citep[e.g.][]{Buck2021} and is heavily debated \citep{2010ApJ...722.1879M,2012MNRAS.426.3282M,2015ApJ...810..137J}.
 
-The goal of this work is to demonstrate how we can use modern machine learning techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
+The goal of this work is to demonstrate how we can use modern machine learning (ML) techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
 Our work focuses on two key parameters; the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF and the rate of SN\,Ia explosions per unit mass.
@@
 
-Our method will make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
-For high-dimensional posterior functions, SBI gives tremendous computational improvements over more traditional methods such as Markov Chain Monte Carlo (MCMC) or even Hamilton Monte Carlo (HMC) methods.
+A methodologically very similar work was previously presented by \cite{Philcox_2019} using Hamilton Monte Carlo sampling for the inference.
+The result of this work was an impressive speed-up over classical Markov Chain Monte Carlo sampling with no loss of accuracy.
+Here we make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
+For high-dimensional posterior functions, SBI gives tremendous computational improvements over more traditional methods such as Markov Chain Monte Carlo (MCMC) or even the more efficient Hamilton Monte Carlo (HMC) methods.
 
-\textcolor{red}{to be adapted...\\
 This paper is structured as follows.
-In Section \ref{sec:methods} we detail our methods.
-Especially, Section \ref{sec:dataset_generation} discusses the methodology used for generating the dataset for our analysis by extracting galaxy data from simulations.
-Section \ref{sec: methods_PCA_on_images} outlines the application of PCA to the image data and the resulting lower-dimensional image space (Section \ref{sec: dimensionality reduction method using pca}).
-Our resulting model is presented in Section \ref{sec: Results} and its accuracy thoroughly tested in Section \ref{sec: model evaluation, reconstruction error}.
-In Section \ref{sec: application} we highlight potential applications of our PCA galaxy model and conclude in Section \ref{sec: conclusion} with a summary and outlook of our results.}
+In Section~\ref{sec:methods} we detail our methods.
+In particular, we discuss the GCE model and the methods behind SBI.
+Our results are presented in Section~\ref{sec: Results} where we test the performance of our approach on mock data and data taken from a hydrodynamical simulation.
+We conclude our study with Section~\ref{sec: conclusion} with a summary and outlook of our results.
 
@@
 
-\paragraph{Simulation-based inference}
-In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
-Once this mapping is trained, we can apply it to real observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
-\ref{fig:flowchart}).
-For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
+In order to establish our new method based on simulation based inference we need two ingredients: A simulator (in our case a Galactic chemical evolution model) that simulates observational data from a set of model parameters (in our case the IMF slope and the Type Ia supernovae normalization) and a flexible way of parametrizing the posterior density conditioned on the observation in order to perform our inference.
+In the next subsections we describe both ingredients in detail.
 
-\paragraph{Galactic chemical evolution models}
-Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}.
+\subsection{Galactic chemical evolution models}
+Our simulator is based on the \texttt{CHEMPY} model \citep{Rybizki_2017}.
 \texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics.
 We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.
-% %
+
+In particular, we are using the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model.
+
+\paragraph{\texttt{CHEMPY} parameters}
 In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
-\vspace{-1em}
-% %\setlist[1]{itemsep=-4pt}
 \begin{enumerate}
      \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$.
-We treat these as star-independent and assume them to be constant across galactic environments and cosmic time.
-\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,Martin2019}.}
+We treat these as star-independent and assume them to be constant across galactic environments and cosmic time\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{Chabrier2014,2016MNRAS.462.2832C,2019MNRAS.482..118G,Martin2019}.}.
      We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}). 
+     %
      \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
@@
 The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
-This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk (which offsets the metallicity for different galactocentric radii).
+This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk \citep[which offsets the metallicity for different galactocentric radii, e.g.][for a simulated example]{Buck2020, Wang2024}.
 Similarly \cite{2019ApJ...874..102W} find that ISM parameter variations are deprojected in the [X/Mg] vs [Mg/H] plane (their Fig.\,17) and that abundance tracks in that space are independent of the stellar sample's spatial position within the Galaxy (their Fig.\,3).
@@
 
+\paragraph{Nucleosynthetic yield tables}
 We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
+To test our method, we aim further at inferring parameters from a sample of stars taken from a hydrodynamical simulation of a MW type galaxy which we take from the Ilustris TNG project \citep{Pillepich2018}.
 To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
@@
 \texttt{Chempy} uses only net yields, such that they provide only newly synthesized material, with the remainder coming from the initial SSP composition.
-These tables may not well-represent true stellar chemistry, and the effects of this are examined in Sec.\,\ref{subsec: mocks_wrong_yield} by performing inference using an alternative set of yields.
+These tables may not well-represent true stellar chemistry, and the effects of this missmatch are examined in Sec.\,\ref{subsec: mocks_wrong_yield} by performing inference using an alternative set of yields that does not match the yield set of the training data.
 For the analysis of observational data, we would want to use the most up-to-date yields, such as \citet{2016ApJ...825...26K} AGB yields, and carefully chose elements which are known to be well reproduced by our current models (e.g.
 shown by \citet{2019ApJ...874..102W,2019arXiv190806113G}), though this is not appropriate in our context.
-To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{1998A&A...334..505P} and do not allow for any `hypernovae' (in contrary to \citetalias{2018ApJ...861...40P}).
+To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{1998A&A...334..505P} and do not allow for any `hypernovae' \citep[in contrary to][]{2018ApJ...861...40P}).
 
@@
 
-%TNG only tracks nine elements in their analysis: C, Fe, H, He, Mg, N, Ne, O and Si, reporting the mass-fractions of each \citep{2018MNRAS.473.4077P}. 
-In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si and principally compare the logarithmic abundances [X/Fe] and [Fe/H] (defined by 
+\paragraph{Chemical elements}
+In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si since these are the only elements traced by TNG.
+We principally compare the logarithmic abundances [X/Fe] and [Fe/H] defined by
 \begin{equation}
@@
 \end{equation}
-for number fraction $N_\mathrm X$ of element X), where $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
-This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}.\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}
+for number fraction $N_\mathrm X$ of element X.
+Here $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
+This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}.
 
-In this paper, \texttt{Chempy} will be used as the principal GCE model, which, with the modifications described above, allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
+With these modifications, \texttt{Chempy} allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
 It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{Chempy} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
-This is a necessary test to prepare for an inference on real data.
+However,  such a test is necessary to prepare for an inference on real data.
 
-\paragraph{Neural network emulator for \texttt{CHEMPY}:}
-Despite the simplifications made by the GCE model \texttt{CHEMPY}, we still have difficulties sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ due to the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space.
-To alleviate this, we follow \cite{Philcox_2019} and implement a \textit{neural network} (NN) emulator of the \texttt{CHEMPY} simulations.
-The NN is a simple feed-forward neural network with 2 hidden layers and 100 neurons in the first and 40 neurons in the second layer.
+\subsection{Neural network emulator for \texttt{CHEMPY}:}
+Despite the simplifications made by the GCE model \texttt{CHEMPY}, the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space incurs some difficulties when sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$.
+To alleviate this, we follow \cite{Philcox_2019} and implement a \textit{neural network} (NN) emulator of the \texttt{CHEMPY} simulator.
+We design the NN as a simple feed-forward neural network with 2 hidden layers and 100 neurons in the first and 40 neurons in the second layer.
 
-In essence, instead of computing the full model for each input parameter set, we pass the parameters to the network which predicts the output abundances to high accuracy.
+In essence, instead of computing the full model for each input parameter set, we pass the parameters to the NN which predicts the output abundances to high accuracy.
 As already argued in \cite{Philcox_2019} this has two benefits;
 \begin{enumerate}
-    \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow posterior sampling.
-With the neural network, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{Chempy}.
-    \item \textbf{Differentiablility:} The neural network has a simple closed-form analytic structure (described in appendix \ref{appen: neural_network}), unlike the complex \texttt{Chempy} model.
-This allows it to be differentiated, so we can sample via advanced methods (cf.\,Sec.\,\ref{sec: sampling}).
+    \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow generation of training data for SBI.
+With the NN emulator, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{Chempy}.
+    \item \textbf{Differentiability:} The NN is written in pytorch which allows for automatic differentiation.
+Additionally the NN has a simple closed-form analytic structure \citep[described in the appendix of][]{Philcox_2019}, unlike the complex \texttt{Chempy} model.
+This allows it to be differentiated, so one can use it to sample via advanced methods such HMC as done in \citep{Philcox_2019}.
 \end{enumerate}
 
-Despite the additional complexity introduced by using multiple stellar data-points, our network simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
-The same network can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
-With the above choices, the network predicts abundances with an absolute percantage error of $1.6_{-0.9}^{+2.4}$\,\%, \textcolor{red}{is this still the right value?
-using showyourwork, we could do a little evaluation script to have always the right number :-)} \textcolor{blue}{Yep, also correct} which is far below typical observational errors and even smaller away from the extremes of parameter space.
-All our code to reproduce our results is publicly avaialble on \href{https://github.com/bGuenes/sbi_chemical_abundances}{github}.
+Despite the additional complexity introduced by using multiple stellar data-points, our NN simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
+The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
+With the above network parameter choices, the NN predicts abundances with an absolute percantage error of $1.6_{-0.9}^{+2.4}$\,\%, which is far below typical observational errors and even smaller away from the extremes of parameter space.
+Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy}{https://github.com/TobiBu/sbi-chempy}.
+\textcolor{red}{adapt to point to final folder on the repo.
+and maybe add zenodo link for the nueral network weights...}
+
+\begin{figure}[]
+     \centering
+     \includegraphics[width=\columnwidth]{figures/ape_NN.png}
+     \vspace{-.5cm}
+     \caption{Absolute percentage error (APE) of the NN emulator for the \texttt{CHEMPY} simulator.}
+     \label{fig:ape_NN}
+     \script{train_torch_chempy.py}
+\end{figure}
 
 % %The larger volume of data should be able to give tighter statistical constraints on those parameters that are held fixed across the galaxy, but complexity is added since we must allow each star to carry its own set of local ISM parameters.
+
+\paragraph{Simulation-based inference}
+In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
+Once this mapping is trained, we can apply it to real observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
+\ref{fig:flowchart}).
+For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
 
@@
 %     % \textcolor{blue}{For training the sbi net i used compgpu11 and it took around 3h. The inference of the stars i did on my laptop (macbook air with a M2 chip) in 5 minutes}
+
+\subsection{Alternative yield set}
+\label{subsec: mocks_wrong_yield}
 

### 2024-12-18 src/scripts/train_torch_chempy.py
save APE calculation to txt file.

@@ -16,15 +16,6 @@ import os
 re_train = False
 
 # ----- Load the data ---------------------------------------------------------------------------------------------------------------------------------------------
-# --- Load in training data ---
-path_training = paths.data / 'chempy_data/chempy_TNG_train_data.npz'
-training_data = np.load(path_training, mmap_mode='r')
-
-elements = training_data['elements']
-train_x = training_data['params']
-train_y = training_data['abundances']
-
-
 # ---  Load in the validation data ---
 path_test = paths.data / 'chempy_data/chempy_TNG_val_data.npz'
 val_data = np.load(path_test, mmap_mode='r')
@@ -48,13 +39,9 @@ def clean_data(x, y):
 
     return x, y
 
-
-train_x, train_y = clean_data(train_x, train_y)
 val_x, val_y     = clean_data(val_x, val_y)
 
 # convert to torch tensors
-train_x = torch.tensor(train_x, dtype=torch.float32)
-train_y = torch.tensor(train_y, dtype=torch.float32)
 val_x = torch.tensor(val_x, dtype=torch.float32)
 val_y = torch.tensor(val_y, dtype=torch.float32)
 
@@ -88,14 +75,27 @@ model = Model_Torch()
 optimizer = torch.optim.Adam(model.parameters(), lr=0.001)
 loss_fn = torch.nn.MSELoss()
 
-# shuffle the data
-index = np.arange(train_x.shape[0])
-np.random.shuffle(index)
-train_x = train_x[index]
-train_y = train_y[index]
-
 # --- Train the neural network ---
 if re_train:
+    # --- Load in training data ---
+    path_training = paths.data / 'chempy_data/chempy_TNG_train_data.npz'
+    training_data = np.load(path_training, mmap_mode='r')
+
+    elements = training_data['elements']
+    train_x = training_data['params']
+    train_y = training_data['abundances']
+
+    train_x, train_y = clean_data(train_x, train_y)
+
+    train_x = torch.tensor(train_x, dtype=torch.float32)
+    train_y = torch.tensor(train_y, dtype=torch.float32)
+
+    # shuffle the data
+    index = np.arange(train_x.shape[0])
+    np.random.shuffle(index)
+    train_x = train_x[index]
+    train_y = train_y[index]
+
     print("Training the model")
     epochs = 20
     batch_size = 64
[…]

### 2024-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
-With the above network parameter choices, the NN predicts abundances with an absolute percantage error of $1.6_{-0.9}^{+2.4}$\,\%, which is far below typical observational errors and even smaller away from the extremes of parameter space.
+With the above network parameter choices, the NN predicts abundances with an absolute percantage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space.
 Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy}{https://github.com/TobiBu/sbi-chempy}.

### 2024-12-18 src/scripts/train_torch_chempy.py
[showyourwork] overleaf sync

@@ -57,9 +57,9 @@ priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_opti
 class Model_Torch(torch.nn.Module):
     def __init__(self):
         super(Model_Torch, self).__init__()
-        self.l1 = torch.nn.Linear(train_x.shape[1], 100)
+        self.l1 = torch.nn.Linear(val_x.shape[1], 100)
         self.l2 = torch.nn.Linear(100, 40)
-        self.l3 = torch.nn.Linear(40, train_y.shape[1])
+        self.l3 = torch.nn.Linear(40, val_y.shape[1])
 
     def forward(self, x):
         x = torch.tanh(self.l1(x))

### 2025-01-09 src/scripts/evaluate_sbi.py
added plotting scripts for SBC plots and APE plots.

@@ -1,0 +1,89 @@
+import paths
+from sbi.analysis.plot import sbc_rank_plot
+from sbi.diagnostics import plot_tarp, check_tarp, run_sbc, run_tarp
+import pickle
+
+from Chempy.parameter import ModelParameters
+
+from plot_functions import *
+
+# ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
+
+name = "NPE_C"
+
+# --- Absolute percentage error plot ---
+
+# --- Define the prior ---
+a = ModelParameters()
+labels_out = a.elements_to_trace
+labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
+priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
+
+combined_priors = utils.MultipleIndependent(
+    [Normal(p[0]*torch.ones(1), p[1]*torch.ones(1)) for p in priors] +
+    [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
+    validate_args=False)
+
+# ----- Save the posterior -------------------------------------------------------------------------------------------------------------------------------------------
+with open(paths.data / f'posterior_{name}.pickle', 'rb') as f:
+    posterior = pickle.load(f)
+
+# --- Simulation based calibration plot ---
+def simulator(params):
+    y = model(params)
+    y = y.detach().numpy()
+
+    # Remove H from data, because it is just used for normalization (output with index 2)
+    y = np.delete(y, 2,1)
+
+    return y
+
+num_sbc_samples = 200  # choose a number of sbc runs, should be ~100s
+# generate ground truth parameters and corresponding simulated observations for SBC.
+thetas = combined_priors.sample((num_sbc_samples,))
+xs = simulator(thetas)
+
+# run SBC: for each inference we draw 1000 posterior samples.
+num_posterior_samples = 1_000
+num_workers = 1
+ranks, dap_samples = run_sbc(
+    thetas, xs, posterior, num_posterior_samples=num_posterior_samples, num_workers=num_workers
+)
+
+# --- SBC rank plot ---
+f, ax = sbc_rank_plot(
+    ranks=ranks,
+    num_posterior_samples=num_posterior_samples,
+    parameter_labels=labels_in,
+    plot_type="hist",
+    num_cols=3,
+    figsize=(15,10),
[…]

### 2025-01-09 src/scripts/plot_functions.py

@@ -1,0 +1,369 @@
+import paths
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.ticker import NullFormatter
+from scipy.stats import norm
+from scipy.stats import multivariate_normal
+
+##########################################################################################################
+# --- Calculate mean and error ---
+def mean_error(x, N_stars):
+    mean = np.mean(x)
+    std = np.std(x)
+    err = std/np.sqrt(N_stars)
+
+    return mean, err, std
+
+##########################################################################################################
+# --- Plot 1D Histogram ---
+def plot_1d_hist(x1, x2, x_true, names, simulations):
+    fig, ax = plt.subplots(1,2, figsize=(12,6))
+    plot_width = 0.3
+    
+    def plot_subplot(i, x, true, name):
+        N_stars = int(len(x)/simulations)
+        mean, err, std = mean_error(x, N_stars)
+
+        # Set the x-axis limits
+        x_min = true - plot_width*true
+        x_max = true + plot_width*true
+
+        # Fit a Gaussian to the data
+        lin = np.linspace(x_min, x_max, 1000)
+        p = norm.pdf(lin, mean, std)
+
+        # Plot the data
+        ax[i].hist(x, bins=500, density=True, alpha=0.6, color='g')
+        ax[i].plot(lin, p, 'k', linewidth=2)
+        ax[i].axvline(x=true, color='r', linestyle='dashed', linewidth=2)
+        
+        ax[i].set_title(fr"{name}: {true:.3f}"
+                        "\n"
+                        fr"Fit: {mean:.3f} $\pm$ {err:.3f}")
+        ax[i].set_xlim(x_min, x_max)
+
+    for i in [0,1]:
+        if i == 0:
+            x = x1
+        else:
+            x = x2
+
+        plot_subplot(i, x, x_true[i], names[i])
+    
+    fig.suptitle(f'{int(len(x1)/simulations)} Stars', fontsize=30)
+    plt.tight_layout()
+    plt.show()
+
+##########################################################################################################
+# --- 2D Histogram preparation ---
[…]

### 2025-01-09 src/scripts/torch_chempy_model.py

@@ -1,0 +1,18 @@
+import torch
+
+# ----- Define the model ------------------------------------------------------------------------------------------------------------------------------------------
+
+class Model_Torch(torch.nn.Module):
+    def __init__(self):
+        super(Model_Torch, self).__init__()
+        self.l1 = torch.nn.Linear(val_x.shape[1], 100)
+        self.l2 = torch.nn.Linear(100, 40)
+        self.l3 = torch.nn.Linear(40, val_y.shape[1])
+
+    def forward(self, x):
+        x = torch.tanh(self.l1(x))
+        x = torch.tanh(self.l2(x))
+        x = self.l3(x)
+        return x
+
+model = Model_Torch()

### 2025-01-09 src/scripts/train_sbi.py

@@ -1,0 +1,154 @@
+import paths
+import numpy as np
+
+from scipy.stats import norm
+
+from Chempy.parameter import ModelParameters
+
+import sbi.utils as utils
+from sbi.utils.user_input_checks import check_sbi_inputs, process_prior, process_simulator
+from sbi.inference import simulate_for_sbi, NPE_C
+from sbi.neural_nets import posterior_nn
+
+import torch
+from torch.distributions.normal import Normal
+from torch.distributions.uniform import Uniform
+
+import time as t
+import pickle
+from tqdm import tqdm
+
+from torch_chempy_model import model
+
+# ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
+
+name = "NPE_C"
+
+# ----- Load the model -------------------------------------------------------------------------------------------------------------------------------------------
+# --- Define the prior ---
+a = ModelParameters()
+labels_out = a.elements_to_trace
+labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
+priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
+
+combined_priors = utils.MultipleIndependent(
+    [Normal(p[0]*torch.ones(1), p[1]*torch.ones(1)) for p in priors] +
+    [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
+    validate_args=False)
+
+
+# --- Load the weights ---
+model.load_state_dict(torch.load(paths.data / 'pytorch_state_dict.pt'))
+model.eval()
+
+
+# ----- Set up the simulator -------------------------------------------------------------------------------------------------------------------------------------------
+def simulator(params):
+    y = model(params)
+    y = y.detach().numpy()
+
+    # Remove H from data, because it is just used for normalization (output with index 2)
+    y = np.delete(y, 2)
+
+    return y
+
+
+prior, num_parameters, prior_returns_numpy = process_prior(combined_priors)
+simulator = process_simulator(simulator, prior, prior_returns_numpy)
+check_sbi_inputs(simulator, prior)
+
+
+# ----- Train the SBI -------------------------------------------------------------------------------------------------------------------------------------------
[…]

### 2025-01-09 src/scripts/train_torch_chempy.py

@@ -10,7 +10,7 @@ from torch.distributions.normal import Normal
 from torch.distributions.uniform import Uniform
 
 import time as t
-import os
+from torch_chempy_model import model
 
 # ---  You want to retrain? ---
 re_train = False
@@ -51,26 +51,8 @@ a = ModelParameters()
 labels = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
 priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
 
-
-# ----- Define the model ------------------------------------------------------------------------------------------------------------------------------------------
-
-class Model_Torch(torch.nn.Module):
-    def __init__(self):
-        super(Model_Torch, self).__init__()
-        self.l1 = torch.nn.Linear(val_x.shape[1], 100)
-        self.l2 = torch.nn.Linear(100, 40)
-        self.l3 = torch.nn.Linear(40, val_y.shape[1])
-
-    def forward(self, x):
-        x = torch.tanh(self.l1(x))
-        x = torch.tanh(self.l2(x))
-        x = self.l3(x)
-        return x
-
-model = Model_Torch()
-
-
 # ----- Train the model -------------------------------------------------------------------------------------------------------------------------------------------
+
 
 optimizer = torch.optim.Adam(model.parameters(), lr=0.001)
 loss_fn = torch.nn.MSELoss()
@@ -189,4 +171,4 @@ else:
     plt.clf()
 
     with open(paths.output / 'ape_NN.txt', 'w') as f:
-        f.write(f'${p2:.1f}^{{+{p3-p2:.1f}}}_{{-{p2-p1:.1f}}}\,\%$')
+        f.write(f'${p2:.1f}^{{+{p3-p2:.1f}}}_{{-{p2-p1:.1f}}}\,\%$%')

### 2025-01-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 With the above network parameter choices, the NN predicts abundances with an absolute percantage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space.
-Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy}{https://github.com/TobiBu/sbi-chempy}.
-\textcolor{red}{adapt to point to final folder on the repo.
-and maybe add zenodo link for the nueral network weights...}
+Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.
 
@@
      \vspace{-.5cm}
-     \caption{Absolute percentage error (APE) of the NN emulator for the \texttt{CHEMPY} simulator.}
+     \caption{Absolute percentage error of the NN emulator for the \texttt{CHEMPY} simulator.
+The orange histogram shows the full distribution of percentage errors with the vertical dashed line indicating the median and the vertical dotted lines indicating the first and third quartile.
+The box plot on the top of the plot extends from the first quartile to the third quartile of the data, with a line at the median.
+The whiskers extend from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box.
+The NN predicts abundances with an absolute percentage error far below typical observational errors.}
      \label{fig:ape_NN}
@@
 \paragraph{Simulation-based inference}
-In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a  Bayesian inference framework --  works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical enrichment model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
-Once this mapping is trained, we can apply it to real observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
+In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a Bayesian inference framework -- works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical evolution model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
+Once this mapping is trained, we can apply it to observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
 \ref{fig:flowchart}).
-For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 5 transformer blocks with 50 hidden features each.
+
+For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 10 hidden features and 1 transformation layer for the normalizing flow.
+
+We train our NPE with $10^5$ data points simulated with the NN emulator described in the previous section until convergence.
+Training takes $\lesssim10$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPUs.
+We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator.
+In order to mimick observational noise, we add 5\% observational uncertainties to the abundances simulated with \texttt{CHEMPY} before feeding them to our NPE.
+
+\begin{figure}[]
+     \centering
+     \includegraphics[width=\columnwidth]{figures/posterior_ape.png}
+     \vspace{-.5cm}
+     \caption{Absolute percentage error of the neural posterior density estimate for a single star.
+Different coloured histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines.
+The box plots show again the first and third quartiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box.
+The global parameters of main interest for this work are shown by the light blue and red histogram.}
+     \label{fig:ape_NN}
+     \script{train_sbi.py}
+\end{figure}
+
+For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate with 1000 data points (see Fig.~\ref{fig:posterior_samples} in the appendix).
+We then compare the posterior mean to the ground truth value.
+For a single star observation, our NPE has an APE of \variable{output/posterior_APE.txt} when looking at all 6 parameters of interest.
+If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/global_posterior_APE_NPE_C.txt} as we show in Fig.~\ref{fig:posterior_APE}.
+Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high.
+However, for the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ we can boost the accuracy by combining the inference for many stars of the same galaxy.
 

### 2025-01-09 src/scripts/torch_chempy_model.py
restructure code

@@ -1,18 +1,0 @@
-import torch
-
-# ----- Define the model ------------------------------------------------------------------------------------------------------------------------------------------
-
-class Model_Torch(torch.nn.Module):
-    def __init__(self):
-        super(Model_Torch, self).__init__()
-        self.l1 = torch.nn.Linear(val_x.shape[1], 100)
-        self.l2 = torch.nn.Linear(100, 40)
-        self.l3 = torch.nn.Linear(40, val_y.shape[1])
-
-    def forward(self, x):
-        x = torch.tanh(self.l1(x))
-        x = torch.tanh(self.l2(x))
-        x = self.l3(x)
-        return x
-
-model = Model_Torch()

### 2025-01-09 src/scripts/train_sbi.py

@@ -36,6 +36,21 @@ combined_priors = utils.MultipleIndependent(
     [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
     validate_args=False)
 
+# --- Set up the model ---
+class Model_Torch(torch.nn.Module):
+    def __init__(self):
+        super(Model_Torch, self).__init__()
+        self.l1 = torch.nn.Linear(len(labels_in), 100)
+        self.l2 = torch.nn.Linear(100, 40)
+        self.l3 = torch.nn.Linear(40, len(labels_out))
+
+    def forward(self, x):
+        x = torch.tanh(self.l1(x))
+        x = torch.tanh(self.l2(x))
+        x = self.l3(x)
+        return x
+
+model = Model_Torch()
 
 # --- Load the weights ---
 model.load_state_dict(torch.load(paths.data / 'pytorch_state_dict.pt'))

### 2025-01-09 src/scripts/train_torch_chempy.py

@@ -10,7 +10,6 @@ from torch.distributions.normal import Normal
 from torch.distributions.uniform import Uniform
 
 import time as t
-from torch_chempy_model import model
 
 # ---  You want to retrain? ---
 re_train = False
@@ -51,8 +50,26 @@ a = ModelParameters()
 labels = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
 priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
 
-# ----- Train the model -------------------------------------------------------------------------------------------------------------------------------------------
 
+# ----- Define the model ------------------------------------------------------------------------------------------------------------------------------------------
+
+class Model_Torch(torch.nn.Module):
+    def __init__(self):
+        super(Model_Torch, self).__init__()
+        self.l1 = torch.nn.Linear(val_x.shape[1], 100)
+        self.l2 = torch.nn.Linear(100, 40)
+        self.l3 = torch.nn.Linear(40, val_y.shape[1])
+
+    def forward(self, x):
+        x = torch.tanh(self.l1(x))
+        x = torch.tanh(self.l2(x))
+        x = self.l3(x)
+        return x
+
+model = Model_Torch()
+
+
+# ----- Train the model -------------------------------------------------------------------------------------------------------------------------------------------
 
 optimizer = torch.optim.Adam(model.parameters(), lr=0.001)
 loss_fn = torch.nn.MSELoss()

### 2025-01-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
+\textcolor{red}{here goes the assumption that stars are iid since each stars is produced by a single call to chempy.
+However, real physics might correlate stars abundacnes by star formation histories and cumulative enrichment...}
 The outcome of this mock data creation is a set of 1000 stars \textcolor{red}{we vary this number, right?}, emulating a real dataset.

### 2025-01-09 src/scripts/train_sbi.py
[showyourwork] overleaf sync

@@ -18,8 +18,6 @@ import time as t
 import pickle
 from tqdm import tqdm
 
-from torch_chempy_model import model
-
 # ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
 
 name = "NPE_C"

### 2025-01-09 src/scripts/evaluate_sbi.py
[showyourwork] overleaf sync

@@ -1,7 +1,12 @@
 import paths
+import pickle
+import torch 
+from torch.distributions.normal import Normal
+from torch.distributions.uniform import Uniform
+
+import sbi.utils as utils
 from sbi.analysis.plot import sbc_rank_plot
 from sbi.diagnostics import plot_tarp, check_tarp, run_sbc, run_tarp
-import pickle
 
 from Chempy.parameter import ModelParameters
 

### 2025-01-09 src/scripts/train_sbi.py

@@ -11,8 +11,6 @@ from sbi.inference import simulate_for_sbi, NPE_C
 from sbi.neural_nets import posterior_nn
 
 import torch
-from torch.distributions.normal import Normal
-from torch.distributions.uniform import Uniform
 
 import time as t
 import pickle

### 2025-01-09 src/scripts/train_torch_chempy.py

@@ -136,7 +136,7 @@ if re_train:
     plt.title('Training and Validation Loss', fontsize=20)
     plt.legend()
     plt.tight_layout()
-    plt.savefig(paths.figures / "loss_NN_simulator.png")
+    plt.savefig(paths.figures / "loss_NN_simulator.pdf")
     plt.clf()
 
 else:
@@ -171,7 +171,7 @@ else:
     plt.xlim(0, 30)
     fig.tight_layout()
 
-    plt.savefig(paths.figures / "ape_NN.png")
+    plt.savefig(paths.figures / "ape_NN.pdf")
     plt.clf()
 
     with open(paths.output / 'ape_NN.txt', 'w') as f:

### 2025-01-09 src/scripts/train_sbi.py
[showyourwork] overleaf sync

@@ -11,6 +11,8 @@ from sbi.inference import simulate_for_sbi, NPE_C
 from sbi.neural_nets import posterior_nn
 
 import torch
+from torch.distributions.normal import Normal
+from torch.distributions.uniform import Uniform
 
 import time as t
 import pickle

### 2025-01-09 src/scripts/evaluate_posterior.py
[showyourwork] overleaf sync

@@ -54,6 +54,8 @@ val_x = torch.tensor(val_x, dtype=torch.float32)
 abundances =  torch.cat([val_x[:,:2], val_x[:,3:]], dim=1)
 
 # add noise to data to simulate observational errors
+pc_ab = 5 # percentage error in abundance
+
 x_err = np.ones_like(abundances)*float(pc_ab)/100.
 abundances = norm.rvs(loc=abundances,scale=x_err)
 abundances = torch.tensor(abundances).float()

### 2025-01-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
      \centering
-     \includegraphics[width=\columnwidth]{figures/posterior_ape.png}
+     \includegraphics[width=\columnwidth]{figures/ape_posterior_NPE_C.pdf}
      \vspace{-.5cm}
@@
      \label{fig:ape_NN}
-     \script{train_sbi.py}
+     \script{evaluate_posterior.py}
 \end{figure}
@@
 For a single star observation, our NPE has an APE of \variable{output/posterior_APE.txt} when looking at all 6 parameters of interest.
-If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/global_posterior_APE_NPE_C.txt} as we show in Fig.~\ref{fig:posterior_APE}.
+If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/global_posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}.
 Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high.

### 2025-01-09 src/scripts/plot_functions.py

@@ -333,7 +333,7 @@ def ape_plot(ape, labels_in, save_path):
     print(f"Total : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
     print("")
 
-    with open(paths.output / f'global_posterior_APE_{name}.txt', 'w') as f:
+    with open(paths.output / f'global_posterior_APE.txt', 'w') as f:
         f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")
 
     for i in range(ape.shape[1]):

### 2025-01-09 src/scripts/evaluate_posterior.py
[showyourwork] overleaf sync

@@ -69,5 +69,19 @@ for index in tqdm(range(len(abundances))):
 ape = torch.abs((val_theta - theta_hat) / val_theta) * 100
 torch.save(ape, paths.data / f'ape_posterior_{name}.pt')
 
-save_path = paths.figures / f'ape_posterior_{name}.pdf'
-ape_plot(ape, labels_in, save_path)
+print("\nAPE of the Posterior:")
+print("Median + upper quantile - lower quantile")
+l_quantile, median, u_quantile = np.percentile(ape, [25, 50, 75])
+print(f"Total : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
+print("")
+
+with open(paths.output / f'global_posterior_APE.txt', 'w') as f:
+    f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")
+
+for i in range(ape.shape[1]):
+    l_quantile, median, u_quantile = np.percentile(ape[:,i], [25, 50, 75])
+    median = np.percentile(ape[:,i], 50)
+    print(labels_in[i] + f" : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
+    if i in [0,1]:
+        with open(paths.output / f'posterior_APE.txt', 'a') as f:
+            f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")

### 2025-01-09 src/scripts/plot_functions.py

@@ -326,27 +326,10 @@ def n_stars_plot_comp(x1, x2, x_true, dat, no_stars= np.array([1, 10, 100, 500, 
 def ape_plot(ape, labels_in, save_path):
     fig, (ax_box, ax_hist) = plt.subplots(2, sharex=True, gridspec_kw={"height_ratios": (.20, .80)})
     colors = ["tomato", "skyblue", "olive", "gold", "teal", "orchid"]
-    
-    print("\nAPE of the Posterior:")
-    print("Median + upper quantile - lower quantile")
-    l_quantile, median, u_quantile = np.percentile(ape, [25, 50, 75])
-    print(f"Total : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
-    print("")
-
-    with open(paths.output / f'global_posterior_APE.txt', 'w') as f:
-        f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")
 
     for i in range(ape.shape[1]):
-        l_quantile, median, u_quantile = np.percentile(ape[:,i], [25, 50, 75])
         ax_hist.hist(ape[:,i], bins=25, density=True, range=(0, 100), label=labels_in[i], color=colors[i], alpha=0.5)
-        median = np.percentile(ape[:,i], 50)
         ax_hist.axvline(median, color=colors[i], linestyle='--')
-        print(labels_in[i] + f" : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
-        if i in [0,1]:
-            with open(paths.output / f'posterior_APE.txt', 'a') as f:
-                f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")
-
-    print()
             
     ax_hist.set_xlabel('Error (%)', fontsize=15)
     ax_hist.set_ylabel('Density', fontsize=15)

### 2025-01-09 src/scripts/plot_posterior_APE.py

@@ -1,0 +1,19 @@
+import paths
+import torch
+from Chempy.parameter import ModelParameters
+
+from plot_functions import ape_plot
+
+# ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
+name = "NPE_C"
+
+# --- Define the prior ---
+a = ModelParameters()
+labels_out = a.elements_to_trace
+labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
+
+
+ape = torch.load(paths.data / f'ape_posterior_{name}.pt')
+
+save_path = paths.figures / f'ape_posterior_{name}.pdf'
+ape_plot(ape, labels_in, save_path)

### 2025-01-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
-With the above network parameter choices, the NN predicts abundances with an absolute percantage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space.
+With the above network parameter choices, the NN predicts abundances with an absolute percantage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
 Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.
@@
 
+\paragraph{Bayesian model}
+\texttt{CHEMPY} effectively is a Bayesian model for stellar abundances given a set of parameters $\{\Vec{\Lambda},\Vec{\Theta},T\}$ and \citet{Philcox_2019} extended the \texttt{CHEMPY} framework to be able to model multiple stellar data-points.
+Consider a given star with index $i$ that is born in some region of the ISM.
+This star will carry its own set of parameters $\{\Vec{\Lambda},\Vec{\Theta}_i,T_i\}$, where $\Vec{\Lambda}$ are star independent and hence taken to be global parameters while the ISM parameters $\Vec{\Theta}_i$ and the birth-time $T_i$ are star specific.
+Using \texttt{CHEMPY} (or the trained neural network emulator) we can easily model the set of $n_\mathrm{el}$ chemical abundances $\{X_i^j\}$ for the $i$-th star as:
+\begin{equation}\label{eq:chempy_function}
+\{X_i^j\} = \texttt{CHEMPY}(\Vec{\Lambda},\Vec{\Theta}_i,T_i),
+\end{equation}
+where $j$ indexes the chemical element. 
+These model abundances can then be compared against observations, with measured abundances $d_i^j$ and corresponding Gaussian errors $\sigma_{i,\mathrm{obs}}^j$ jointly denoted as $D_i=\{d_i^j,\sigma_{i,\mathrm{obs}}^j\}$.
+
+As stated in \citet{Philcox_2019} the full posterior for this case is given by 
+\begin{eqnarray}\label{eq:posterior}
+    \mathbb{P}(\vec\Lambda,\{\vec\Theta_i\},\{T_i\}|\{D_i\}) &\propto&  \left[\prod_{i=1}^{n_\mathrm{star}}p_{\vec\Theta}(\vec\Theta_i)p_{T_i}(T_i)\right]
+    \times p_{\vec\Lambda}(\vec\Lambda)\\
+    \nonumber
+    &\times& \mathcal{L}(\{D_i\}|\vec\Lambda,\{\vec\Theta_i\},\{T_i\})
+\end{eqnarray}
+where $p_V(V_i)$ is the prior on variable $V_i$ (belonging to the set $V$).
+
+In order to determine the optimal values of the global galactic parameters ($\Vec{\Lambda}$) one has to sample the posterior of Eq.~\ref{eq:posterior}.
+This is in practices a fairly costly computation and even in the case of advanced techniques such as Hamilton Monte Carlo sampling the posterior can only be evaluated for a small set of stars ($\lesssim200$) using fairly long compute times \citep[$\sim40$ hours][]{Philcox_2019}.
+%
+However, recent advances in implicit-likelihood inference or simulation-based inference \citep{Cranmer2020} offer another very efficient approach to approximate the posterior (see next paragraph for more details).
+These methods train a neural conditional density estimator to represent the conditional posterior, $\mathbb{P}(\vec\Lambda|\{D_i\})$, which can be very efficiently evaluated given observational data $\{D_i\}$.
+
+In particular, if we marginalize over the star specific parameters and solely focus on the global parameters $\Vec{\Lambda}$ we can make the assumption that individual observations of stars are identically and independently distributed (i.i.d.) and factorize the joint posterior from above to simply express it as:  %
+\begin{eqnarray}\label{eq:posterior}
+\mathbb{P}(\vec\Lambda|\{D_i\}) &\propto& 
+\mathbb{P}(\vec\Lambda)\mathbb{P}(\{D_i\}|\Lambda)\quad\quad\quad\quad\quad\quad\,\, \text{(Bayes rule)}
+\\\nonumber
+&=&
+\mathbb{P}(\vec\Lambda)\mathbb{P}
+(D_1,...,D_{n_\mathrm{star}}|\vec\Lambda)
+\\\nonumber
+&=& \mathbb{P}(\vec\Lambda)\prod_{j=1}^{n_\mathrm{star}}\mathbb{P}(D_j|\vec\Lambda) 
+\\\nonumber
+&\propto& \mathbb{P}(\vec\Lambda) \prod_{j=1}^{n_\mathrm{star}}\frac{\mathbb{P}(\vec\Lambda|D_j)}{\mathbb{P}(\vec\Lambda)} \quad\quad\quad\quad\quad\text{(Bayes rule)}
+\\\nonumber
+&=& \mathbb{P}(\vec\Lambda)^{1-n_\mathrm{star}}\ \prod_{j=1}^{n_\mathrm{star}}\mathbb{P}(\vec\Lambda|D_j) 
+\end{eqnarray}
+
+This factorization of the joint posterior into single star posteriors $\mathbb{P}(\vec\Lambda|D_j)$ has the advantage that we do not need to specify the exact number of observational data points before.
+We can simply train our neural density estimator on single star observations and then at evaluation time, we simply combine as many posterior estimates as we have observational data.
+
 \paragraph{Simulation-based inference}
-In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a Bayesian inference framework -- works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\theta$ (in our case a galactic chemical evolution model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\theta|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\theta$ that reproduce the simulated observations $\Vec{X}$.
-Once this mapping is trained, we can apply it to observations of stellar abundances $\Vec{X_R}$ to infer $p$($\theta|\Vec{X_R}$, $\mathcal{M}$)(see Fig.
-\ref{fig:flowchart}).
+In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a Bayesian inference framework -- works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\Vec{\theta}$ (in our case a galactic chemical evolution model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\Vec{\theta}|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\Vec{\theta}$ that reproduce the simulated observations $\Vec{X}$.
+Once this mapping is trained, we can apply it to observations of stellar abundances $\Vec{X_R}$ to infer $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$) (see Fig.
+\ref{fig:flowchart} for a schematic visual representation of our method).
+Note, we do not need to know anything about $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$), we solely need to be able to sample from it.
 
-For this we use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 10 hidden features and 1 transformation layer for the normalizing flow.
+We use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 10 hidden features and 1 transformation layer for the normalizing flow.
 
@@
 The global parameters of main interest for this work are shown by the light blue and red histogram.}
-     \label{fig:ape_NN}
-     \script{evaluate_posterior.py}
+     \label{fig:posterior_APE}
+     \script{plot_posterior_APE.py}
 \end{figure}
 
-For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate with 1000 data points (see Fig.~\ref{fig:posterior_samples} in the appendix).
-We then compare the posterior mean to the ground truth value.
-For a single star observation, our NPE has an APE of \variable{output/posterior_APE.txt} when looking at all 6 parameters of interest.
-If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/global_posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}.
+For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate with 1000 points. %(see Fig.~\ref{fig:posterior_samples} in the appendix)
+We then compare the posterior mean to the ground truth value for each observation and calculate the absolute percentage error (APE, see Fig.~\ref{fig:posterior_APE}).
+For a single star observation, our NPE has an APE of \variable{output/global_posterior_APE.txt} when looking at all 6 parameters of interest.
+If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}.
 Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high.
@@
 \label{sec:appendix_code_and_data}
-%The source code for this manuscript is made available under an open source license and is publicly hosted on GitHub under this url: \GitHubURL.\\
-% The github repository accompanying this article includes comprehensive documentation %hosted on \emph{ReadTheDocs} that provides an overview of the project, installation instructions, and a guide on how to use the software.
 
@@
 
-% \textcolor{red}{To be finalized, do we want to upload data???}
-% The cleaned data set with outliers removed is publicly available on Zenodo.%\footnote{URL: \url{https://zenodo.org/record/8375344}}
+\section{Simulation-based calibration}
+\label{sec:sbc}
+
+\begin{figure*}[]
+     \centering
+     \includegraphics[width=\columnwidth]{figures/sbc_rank_plot_NPE_C.pdf}
+     \includegraphics[width=\columnwidth]{figures/tarp_plot_NPE_C.pdf}
+     \vspace{-.5cm}
+     \caption{\emph{Left:} SBC ranks of ground truth parameters under the inferred posterior samples for each of the six parameters (red bars).
+The grey area shows the 99\% confidence interval of a uniform distribution given the number of samples provided.
+\emph{Right:} }
+     \label{fig:sbc}
+     \script{evaluate_sbi.py}
+\end{figure*}
+
+Since SBI relies in neural networks to approximate posterior densities one important point is to check that neural network hyperparameters are well chosen and that posterior estimates are trustable.
+
+After a density estimator has been trained with simulated data to obtain a posterior, the estimator should be made subject to several diagnostic tests.
+This needs to be performed before being used for inference given the actual observed data.
+Posterior Predictive Checks provide one way to “critique” a trained estimator based on its predictive performance.
+Another important approach to such diagnostics is simulation-based calibration as developed by \citet{Cook2006} and \citet{Talts2018}.
+
+\paragraph{Simulation-based calibration}
+Simulation-based calibration (SBC) provides a (qualitative) view and a quantitive measure to check, whether the variances of the posterior are balanced, i.e.
+it is neither over-confident nor under-confident.
+As such, SBC can be viewed as a necessary condition (but not sufficient) for a valid inference algorithm: If SBC checks fail, this tells you that your inference is invalid.
+If SBC checks pass, this is no guarantee that the posterior estimation is working.
+
+To perform SBC, we sample some $\theta_i^o$ values from the parameter prior of the problem at hand and simulate “observations” from these parameters: 
+\begin{equation}
+    x_i = \text{simulator}(\theta_i^o)
+\end{equation}
+Then we perform inference given each observation $x_i$ which produces a separate posterior $p_i(\theta|x_i)$ for each $x_i$.
+The key step for SBC is to generate a set of posterior samples $\{\theta\}_i$ from each posterior.
+We call this $\theta_i^s$, referring to $s$ samples from the posterior $p_i(\theta|x_i)$.
+Next, we rank the corresponding $\theta_i^o$ under this set of samples.
+A rank is computed by counting how many samples $\theta_i^s$ fall below their corresponding $\theta_i^o$ value \citep[see section 4.1 in][]{Talts2018}.
+These ranks are then used to perform the SBC check itself.
+
+The core idea behind SBC is two fold: (i) SBC ranks of ground truth parameters under the inferred posterior samples follow a uniform distribution (If the SBC ranks are not uniformly distributed, the posterior is not well calibrated); and (ii) samples from the data averaged posterior (ensemble of randomly chosen posterior samples given multiple distinct observations $x_o$) are distributed according to the prior.
+
+Hence, SBC can tell us whether the SBI method applied to the problem at hand produces posteriors that have well-calibrated uncertainties, and if the posteriors have uncalibrated uncertainties, SBC surfaces what kind of systematic bias is present: negative or positive bias (shift in the mean of the predictions) or over- or under-dispersion (too large or too small variance).
+
+In the left panel of Fig.~\ref{fig:sbc} we show the distribution of ranks (depicted in red) in each dimension.
+Highlighted in grey, you see the 99\% confidence interval of a uniform distribution given the number of samples provided.
+In plain english: for a uniform distribution, we would expect 1 out of 100 (red) bars to lie outside the grey area.
+This figure shows that overall our posteriors are decently calibrated.
+Only for the parameters $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ and stellar age $T_i$ see single bars outside the 99\% confidence interval.
+But most importantly for the parameters of interest here, $\log_{10}\left(N_\mathrm{Ia}\right)$ and $\alpha_\mathrm{IMF}$ we have a well calibrated posterior.
+
+\paragraph{Tests of Accuracy with Random Points (TARP)}
+
+TARP \citep{Lemos2023} is an alternative calibration check.
+Given a test set $(\theta^*,x^*)$ and a set of reference points $\theta_r$, TARP calculates statistics for posterior calibration by - drawing posterior samples $\theta$ given each observation $x^*$ and calculating the distance $r$ between $\theta^*$ and $\theta_r$ counting for how many of the posterior samples the distance to $\theta_r$ is smaller than $r$ \citep[see e.g.
+Fig.~2 in][for an illustration]{Lemos2023}.
+
+For each given coverage level $\alpha$, one can then calculate the corresponding average counts and check, whether they correspond to the given $\alpha$.
+The visualization and interpretation of TARP values is therefore similar to that of SBC.
+However, in contrast to SBC TARP provides a necessary and sufficient condition for posterior accuracy, i.e., it can also detect inaccurate posterior estimators.
+
+Note, however, that this property depends on the choice of reference point distribution: to obtain the full diagnostic power of TARP, one would need to sample reference points from a distribution that depends on $x$.
+Thus, in general, it is recommended using and interpreting TARP like SBC and complementing coverage checks with posterior predictive checks.
 

### 2025-01-17 src/scripts/evaluate_sbi.py

@@ -10,6 +10,7 @@ from sbi.diagnostics import plot_tarp, check_tarp, run_sbc, run_tarp
 
 from Chempy.parameter import ModelParameters
 
+from chempy_torch_model import Model_Torch
 from plot_functions import *
 
 # ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
@@ -32,6 +33,13 @@ combined_priors = utils.MultipleIndependent(
 # ----- load the posterior -------------------------------------------------------------------------------------------------------------------------------------------
 with open(paths.data / f'posterior_{name}.pickle', 'rb') as f:
     posterior = pickle.load(f)
+
+# --- Set up the model ---
+model = Model_Torch(len(labels_in), len(labels_out))
+
+# --- Load the weights ---
+model.load_state_dict(torch.load(paths.data / 'pytorch_state_dict.pt'))
+model.eval()
 
 # --- Simulation based calibration plot ---
 def simulator(params):

### 2025-01-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
      \centering
-     \includegraphics[width=\columnwidth]{figures/ape_NN.png}
+     \includegraphics[width=\columnwidth]{figures/ape_NN.pdf}
      \vspace{-.5cm}
@@
 This needs to be performed before being used for inference given the actual observed data.
-Posterior Predictive Checks provide one way to “critique” a trained estimator based on its predictive performance.
+Posterior Predictive Checks provide one way to "critique" a trained estimator based on its predictive performance.
 Another important approach to such diagnostics is simulation-based calibration as developed by \citet{Cook2006} and \citet{Talts2018}.
@@
 
-To perform SBC, we sample some $\theta_i^o$ values from the parameter prior of the problem at hand and simulate “observations” from these parameters: 
+To perform SBC, we sample some $\theta_i^o$ values from the parameter prior of the problem at hand and simulate "observations" from these parameters: 
 \begin{equation}

### 2025-01-17 src/scripts/evaluate_sbi.py

@@ -5,8 +5,8 @@ from torch.distributions.normal import Normal
 from torch.distributions.uniform import Uniform
 
 import sbi.utils as utils
-from sbi.analysis.plot import sbc_rank_plot
-from sbi.diagnostics import plot_tarp, check_tarp, run_sbc, run_tarp
+from sbi.analysis.plot import sbc_rank_plot, plot_tarp
+from sbi.diagnostics import check_tarp, run_sbc, run_tarp
 
 from Chempy.parameter import ModelParameters
 

### 2025-01-17 src/scripts/evaluate_sbi.py
[showyourwork] overleaf sync

@@ -76,7 +76,7 @@ f, ax = sbc_rank_plot(
 
 f.suptitle("SBC rank plot", fontsize=36)
 plt.tight_layout()
-plt.savefig(paths.figures + f'sbc_rank_plot_{name}.pdf')
+plt.savefig(paths.figures / f'sbc_rank_plot_{name}.pdf')
 plt.clf()
 
 # --- TARP plot ---
@@ -98,5 +98,5 @@ print(ks_pval, "Should be larger than 0.05")
 # Or, we can perform a visual check.
 f, ax = plot_tarp(ecp, alpha)
 plt.tight_layout()
-plt.savefig(paths.figures + f'tarp_plot_{name}.pdf')
+plt.savefig(paths.figures / f'tarp_plot_{name}.pdf')
 plt.clf()

### 2025-01-17 src/scripts/evaluate_sbi.py
[showyourwork] overleaf sync

@@ -70,11 +70,12 @@ f, ax = sbc_rank_plot(
     parameter_labels=labels_in,
     plot_type="hist",
     num_cols=3,
-    figsize=(15,10),
+    figsize=(5,3.3),
     num_bins=None,  # by passing None we use a heuristic for the number of bins.
 )
 
-f.suptitle("SBC rank plot", fontsize=36)
+
+f.suptitle("SBC rank plot", fontsize=17)
 plt.tight_layout()
 plt.savefig(paths.figures / f'sbc_rank_plot_{name}.pdf')
 plt.clf()

### 2025-01-17 src/scripts/evaluate_sbi.py
[showyourwork] overleaf sync

@@ -70,7 +70,7 @@ f, ax = sbc_rank_plot(
     parameter_labels=labels_in,
     plot_type="hist",
     num_cols=3,
-    figsize=(5,3.3),
+    figsize=(8,5),
     num_bins=None,  # by passing None we use a heuristic for the number of bins.
 )
 
@@ -97,7 +97,7 @@ print(atc, "Should be close to 0")
 print(ks_pval, "Should be larger than 0.05")
 
 # Or, we can perform a visual check.
-f, ax = plot_tarp(ecp, alpha)
+f, ax = plot_tarp(ecp, alpha, figsize=(10, 10))
 plt.tight_layout()
 plt.savefig(paths.figures / f'tarp_plot_{name}.pdf')
 plt.clf()

### 2025-01-17 src/scripts/evaluate_sbi.py
[showyourwork] overleaf sync

@@ -97,7 +97,7 @@ print(atc, "Should be close to 0")
 print(ks_pval, "Should be larger than 0.05")
 
 # Or, we can perform a visual check.
-f, ax = plot_tarp(ecp, alpha, figsize=(10, 10))
+f, ax = plot_tarp(ecp, alpha)
 plt.tight_layout()
 plt.savefig(paths.figures / f'tarp_plot_{name}.pdf')
 plt.clf()

### 2025-01-17 src/scripts/evaluate_sbi.py
[showyourwork] overleaf sync

@@ -70,7 +70,7 @@ f, ax = sbc_rank_plot(
     parameter_labels=labels_in,
     plot_type="hist",
     num_cols=3,
-    figsize=(8,5),
+    figsize=(10,6.67),
     num_bins=None,  # by passing None we use a heuristic for the number of bins.
 )
 

### 2025-01-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
      \centering
-     \includegraphics[width=\columnwidth]{figures/sbc_rank_plot_NPE_C.pdf}
-     \includegraphics[width=\columnwidth]{figures/tarp_plot_NPE_C.pdf}
-     \vspace{-.5cm}
-     \caption{\emph{Left:} SBC ranks of ground truth parameters under the inferred posterior samples for each of the six parameters (red bars).
-The grey area shows the 99\% confidence interval of a uniform distribution given the number of samples provided.
-\emph{Right:} }
+     \includegraphics[width=\textwidth]{figures/sbc_rank_plot_NPE_C.pdf}
+     %\vspace{-.5cm}
+     \caption{SBC ranks of ground truth parameters under the inferred posterior samples for each of the six parameters (red bars).
+The grey area shows the 99\% confidence interval of a uniform distribution given the number of samples provided.}
      \label{fig:sbc}
@@
 \end{figure*}
+
+\begin{figure}[]
+     \centering
+     \includegraphics[width=\columnwidth]{figures/tarp_plot_NPE_C.pdf}
+     %\vspace{-.5cm}
+     \caption{TARP plot.}
+     \label{fig:sbc}
+     \script{evaluate_sbi.py}
+\end{figure}
 

### 2025-01-17 src/scripts/evaluate_sbi.py

@@ -75,7 +75,7 @@ f, ax = sbc_rank_plot(
 )
 
 
-f.suptitle("SBC rank plot", fontsize=17)
+f.suptitle("SBC rank plot", fontsize=13)
 plt.tight_layout()
 plt.savefig(paths.figures / f'sbc_rank_plot_{name}.pdf')
 plt.clf()

### 2025-01-21 src/tex/ms.tex
test new python version
[showyourwork] fix for overlef sync
[showyourwork] overleaf sync

@@
 We design the NN as a simple feed-forward neural network with 2 hidden layers and 100 neurons in the first and 40 neurons in the second layer.
+Training this tiny emulator takes about 50s on the CPU.
 
@@
 In particular, if we marginalize over the star specific parameters and solely focus on the global parameters $\Vec{\Lambda}$ we can make the assumption that individual observations of stars are identically and independently distributed (i.i.d.) and factorize the joint posterior from above to simply express it as:  %
-\begin{eqnarray}\label{eq:posterior}
+\begin{eqnarray}
 \mathbb{P}(\vec\Lambda|\{D_i\}) &\propto& 
@@
 &\propto& \mathbb{P}(\vec\Lambda) \prod_{j=1}^{n_\mathrm{star}}\frac{\mathbb{P}(\vec\Lambda|D_j)}{\mathbb{P}(\vec\Lambda)} \quad\quad\quad\quad\quad\text{(Bayes rule)}
-\\\nonumber
+\\%\nonumber
+\label{eq:posterior}
 &=& \mathbb{P}(\vec\Lambda)^{1-n_\mathrm{star}}\ \prod_{j=1}^{n_\mathrm{star}}\mathbb{P}(\vec\Lambda|D_j) 
@@
 
-We train our NPE with $10^5$ data points simulated with the NN emulator described in the previous section until convergence.
+We train our NPE with $10^5$ data points consisting of $n_{\rm elemets}=8$ simulated with the NN emulator described in the previous section with inputs sampled from the priors shown in Tab.
+\ref{tab:priors}.
 Training takes $\lesssim10$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPUs.
@@
 In order to mimick observational noise, we add 5\% observational uncertainties to the abundances simulated with \texttt{CHEMPY} before feeding them to our NPE.
+We have made sure that our NPE is well calibrated and posterior distributions are trustable. For more details on simulation-based calibration see appendix \ref{sec:sbc}.
 
@@
 
-For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate with 1000 points. %(see Fig.~\ref{fig:posterior_samples} in the appendix)
-We then compare the posterior mean to the ground truth value for each observation and calculate the absolute percentage error (APE, see Fig.~\ref{fig:posterior_APE}).
+\paragraph{Multi-star inference}
+Following equation~\ref{eq:posterior} we can calculate the joint posterior for a combined inference using abundance observations of multiple stars.
+For this, we will condition our NPE individually on single star observations and sample the posterior.
+Then, according to equation~\ref{eq:posterior} we simply have to multiply individual posteriors and account for the inverse weighting by some power of the prior which we take from table~\ref{tab:priors}.
+
+Note however, there is one issue here.
+Since we only have access to samples from the posterior and not the posterior itself it is difficult to evaluate equation~\ref{eq:posterior}.
+We circumvent this by approximating each single star posterior by a Gaussian and fit for the parameters of mean and covariance.
+With this it is straight forward to evaluate equation~\ref{eq:posterior} analytically.
+In fact, under our assumption the combined posterior is a product of the prior and the product of multiple Gaussians for the individual star posteriors.
+The latter product is also a Gaussian with mean $\mathbf{\mu'}$ and variance $\mathbf{\sigma'}$:
+\begin{align}
+    \mathbf{\mu'} &= \frac{\sum_{i=1}^{N_{stars}} \frac{\mu_i}{\sigma_i^2}}{\sum_{i=1}^{N_{stars}} \frac1{\sigma_i^2}} \\
+\mathbf{\sigma'}^2 &= \frac1 {\sum_{i=1}^{N_{stars}} \frac1{\sigma_i^2}}
+\end{align}
+
+Further, in our case the prior for the galactic parameters $\Lambda$ is Gaussian as well.
+Therefore the resulting factorized posterior from eq.~\ref{eq:posterior} is again a Gaussian and can be expressed with mean $\mathbf{\mu}$ and variance $\mathbf{\sigma}$ as:
+\begin{align}
+\mathbf{\mu} &= \frac{\frac{\mu'}{\sigma'^2}-\frac{(1-N)\mu_ {prior}}{\sigma_ {prior}^2}}{\frac1{\sigma'^2}-\frac{(1-N)}{\sigma_ {prior}^2}} \\
+\mathbf{\sigma}^2 &= \frac1 {\frac1{\sigma'^2}-\frac{(1-N)}{\sigma_ {prior}^2}}
+\end{align}
+
+Note, given the tiny and simple neural network that represents our NPE the above assumption of Gaussian single star posteriors and the consecutive simplification is a decent one to make.
+We have further empirically verified that single star posteriors are indeed close to Gaussian.
+However, in future work we plan to alleviate this simplification and directly approximate the joint posterior of a multi-star inference.
+
+\section{Results}
+\label{sec: Results}
+
+\textcolor{red}{see philcox results for boiler plate text...}
+
+We test our analysis using mock observations drawn firstly from \texttt{CHEMPY} from the same yield set as the training data, secondly using a different yield set to test for potential biases through model missspecification in our SBI setup and then lastly from a large-scale hydrodynamical simulation taken from the IlustrisTNG suite \citep{Pillepich2018} but with the same yield set as our \texttt{CHEMPY} training data to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
+The methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
+
+\subsection{\texttt{CHEMPY} Mock Observational Data}
+Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
+\ref{tab:priors}, additionally drawing $T_i$ uniformly in the range $[2,12.8]$ Gyr, to minimize overlap with the neural network training birth-time limits when observational uncertainties are included.
+Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
+
+For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with 1000 points.
+This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
+Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1000 stars (each time sampling the single star posterior with 1000 points and fitting for the Gaussian parameters).
+Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}.
+Hence, shorter computing times make it feasible to use orders of magnitudes more observations.
+
+\paragraph{Validation through absolute percentage error}
+To evaluate our method quantitatively, we compare the posterior mean to the ground truth value for each observation and calculate the absolute percentage error (APE, see Fig.~\ref{fig:posterior_APE}).
 For a single star observation, our NPE has an APE of \variable{output/global_posterior_APE.txt} when looking at all 6 parameters of interest.
@@
 
-\section{Results}
-\label{sec: Results}
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
+    \vspace{-.25cm}
+    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) as done by \cite{Philcox_2019}.}
+    \label{fig:CHEMPY_TNG_N_star_analysis}
+    \script{chempy_tng_inference.py}
+\end{figure}
 
-We test our analysis using mock observations drawn firstly from \texttt{Chempy} then from large-scale hydrodynamical simulations taken from the IlustrisTNG suite to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
-The methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
+\paragraph{Inference on the \texttt{CHEMPY} TNG yield set}
 
-% % \begin{figure}
-% %     \centering
-% %     \includegraphics[width=0.9\linewidth]{figures/sbi_Nstar_comp.png}
-% %     \vspace{-.25cm}
-% %     \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) \cite{Philcox_2019} .}
-% %     \label{fig:N_star_analysis}
-% % \end{figure}
+\begin{figure}
+%\vspace{-1.3cm}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields.pdf}
+    \vspace{-.25cm}
+    \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
+    \label{fig:CHEMPY_TNG_sbi} 
+    \script{chempy_tng_inference.py}
+\end{figure}
 
-We train the NPE with a set of $n_{\rm stars}=10^6$ abundances for $n_{\rm elemets}=8$ with an assumed observational error of $0.05$ dex, generated with the \textit{neural network} emulator, with inputs sampled from the priors shown in Tab.
-\ref{tab:priors}.
-The sole purpose of the emulator is to highly reduce the computational time compared to full \texttt{CHEMPY} simulations.
-Training the NPE takes $~3$h on a single H100 GPU.
-
-\paragraph{Mock Observational Data}
-Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
-\ref{tab:priors}, additionally drawing $T_i$ uniformly in the range $[2,12.8]$ Gyr, to minimize overlap with the neural network training birth-time limits when observational uncertainties are included.
-Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
-\textcolor{red}{here goes the assumption that stars are iid since each stars is produced by a single call to chempy.
-However, real physics might correlate stars abundacnes by star formation histories and cumulative enrichment...}
-The outcome of this mock data creation is a set of 1000 stars \textcolor{red}{we vary this number, right?}, emulating a real dataset.
-
-\paragraph{Inference}
-
-% % \begin{figure}
-% % \vspace{-1.3cm}
-% %     \centering
-% %     \includegraphics[width=0.49\textwidth]{figures/posterior_2.png}
-% %     \vspace{-.25cm}
-% %     \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
-% %     \label{fig:sbi}   
-% % \end{figure}
-
-For each set of mock abundances we infer the posterior distribution for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with the NPE by sampling 1000 times and fitting the posterior for each parameter.
-This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
-Our method takes around $5$ minutes to build a posterior for all six parameters for a dataset size of 1000 stars on a Macbook Air M2 chip, making it around $2400$x faster than current HMC methods \citep[cf.][who needs $40$h for 200 stars with HMC]{Philcox_2019}.
-Hence, shorter computing times make it feasible to use orders of magnitudes more observations.
 %
@@
 
-%     % \textcolor{blue}{Number of stars is not limited.
-I used $1000$ stars with an inference time of around $5$ minutes, but accuracy doesn't improve much after $\sim200$ stars [see \ref{fig:N_star_analysis}].
-Philcox inferred $200$ stars in $\sim40$h}
-%     % \textcolor{red}{awesome! that gives us a nice metric to quote speedup. can we specify the architecture we used? I.e. which GPU on the clyster did you use?}
-%     % \textcolor{blue}{For training the sbi net i used compgpu11 and it took around 3h. The inference of the stars i did on my laptop (macbook air with a M2 chip) in 5 minutes}
-
 \subsection{Alternative yield set}
 \label{subsec: mocks_wrong_yield}
+
+\section{Discussion}
+\label{sec: conclusion}
+
+\textcolor{red}{here goes the assumption that stars are iid since each stars is produced by a single call to chempy.
+However, real physics might correlate stars abundacnes by star formation histories and cumulative enrichment...}
 
@@
      %\vspace{-.5cm}
-     \caption{TARP plot.}
-     \label{fig:sbc}
+     \caption{TARP plot showing the expected coverage probability vs.
+the credibility level $\alpha$.
+The dashed black 1:1-line shows an ideal calibrated posterior and the blue solid line shows the TARP value for our NPE.}
+     \label{fig:tarp}
      \script{evaluate_sbi.py}
@@
 This figure shows that overall our posteriors are decently calibrated.
-Only for the parameters $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ and stellar age $T_i$ see single bars outside the 99\% confidence interval.
+Only for the parameter $\log_{10}\left(\mathrm{SFE}\right)$ we see a single bar outside the 99\% confidence interval.
 But most importantly for the parameters of interest here, $\log_{10}\left(N_\mathrm{Ia}\right)$ and $\alpha_\mathrm{IMF}$ we have a well calibrated posterior.
@@
 The visualization and interpretation of TARP values is therefore similar to that of SBC.
-However, in contrast to SBC TARP provides a necessary and sufficient condition for posterior accuracy, i.e., it can also detect inaccurate posterior estimators.
+However, in contrast to SBC, TARP provides a necessary and sufficient condition for posterior accuracy, i.e., it can also detect inaccurate posterior estimators.
+In Fig.~\ref{fig:tarp} we show the result for our NPE in blue in comparison to the ideal line shown in black dashed style.
+This figure clearly shows that our NPE is well calibrated.
 

### 2025-01-22 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -251,13 +251,12 @@ def n_stars_plot(x1, x2, x_true, no_stars= np.array([1, 10, 100, 500, 1000]), si
         ax.tick_params(which='minor', size=5, width=2)
 
     for i, name in enumerate([r'$\alpha_{\rm IMF}$', r'$\log_{10} N_{\rm Ia}$']):
-        plot(fit[:,i], err[:,i], x_true[i], ax[i], name)
+        plot(fit[:,i], err[:,i], x_true[0,i], ax[i], name)
 
     ax[0].legend(fontsize=15, fancybox=True, shadow=True)
     
     plt.tight_layout()
     plt.savefig(paths.figures / f'{save_name}.pdf')
-    plt.show()
 
 ##########################################################################################################
 # --- N-Star comparison plot ---
@@ -297,7 +296,7 @@ def n_stars_plot_comp(x1, x2, x_true, dat, save_name, no_stars = np.array([1, 10
     err = np.array(err)
 
     # --- Plot the data ---
-    fig,ax=plt.subplots(nrows=1,ncols=2,figsize=(26,6))
+    fig, ax = plt.subplots(nrows=1,ncols=2,figsize=(26,6))
 
     def plot(fit, err, x_true, ax, name):
         ax.plot(no_stars, fit, color="b", label="Fit")
@@ -319,13 +318,13 @@ def n_stars_plot_comp(x1, x2, x_true, dat, save_name, no_stars = np.array([1, 10
         ax.fill_between(n_stars,lo2[:,i],up2[:,i],alpha=0.1,color='r')
 
     for i, name in enumerate([r'$\alpha_{\rm IMF}$', r'$\log_{10} N_{\rm Ia}$']):
-        plot(fit[:,i], err[:,i], x_true[i], ax[i], name)
+        plot(fit[:,i], err[:,i], x_true[0,i], ax[i], name)
 
     ax[0].legend(fontsize=20, fancybox=True, shadow=True)
     
     plt.tight_layout()
     plt.savefig(paths.figures / f'{save_name}.pdf')
-    plt.show()
+    plt.clf()
 
 
 ##########################################################################################################
@@ -349,12 +348,12 @@ def n_stars_plot_comp2(x1, x2, x_true, philcox, save_name, no_stars = np.array([
     err = np.array(err)
 
     # --- Plot the data ---
-    fig,ax=plt.subplots(nrows=1,ncols=2,figsize=(26,6))
[…]

### 2025-01-22 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -365,6 +365,7 @@ def n_stars_plot_comp2(x1, x2, x_true, philcox, save_name, no_stars = np.array([
         ax.tick_params(labelsize=30, size=10, width=3)
         ax.tick_params(which='minor', size=5, width=2)
 
+        print(philcox['med'][:,i])
         # Add Philcox
         ax.plot(philcox['n_stars'],philcox['med'][:,i],c='r', label="HMC")
         ax.fill_between(philcox['n_stars'],philcox['lo'][:,i],philcox['up'][:,i],alpha=0.2,color='r')
@@ -467,4 +468,3 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
 
     plt.tight_layout()
     plt.savefig(paths.figures / f'{title}.pdf')
-    plt.show()

### 2025-01-22 src/scripts/chempy_tng_inference.py
[showyourwork] overleaf sync

@@ -43,10 +43,10 @@ gaussian_posterior_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHE
 
 # --- Compare to HMC ---
 philcox = {}
-philcox["n_stars"] = [1,10,100]
-philcox["med"] = [[-2.29,-2.87],[-2.31,-2.90],[-2.31,-2.90]]
-philcox["up"] = [[-2.21,-2.76],[-2.29,-2.87],[-2.30,-2.89]]
-philcox["lo"] = [[-2.37,-2.98],[-2.33,-2.93],[-2.32,-2.91]]
+philcox["n_stars"] = np.array([1,10,100])
+philcox["med"] = np.array([[-2.29,-2.87],[-2.31,-2.90],[-2.31,-2.90]])
+philcox["up"] = np.array([[-2.21,-2.76],[-2.29,-2.87],[-2.30,-2.89]])
+philcox["lo"] = np.array([[-2.37,-2.98],[-2.33,-2.93],[-2.32,-2.91]])
 
 stars = np.arange(1,1000)
 n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_N_star_comp")

### 2025-01-22 src/scripts/chempy_tng_inference.py
[showyourwork] overleaf sync

@@ -49,4 +49,4 @@ philcox["up"] = np.array([[-2.21,-2.76],[-2.29,-2.87],[-2.30,-2.89]])
 philcox["lo"] = np.array([[-2.37,-2.98],[-2.33,-2.93],[-2.32,-2.91]])
 
 stars = np.arange(1,1000)
-n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_N_star_comp")
+n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_Nstar_comp")

### 2025-01-22 src/scripts/plot_functions.py

@@ -336,7 +336,6 @@ def n_stars_plot_comp2(x1, x2, x_true, philcox, save_name, no_stars = np.array([
 
     # --- Fit a 2D Gaussian to the data ---
     for n in no_stars:
-        print(n.dtype)
         mu_alpha, sigma_alpha = mean_std(x1[:n], x_true[0,0], x_true[1,0])
         mu_logNIa, sigma_logNIa = mean_std(x2[:n], x_true[0,1], x_true[1,1])
 

### 2025-01-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}.
+But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim72.000$.
+\textcolor{red}{check if those speed-up values are correctly calculated...}
 Hence, shorter computing times make it feasible to use orders of magnitudes more observations.

### 2025-01-22 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -463,7 +463,7 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
     plt.gca().add_artist(legend_fit)
     plt.gca().add_artist(legend_true)
     
-    plt.title(title, fontsize=60)
+    plt.title(title, fontsize=40)
 
     plt.tight_layout()
-    plt.savefig(paths.figures / f'{title}.pdf')
+    plt.savefig(paths.figures / f'{title}.png')

### 2025-01-22 src/scripts/chempy_tng_inference.py
[showyourwork] overleaf sync

@@ -49,4 +49,4 @@ philcox["up"] = np.array([[-2.21,-2.76],[-2.29,-2.87],[-2.30,-2.89]])
 philcox["lo"] = np.array([[-2.37,-2.98],[-2.33,-2.93],[-2.32,-2.91]])
 
 stars = np.arange(1,1000)
-n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_Nstar_comp")
+n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_Nstar_comp", stars)

### 2025-01-22 src/scripts/plot_functions.py

@@ -440,8 +440,8 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
     legend_true = plt.scatter(global_params[0,0], global_params[0,1], color='red', label=label_gt, s=100)
     legend_fit = plt.scatter(mu_alpha, mu_log10N_Ia, color='k', label=label_fit, s=100)
     
-    legend_fit = plt.legend(handles=[legend_fit], fontsize=15, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.9))
-    legend_true = plt.legend(handles=[legend_true], fontsize=15, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.99))
+    legend_fit = plt.legend(handles=[legend_fit], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.1))
+    legend_true = plt.legend(handles=[legend_true], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.19))
     
 
     # Sigma levels

### 2025-01-22 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -440,8 +440,8 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
     legend_true = plt.scatter(global_params[0,0], global_params[0,1], color='red', label=label_gt, s=100)
     legend_fit = plt.scatter(mu_alpha, mu_log10N_Ia, color='k', label=label_fit, s=100)
     
-    legend_fit = plt.legend(handles=[legend_fit], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.1))
-    legend_true = plt.legend(handles=[legend_true], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.19))
+    legend_fit = plt.legend(handles=[legend_fit], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.2))
+    legend_true = plt.legend(handles=[legend_true], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.4))
     
 
     # Sigma levels

### 2025-01-22 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -441,7 +441,7 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
     legend_fit = plt.scatter(mu_alpha, mu_log10N_Ia, color='k', label=label_fit, s=100)
     
     legend_fit = plt.legend(handles=[legend_fit], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.2))
-    legend_true = plt.legend(handles=[legend_true], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.4))
+    legend_true = plt.legend(handles=[legend_true], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.35))
     
 
     # Sigma levels

### 2025-01-22 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -467,3 +467,4 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
 
     plt.tight_layout()
     plt.savefig(paths.figures / f'{title}.png')
+    plt.clf()

### 2025-01-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\textcolor{red}{see philcox results for boiler plate text...}
+We use our SBI method described in the previous section to infer the global galactic parameters $\alpha_{\rm {IMF}}$ and $\log_{10}(N_{\rm{Ia}})$.
+In order to demonstrate the performance and robustness of our methods we use three mock data-sets:
+\begin{enumerate}
+    \item Mock observations drawn from \texttt{CHEMPY} from the same yield set as the training data for the neural network emulator.
+With this we esnure to test our training strategy and the performance without any systematic distribution shifts.
+    \item  \texttt{CHEMPY} mock data using a different yield set to test for potential biases through model missspecification in our SBI setup.
+    \item Simulated data from a large-scale hydrodynamical simulation taken from the IlustrisTNG suite \citep{Pillepich2018} but with the same yield set as our \texttt{CHEMPY} training data to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
+\end{enumerate}
 
-We test our analysis using mock observations drawn firstly from \texttt{CHEMPY} from the same yield set as the training data, secondly using a different yield set to test for potential biases through model missspecification in our SBI setup and then lastly from a large-scale hydrodynamical simulation taken from the IlustrisTNG suite \citep{Pillepich2018} but with the same yield set as our \texttt{CHEMPY} training data to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
-The methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
+Note, the methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
 
 \subsection{\texttt{CHEMPY} Mock Observational Data}
-Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and the local parameters $\vec{\Theta}_i$ from Tab.
-\ref{tab:priors}, additionally drawing $T_i$ uniformly in the range $[2,12.8]$ Gyr, to minimize overlap with the neural network training birth-time limits when observational uncertainties are included.
-Each set of parameters is passed to the NN, producing eight true chemical abundances, to which we add a Gaussian error of $0.05$ dex.
+Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and varying local parameters $\vec{\Theta}_i$ using priors from Tab.
+\ref{tab:priors}.
+Additionally we draw $T_i$ uniformly in the range $[2,12.8]$ Gyr to minimize overlap with the neural network training birth-time limits when observational uncertainties are included.
+Each set of parameters is passed to our \texttt{CHEMPY} emulator, producing eight true chemical abundances.
+In order to fully represent observational data, we augment this data with observational uncertainties by adding a Gaussian error of $0.05$ dex for the abundances representative for typical APOGEE data \citep{Majewski2016}.
+All our datasets and accompanying code to reproduce our data generation process and analysis is publicly available\footnote{\url{https://github.com/TobiBu/sbi-chempy}}.
 
@@
 \paragraph{Validation through absolute percentage error}
+We start evaluating our method by using \texttt{CHEMPY} to produce a mock observational dataset to ensure no systematical shft between training and testing data in terms of physics parameters.
+
 To evaluate our method quantitatively, we compare the posterior mean to the ground truth value for each observation and calculate the absolute percentage error (APE, see Fig.~\ref{fig:posterior_APE}).
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
-    \vspace{-.25cm}
-    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue) to the inferred values using HMC (red) as done by \cite{Philcox_2019}.}
+    \includegraphics[width=\textwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
+    \vspace{-.5cm}
+    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
+For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
     \label{fig:CHEMPY_TNG_N_star_analysis}
     \script{chempy_tng_inference.py}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields.pdf}
+    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields.png}
     \vspace{-.25cm}
-    \caption{Sampled posteriors for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.}
+    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.
+The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
     \label{fig:CHEMPY_TNG_sbi} 
@@
 %
-Taking the means of multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$ (Fig.
-\ref{fig:N_star_analysis}).
-However, we also see that in the limit of very few stars (less than $\sim10$) SBI shows a larger bias than HMC results.
+Combining the inference on multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$.
+We therefore perform inference using a range of stars $n\in[1,1000]$ (see Fig.
+\ref{fig:CHEMPY_TNG_N_star_analysis}) to show how inference accuracy increases with number of observations.
+We find that in the limit of few stars (less than $\sim100$) SBI shows a larger uncertainties than the HMC results of \citet{Philcox_2019}.
 But in the limit of large numbers of stars (few hundred) the accuracy and precision is superior compared to HMC.
+In particular, after using a few ten stars, the NPE estimate is less biased than the HMC results.
+Finally, given our computational advantage we will be able to use orders of magnitude more stars for our inference.
+This is particular important since sample variances play an important role when using small samples of stars as also noted in \citet{Philcox_2019}.
 %
-To infer the global galactic parameters $\vec\Lambda$ of the stars as accurate as possible we take the mean of all inferred posteriors from $n_{\rm stars}=1000$ (Fig.
-\ref{fig:sbi}) resulting in $\alpha_{\rm IMF}=-2.292\pm0.003$ and $\log_{10}(\rm N_{Ia})=-2.883\pm0.005$ which deviates less than $0.4\%$ from the correct value.
+
+In Fig.~\ref{fig:CHEMPY_TNG_sbi} we show the joint posterior for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ for our inference.
+The red star indicates the ground truth value, the black dot shows our posterior mean and white contours indicate $1-3\sigma levels$.
+
+Using a sample of 1000 stars we infer \variable{output/CHEMPY_TNG_sbi.txt} which deviates less than $\sim0.04\%$ from the correct value.
+
+Note, as shown in Fig.~\ref{fig:posterior_APE} in principle our analysis is also able to infer the local parameters $\Vec{\Theta}_i$ and $T_i$.
+This would allow us to estimate/infer stellar ages as well.
+However, currently our NPE network is not particularly good at estimating ages from abundances alone.
+On avergae our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during mock up of our data.
+
+In summary, our SBI pipeline is quite capable of correctly and precisely inferring  global parameters of chemical enrichment models from stellar abundance alone when using the same physical model and yield tables as during training.
+Next, we will check what happens if for some reason the training data was generated with a different yield set.
 

### 2025-01-23 src/scripts/chempy_tng_inference.py

@@ -29,6 +29,9 @@ sigma_alpha_combined = 1/np.sqrt(np.sum(1/sigma_alpha**2))
 mu_N_Ia_combined = np.sum(mu_N_Ia/sigma_N_Ia**2)/np.sum(1/sigma_N_Ia**2)
 sigma_N_Ia_combined = 1/np.sqrt(np.sum(1/sigma_N_Ia**2))
 
+with open(paths.output / f'CHEMPY_TNG_sbi.txt', 'w') as f:
+    f.write(f"$\alpha_{\rm IMF}={mu_alpha_combined:.3f}\pm{sigma_alpha_combined:.3f}$ and $\log_{10}(\rm N_{Ia})={mu_N_Ia_combined:.3f} \pm {sigma_N_Ia_combined:.3f}$%")
+
 print(f"alpha_IMF = {mu_alpha_combined:.3f} +/- {sigma_alpha_combined:.3f}")
 print(f"log10_N_Ia = {mu_N_Ia_combined:.3f} +/- {sigma_N_Ia_combined:.3f}")
 

### 2025-01-23 src/scripts/plot_functions.py

@@ -437,7 +437,7 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
     label_gt = r'Ground Truth' + f"\n" + r"$\alpha_{\rm IMF} = $" + f'${global_params[0,0]:.2f}$' + f"\n" + r"$\log_{10} N_{\rm Ia} = $" + f'${global_params[0,1]:.2f}$'
     label_fit = r'Fit' + f"\n" + r"$\alpha_{\rm IMF} = $" + f'${mu_alpha:.3f} \\pm {sigma_alpha:.3f}$' + f"\n" + r"$\log_{10} N_{\rm Ia} = $" + f'${mu_log10N_Ia:.3f} \\pm {sigma_log10N_Ia:.3f}$'
     
-    legend_true = plt.scatter(global_params[0,0], global_params[0,1], color='red', label=label_gt, s=100)
+    legend_true = plt.scatter(global_params[0,0], global_params[0,1], marker='*', color='red', label=label_gt, s=100)
     legend_fit = plt.scatter(mu_alpha, mu_log10N_Ia, color='k', label=label_fit, s=100)
     
     legend_fit = plt.legend(handles=[legend_fit], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.2))
@@ -446,11 +446,12 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
 
     # Sigma levels
     levels = []
+    colors = ['w', 'w', 'k']
     sigma = np.array([3,2,1], dtype=float)
     for n in sigma:
         levels.append(posterior.pdf([mu_alpha+n*sigma_alpha, mu_log10N_Ia+n*sigma_log10N_Ia]))
-    CS = plt.contour(x, y, posterior.pdf(pos), levels=levels, colors='k', linestyles='dashed')
-    text = plt.clabel(CS, inline=True, fontsize=15)
+    CS = plt.contour(x, y, posterior.pdf(pos), levels=levels, colors='w', linestyles='dashed')
+    text = plt.clabel(CS, inline=True, fontsize=25)
     for t in text:
         i = np.abs(np.array(levels) - float(t._text)).argmin()
         s = int(sigma[i])

### 2025-01-23 src/scripts/chempy_alt_yield_inference.py
[showyourwork] overleaf sync

@@ -1,0 +1,93 @@
+import paths
+import numpy as np
+
+from scipy.stats import norm
+from tqdm import tqdm
+
+import time as t
+import pickle
+
+from plot_functions import *
+
+# ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
+name = "NPE_C"
+
+global_params = np.array([[-2.3, -2.89], [0.3, 0.3]])
+
+# --- Clean the data ---
+# Chempy sometimes returns zeros or infinite values, which need to removed
+def clean_data(x, y):
+    # Remove all zeros from the training data
+    index = np.where((y == 0).all(axis=1))[0]
+    x = np.delete(x, index, axis=0)
+    y = np.delete(y, index, axis=0)
+
+    # Remove all infinite values from the training data
+    index = np.where(np.isfinite(y).all(axis=1))[0]
+    x = x[index]
+    y = y[index]
+
+    return x, y
+
+# ----- Add noise to data -------------------------------------------------------------------------------------------------------------------------------------------
+def add_noise(true_abundances):
+    # Define observational erorrs
+    pc_ab = 5 # percentage error in abundance
+
+    # Jitter true abundances and birth-times by these errors to create mock observational values.
+    obs_ab_errors = np.ones_like(true_abundances)*float(pc_ab)/100.
+    obs_abundances = norm.rvs(loc=true_abundances,scale=obs_ab_errors)
+
+    return obs_abundances
+
+# ----- Load the data -------------------------------------------------------------------------------------------------------------------------------------------
+# CHEMPY data with alternative yields
+data_alternative_yields = np.load(paths.data / "chempy_data/chempy_alternative_yields.npz", mmap_mode='r')
+
+elements = data_alternative_yields['elements']
+alt_x = data_alternative_yields['params']
+alt_y = data_alternative_yields['abundances']
+
+alt_x, alt_y = clean_data(alt_x, alt_y)
+
+# Remove H from data, because it is just used for normalization (output with index 2)
+elements = np.delete(elements, 2)
+alt_y = np.delete(alt_y, 2, 1)
+
+alt_y_obs = add_noise(alt_y)
+
+# ----- Load posterior -----
[…]

### 2025-01-23 src/scripts/chempy_tng_inference.py

@@ -30,7 +30,7 @@ mu_N_Ia_combined = np.sum(mu_N_Ia/sigma_N_Ia**2)/np.sum(1/sigma_N_Ia**2)
 sigma_N_Ia_combined = 1/np.sqrt(np.sum(1/sigma_N_Ia**2))
 
 with open(paths.output / f'CHEMPY_TNG_sbi.txt', 'w') as f:
-    f.write(f"$\alpha_{{\rm IMF}}={mu_alpha_combined:.3f}\pm{sigma_alpha_combined:.3f}$ and $\log_{{10}}(\rm N_{{Ia}})={mu_N_Ia_combined:.3f} \pm {sigma_N_Ia_combined:.3f}$%")
+    f.write(f"$\\alpha_{{\\rm IMF}}={mu_alpha_combined:.3f}\pm{sigma_alpha_combined:.3f}$ and $\\log_{{10}}(\\rm N_{{Ia}})={mu_N_Ia_combined:.3f} \\pm {sigma_N_Ia_combined:.3f}$%")
 
 print(f"alpha_IMF = {mu_alpha_combined:.3f} +/- {sigma_alpha_combined:.3f}")
 print(f"log10_N_Ia = {mu_N_Ia_combined:.3f} +/- {sigma_N_Ia_combined:.3f}")

### 2025-01-23 src/scripts/chempy_alt_yield_inference.py
[showyourwork] overleaf sync

@@ -57,7 +57,7 @@ alt_y = np.delete(alt_y, 2, 1)
 alt_y_obs = add_noise(alt_y)
 
 # ----- Load posterior -----
-with open(f'data/posterior_{name}.pickle', 'rb') as f:
+with open(paths.data / f'posterior_{name}.pickle', 'rb') as f:
     posterior = pickle.load(f)
 
 # ----- Evaluate the posterior -------------------------------------------------------------------------------------------------------------------------------------------

### 2025-01-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/CHEMPY_alternative_yields_Nstar_comp.pdf}
+    \vspace{-.5cm}
+    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
+For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
+    \label{fig:CHEMPY_TNG_N_star_analysis}
+    \script{chempy_alt_yield_inference.py}
+\end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/TNG_sim_Nstar_comp.pdf}
+    \vspace{-.5cm}
+    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
+For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
+    \label{fig:TNG_N_star_analysis}
+    \script{tng_inference.py}
+\end{figure*}
 
-\paragraph{Inference on the \texttt{CHEMPY} TNG yield set}
+\paragraph{Inference on mock data from \texttt{CHEMPY} with TNG yield set}
 
@@
 
-\subsection{Alternative yield set}
+\subsection{Inference on mock data from \texttt{CHEMPY} with incorrect yield set}
 \label{subsec: mocks_wrong_yield}
+
+\begin{figure}
+%\vspace{-1.3cm}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/CHEMPY_alternative_yields.png}
+    \vspace{-.25cm}
+    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.
+The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
+    \label{fig:CHEMPY_TNG_sbi} 
+    \script{chempy_alt_yield_inference.py}
+\end{figure}
+
+\subsection{Inference on mock data from the IllustrisTNG simulation}
+\label{subsec: tng_sim}
+
+\begin{figure}
+%\vspace{-1.3cm}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/TNG_simulation.png}
+    \vspace{-.25cm}
+    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.
+The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
+    \label{fig:CHEMPY_TNG_sbi} 
+    \script{tng_inference.py}
+\end{figure}
 

### 2025-01-24 src/scripts/tng_inference.py

@@ -41,42 +41,34 @@ def add_noise(true_abundances):
     return obs_abundances
 
 # ----- Load the data -------------------------------------------------------------------------------------------------------------------------------------------
-# CHEMPY data with alternative yields
-data_alternative_yields = np.load(paths.data / "chempy_data/chempy_alternative_yields.npz", mmap_mode='r')
+# TNG simulation data
+data_tng = np.load("data/chempy_data_paper/Mock_Data_TNG.npz", mmap_mode='r')
 
-elements = data_alternative_yields['elements']
-alt_x = data_alternative_yields['params']
-alt_y = data_alternative_yields['abundances']
+tng_y = data_tng['true_abuns']
 
-alt_x, alt_y = clean_data(alt_x, alt_y)
-
-# Remove H from data, because it is just used for normalization (output with index 2)
-elements = np.delete(elements, 2)
-alt_y = np.delete(alt_y, 2, 1)
-
-alt_y_obs = add_noise(alt_y)
+tng_y_obs = add_noise(tng_y)
 
 # ----- Load posterior -----
 with open(paths.data / f'posterior_{name}.pickle', 'rb') as f:
     posterior = pickle.load(f)
 
 # ----- Evaluate the posterior -------------------------------------------------------------------------------------------------------------------------------------------
-alpha_IMF_alt_obs = []
-log10_N_Ia_alt_obs = []
+alpha_IMF_tng_obs = []
+log10_N_Ia_tng_obs = []
 simulations = 1000
-N_stars = len(alt_y_obs)
+N_stars = len(tng_y_obs)
 
 start = t.time()
-for i in tqdm(range(len(alt_y_obs))):
-    x = add_noise(alt_y_obs[i])
+for i in tqdm(range(len(tng_y_obs))):
+    x = add_noise(tng_y_obs[i])
     alpha, N_Ia = posterior.sample((simulations,), x=x, show_progress_bars=False)[:,0:2].T
-    alpha_IMF_alt_obs.append(alpha)
-    log10_N_Ia_alt_obs.append(N_Ia)
+    alpha_IMF_tng_obs.append(alpha)
+    log10_N_Ia_tng_obs.append(N_Ia)
 end = t.time()
[…]

### 2025-01-24 src/scripts/tng_inference.py
[showyourwork] overleaf sync

@@ -42,7 +42,7 @@ def add_noise(true_abundances):
 
 # ----- Load the data -------------------------------------------------------------------------------------------------------------------------------------------
 # TNG simulation data
-data_tng = np.load("data/chempy_data_paper/Mock_Data_TNG.npz", mmap_mode='r')
+data_tng = np.load(paths.data / "Mock_Data_TNG.npz", mmap_mode='r')
 
 tng_y = data_tng['true_abuns']
 

### 2025-01-27 src/scripts/chempy_tng_inference.py
add ltu-ili calibration plots

@@ -38,7 +38,6 @@ print(f"log10_N_Ia = {mu_N_Ia_combined:.3f} +/- {sigma_N_Ia_combined:.3f}")
 
 # --- Plot the data ---
 gaussian_posterior_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields")
-
 #stars = np.arange(1,1000)
 #n_stars_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, "CHEMPY_TNG_yields_N_star", stars)
 

### 2025-01-27 src/scripts/evaluate_posterior.py

@@ -84,3 +84,19 @@ for i in range(ape.shape[1]):
     if i in [0,1]:
         with open(paths.output / f'posterior_APE.txt', 'a') as f:
             f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")
+
+# --- Plot calbration using ltu-ili ---
+from ili.validation.metrics import PosteriorCoverage
+
+plot_hist = ["coverage", "histogram", "predictions", "tarp"]
+metric = PosteriorCoverage(
+    num_samples=1000, sample_method='direct',
+    labels=[rf'$\log_{{10}}(M_{{s}}) [M_{{\odot}}]$', rf'$\log_{{10}}(\tau) [Gyr]$'], plot_list = plot_hist
+)
+
+fig = metric(
+    posterior=posterior,
+    x=xs, theta=thetas)
+
+for plot in plot_hist:
+    plt.savefig(paths.figures / f'ili_{plot}.pdf')

### 2025-01-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
           Berkay Günes\inst{1,2}
+          \and
+          et al.
           }
@@
   % {} leave it empty if necessary  
-   {Chemical abundances of stars are able to reveal important galactic parameters, such as the initial mass function high-mass slope and the frequency of type Ia supernovae.
+   {Galactic chemical abundances provide crucial insights into fundamental galactic parameters, such as the high-mass slope of the initial mass function (IMF) and the normalization of Type Ia supernova (SN\,Ia) rates.
+Constraining these parameters is essential for advancing our understanding of stellar feedback, metal enrichment, and galaxy formation processes.
+However, traditional Bayesian inference techniques, such as Hamiltonian Monte Carlo (HMC), are computationally prohibitive when applied to the large datasets produced by modern stellar surveys.}
+  % aims heading (mandatory)
+   {This study aims to leverage simulation-based inference (SBI) as a scalable, robust, and efficient method for constraining galactic parameters from stellar chemical abundances.
+By focusing on the high-mass IMF slope and SN\,Ia normalization, we aim to demonstrate the advantages of SBI over HMC in terms of speed, scalability, and robustness against model misspecifications.}
+  % methods heading (mandatory)
+   {We combined a Galactic Chemical Evolution (GCE) model, \texttt{CHEMPY}, with a neural network emulator and a Neural Posterior Estimator (NPE) to train an SBI pipeline.
+Mock datasets were generated using \texttt{CHEMPY}, including scenarios with mismatched nucleosynthetic yields, and additional tests were conducted on data from a Milky Way-like galaxy taken from a hydrodynamical simulation.
+SBI results were benchmarked against HMC-based inference, focusing on computational performance, accuracy, and resilience to systematic discrepancies.}
+  % results heading (mandatory)
+   {Our SBI approach achieves a $\sim1,000\times$ ($\sim72,000\times$) speed-up compared to HMC, reducing inference runtime from $\sim40$ hours for 200 stars to mere minutes for thousands of stars.
+Inference on $1,000$ stars yielded precise estimates for the IMF slope ($\alpha_{\rm IMF} = −2.296 \pm 0.009$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = −2.889 \pm 0.010$), deviating by less than 0.04\% from the ground truth.
+SBI also demonstrated greater robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation.}
+  % conclusions heading (optional), leave it empty if necessary 
+   {SBI represents a paradigm shift in galactic chemical enrichment studies, enabling efficient and precise analysis of massive stellar datasets.
+By outperforming HMC in speed, scalability, and robustness, SBI is poised to become a cornerstone methodology for future spectroscopic surveys, such as 4MOST and WEAVE, facilitating deeper insights into the chemical and dynamical evolution of galaxies.}
+
+%Chemical abundances of stars are able to reveal important galactic parameters, such as the initial mass function high-mass slope and the frequency of type Ia supernovae.
 Constraining these parameters is of critical importance in order to create realistic hydrodynamical simulations and understand galactic physics.
@@
 Here we present a new method using simulation-based inference (SBI) to circumvent previous limitations and enable inference from massive stellar surveys.
-    SBI enables a new approach to the inference of the posterior distribution of Galactic parameters under an intractable likelihood function by utilizing forward simulations of stellar abundance distributions.
-    Using SBI we are able to predict global galactic parameters of great accuracy and precision from chemical abundances of multiple stars, faster than conventional methods like Hamiltonian Monte Carlo inference.
-Our method easily scales to hundred-thousands of stars and beats previous approaches by orders of magnitude in compute time and accuracy.}
-  % aims heading (mandatory)
-   {}
-  % methods heading (mandatory)
-   {}
-  % results heading (mandatory)
-   {}
-  % conclusions heading (optional), leave it empty if necessary 
-   {}
+%SBI enables a new approach to the inference of the posterior distribution of Galactic parameters under an intractable likelihood function by utilizing forward simulations of stellar abundance distributions.
+%Using SBI we are able to predict global galactic parameters of great accuracy and precision from chemical abundances of multiple stars, faster than conventional methods like Hamiltonian Monte Carlo inference.
+Our method easily scales to hundred-thousands of stars and beats previous approaches by orders of magnitude in compute time and accuracy.
 
@@
 
-Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars or the production of heavy elements inside stars \citep[e.g.][]{Buck2021}.
-Two crucial unknowns are the shape of the initial mass function (IMF), setting the mass distribution of stars born from the interstellar medium (ISM), and the rate of Type Ia supernovae (SN\,Ia) explosions.
+Understanding the chemical enrichment of galaxies is fundamental to deciphering their formation and evolution.
+Chemical abundances of stars offer a wealth of information about galactic parameters, such as the high-mass slope of the initial mass function (IMF) and the normalization of Type Ia supernova (SN\,Ia) rates.
+These parameters critically influence the production of heavy elements \citep[e.g.][]{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, stellar feedback, and star formation histories, making their accurate determination essential for realistic hydrodynamical simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Buck2021,Font2020,Agertz2021}.
+Despite their importance, constraining these parameters has proven challenging due to limited observational data and the computational demands of traditional inference techniques.
 
-The exact values of parameters sensitively affect chemical evolution tracks \citep{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, yet their observational constraints remain weak.
-%
 For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,Rybizki2015,Chabrier2014}.
@@
 
-The goal of this work is to demonstrate how we can use modern machine learning (ML) techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars.
-Our work focuses on two key parameters; the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF and the rate of SN\,Ia explosions per unit mass.
-We assume that both are constant across the galaxy and time.
+Recent advances in stellar spectroscopic surveys, such as APOGEE and GALAH, have produced unprecedented datasets of stellar chemical abundances.
+These datasets hold the potential to unlock detailed constraints on galactic parameters across diverse environments.
+However, traditional Bayesian inference methods, such as Markov Chain Monte Carlo (MCMC) and Hamiltonian Monte Carlo (HMC), struggle to scale to these large datasets.
+Such methods are computationally expensive, requiring hours of runtime for even modest sample sizes, and are susceptible to biases when confronted with high-dimensional posterior distributions.
 
-A methodologically very similar work was previously presented by \cite{Philcox_2019} using Hamilton Monte Carlo sampling for the inference.
-The result of this work was an impressive speed-up over classical Markov Chain Monte Carlo sampling with no loss of accuracy.
-Here we make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}.
-For high-dimensional posterior functions, SBI gives tremendous computational improvements over more traditional methods such as Markov Chain Monte Carlo (MCMC) or even the more efficient Hamilton Monte Carlo (HMC) methods.
+In this work, we present a novel approach leveraging simulation-based inference \citep[SBI, e.g.][]{Cranmer2020} to address these limitations.
+SBI bypasses the need for explicit likelihood functions, enabling efficient and scalable inference of galactic parameters directly from simulated stellar abundances.
+By combining a neural network emulator for the \texttt{CHEMPY} Galactic Chemical Evolution (GCE) model with a Neural Posterior Estimator (NPE), we achieve rapid and robust inference.
+Unlike HMC, which requires extensive sampling for each dataset, our method amortizes the computational cost during training, allowing subsequent inference to scale seamlessly to larger datasets.
 
-This paper is structured as follows.
-In Section~\ref{sec:methods} we detail our methods.
-In particular, we discuss the GCE model and the methods behind SBI.
-Our results are presented in Section~\ref{sec: Results} where we test the performance of our approach on mock data and data taken from a hydrodynamical simulation.
-We conclude our study with Section~\ref{sec: conclusion} with a summary and outlook of our results.
+This study focuses on two critical global galactic parameters: 
+the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF ($\alpha_{\rm IMF}$) and the SN\,Ia normalization ($\log_{10}(N_{\rm Ia})$), the rate of SN\,Ia explosions per unit mass.
+We demonstrate the accuracy, scalability, and robustness of SBI through tests on mock datasets generated by \texttt{CHEMPY} \citep{Rybizki_2017}, as well as on data from hydrodynamical simulations.
+Additionally, we compare our results to those obtained using HMC-based inference on the same datasets \citep[see][]{Philcox_2019}, highlighting SBI's superior performance in terms of speed, precision, and resilience to model misspecification.
+
+The structure of this paper is as follows: In Section~\ref{sec:methods}, we outline the methods used, including the GCE model and SBI framework.
+Section~\ref{sec: Results} presents our results on both \texttt{CHEMPY} and IllustrisTNG \citep{Pillepich2018} mock data, emphasizing SBI's advantages over traditional approaches.
+Finally, in Section~\ref{sec: discussion}, we discuss the broader implications of our findings and outline potential future applications, before concluding in Section~\ref{sec: conclusion}.
 
@@
 We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
-To test our method, we aim further at inferring parameters from a sample of stars taken from a hydrodynamical simulation of a MW type galaxy which we take from the Ilustris TNG project \citep{Pillepich2018}.
+To test our method, we aim further at inferring parameters from a sample of stars taken from a hydrodynamical simulation of a MW type galaxy which we take from the IlustrisTNG project \citep{Pillepich2018}.
 To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
@@
 \texttt{Chempy} uses only net yields, such that they provide only newly synthesized material, with the remainder coming from the initial SSP composition.
-These tables may not well-represent true stellar chemistry, and the effects of this missmatch are examined in Sec.\,\ref{subsec: mocks_wrong_yield} by performing inference using an alternative set of yields that does not match the yield set of the training data.
+These tables may not well-represent true stellar chemistry, and the effects of this mismatch are examined in Sec.\,\ref{subsec:mocks_wrong_yield} by performing inference using an alternative set of yields that does not match the yield set of the training data.
 For the analysis of observational data, we would want to use the most up-to-date yields, such as \citet{2016ApJ...825...26K} AGB yields, and carefully chose elements which are known to be well reproduced by our current models (e.g.
 shown by \citet{2019ApJ...874..102W,2019arXiv190806113G}), though this is not appropriate in our context.
-To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{1998A&A...334..505P} and do not allow for any `hypernovae' \citep[in contrary to][]{2018ApJ...861...40P}).
+To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{portinari} and do not allow for any `hypernovae' \citep[in contrary to][]{2018ApJ...861...40P}).
 
@@
 We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator.
-In order to mimick observational noise, we add 5\% observational uncertainties to the abundances simulated with \texttt{CHEMPY} before feeding them to our NPE.
+In order to mimic observational noise, we add 5\% observational uncertainties to the abundances simulated with \texttt{CHEMPY} before feeding them to our NPE.
 We have made sure that our NPE is well calibrated and posterior distributions are trustable. For more details on simulation-based calibration see appendix \ref{sec:sbc}.
@@
      \caption{Absolute percentage error of the neural posterior density estimate for a single star.
-Different coloured histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines.
-The box plots show again the first and third quartiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box.
+Different colored histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines.
+The box plots show again the first and third quantiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box.
 The global parameters of main interest for this work are shown by the light blue and red histogram.}
@@
 
-For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with 1000 points.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
+    \vspace{-.5cm}
+    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
+For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainties as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).
+See Sec.~\ref{subsec:chempy_tng}.}
+    \label{fig:CHEMPY_TNG_N_star_analysis}
+    \script{chempy_tng_inference.py}
+\end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/CHEMPY_alternative_yields_Nstar_comp.pdf}
+    \vspace{-.5cm}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for the mock data created with a different yield set than the training data. See Sec.~\ref{subsec:mocks_wrong_yield}.}
+    \label{fig:CHEMPY_alt_N_star_analysis}
+    \script{chempy_alt_yield_inference.py}
+\end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/TNG_sim_Nstar_comp.pdf}
+    \vspace{-.5cm}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for the mock data taken from an IlustrisTNG Milky Way-like galaxy. See Sec.~\ref{subsec:tng_sim}.}
+    \label{fig:TNG_N_star_analysis}
+    \script{tng_inference.py}
+\end{figure*}
+
+For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with 1,000 points.
 This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
-Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1000 stars (each time sampling the single star posterior with 1000 points and fitting for the Gaussian parameters).
-Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}.
-But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim72.000$.
+Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
+Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1,000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}.
+But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim72,000$.
 \textcolor{red}{check if those speed-up values are correctly calculated...}
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
-    \vspace{-.5cm}
-    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
-For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
-    \label{fig:CHEMPY_TNG_N_star_analysis}
-    \script{chempy_tng_inference.py}
-\end{figure*}
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figures/CHEMPY_alternative_yields_Nstar_comp.pdf}
-    \vspace{-.5cm}
-    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
-For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
-    \label{fig:CHEMPY_TNG_N_star_analysis}
-    \script{chempy_alt_yield_inference.py}
-\end{figure*}
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figures/TNG_sim_Nstar_comp.pdf}
-    \vspace{-.5cm}
-    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
-For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
-    \label{fig:TNG_N_star_analysis}
-    \script{tng_inference.py}
-\end{figure*}
-
-\paragraph{Inference on mock data from \texttt{CHEMPY} with TNG yield set}
+\subsection{Inference on mock data from \texttt{CHEMPY} with TNG yield set}
+\label{subsec:chempy_tng}
 
@@
     \vspace{-.25cm}
-    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.
+    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1,000$ stars.
 The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
@@
 In Fig.~\ref{fig:CHEMPY_TNG_sbi} we show the joint posterior for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ for our inference.
-The red star indicates the ground truth value, the black dot shows our posterior mean and white contours indicate $1-3\sigma levels$.
-
-Using a sample of 1000 stars we infer \variable{output/CHEMPY_TNG_sbi.txt} which deviates less than $\sim0.04\%$ from the correct value.
+The red star indicates the ground truth value, the black dot shows our posterior mean and white contours indicate $1-3\sigma$ levels.
+%
+Using a sample of 1,000 stars we infer \variable{output/CHEMPY_TNG_sbi.txt} which deviates less than $\sim0.04\%$ from the correct value.
 
@@
 
-\subsection{Inference on mock data from \texttt{CHEMPY} with incorrect yield set}
-\label{subsec: mocks_wrong_yield}
+\subsection{Inference on mock \texttt{CHEMPY} data with incorrect yield set}
+\label{subsec:mocks_wrong_yield}
 
@@
     \vspace{-.25cm}
-    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.
-The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
-    \label{fig:CHEMPY_TNG_sbi} 
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_sbi} but for the mock data created with a different yield set than the training data. See Sec.~\ref{subsec:mocks_wrong_yield}.}
+    \label{fig:CHEMPY_alt_sbi} 
     \script{chempy_alt_yield_inference.py}
@@
 
-\subsection{Inference on mock data from the IllustrisTNG simulation}
-\label{subsec: tng_sim}
+There is quite some discussion in the literature about stellar nucleosynthesis with various different yield sets proposed \citep[see e.g.
+discussion in][]{Rybizki_2017}.
+In fact, any tabulated yield set will differ from reality and during an application of our inference pipeline it will not be clear which tabulated yield set most closely matches reality and should be used.
+In order to investigate how sensitive our method is to a model missspecification by using an incorrect yield set, we create another set of mock data using \texttt{CHEMPY} with a different yield set than during training of our NPE.
+For better cross-comparison we decided to use the same alternative yield set as presented in Tab.~5 of \citep{Philcox_2019}.
+In short this yield set uses SN Ia yields from \citet{2003NuPhA.718..139T}, SN II yields from \citet{2013ARA&A..51..457N} and AGB yields from \citet{2016ApJ...825...26K}.
+By chosing this set of yields we have made sure that contributions to all three processes differ by $\mathcal{O}(10\%)$ \citep[see Sec.~6.2 of][for more details]{Philcox_2019}.
 
@@
     \vspace{-.25cm}
-    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars.
-The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
-    \label{fig:CHEMPY_TNG_sbi} 
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_sim_sbi} but for the mock data taken from an IlustrisTNG Milky Way-like galaxy. See Sec.~\ref{subsec:tng_sim}.}
+    \label{fig:CHEMPY_TNG_sim_sbi} 
     \script{tng_inference.py}
@@
 
+Our mock data generation and inference then follows the one of Sec.~\ref{subsec:chempy_tng}.
+Note, this means we apply our NPE trained on data with the TNG yield set to data created with the above mentioned alternative yield set.
+This effectively probes the effect of model missspecification on the inference results.
+
+We show our inference results for this setup for varying number of stars (blue) in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis} comparing again against HMC results (red) from \citet{Philcox_2019}.
+Note that the SBI results are less biased than the HMC results especially for $\alpha_{\rm IMF}$.
+In fact with 200 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively.
+Again, the joint posterior is shown in Fig.~\ref{fig:CHEMPY_alt_sbi} and shows that while individually parameter inferences look good, jointly taken the SBI results are on the edge of being $3\sigma$ away from the ground truth.
+
+Nevertheless, the performance of our inference is still very good.
+We see that an increasing number of observations helps to decrease the models uncertainty just as before.
+However, we also note that our inference is now slightly biased as the observational data does not match the training data.
+Looking at Fig.~\ref{fig:CHEMPY_alt_sbi} we see that our inference is still consistent with the ground truth within the $3\sigma$ levels.
+Comparing this to the HMC results, we see that SBI is performing much better when the model is misspecified and leads to significantly less biased results.
+This is next to drastically reduced compute times another a key advantage of our SBI method compared to more standard approaches such as HMC.
+
+In an accompanying paper we more closely investigate measures of model misspecification and inference of best fitting models next to just parameter inference.
+
+\subsection{Inference on mock data from the IllustrisTNG simulation}
+\label{subsec:tng_sim}
+
+As a galactic chemical evolution code, \texttt{CHEMPY} is a one-zone model with simplified ISM physics that only approximately describes star formation and feedback as well as metal mixing in the ISM.
+In the parametrization of \texttt{CHEMPY} as used here, we can assign each star to its own ISM environment, but we cannot exchange gas between environments and do not model sudden star formation or infall events.
+Hence, this section is dedicated to investigate whether this significantly biases our inference of the SSP parameters (noting that results from \citet{2019ApJ...874..102W} justify the treatment of ISM parameters as latent variables).
+
+In order to explore what effect this simplified treatment of ISM physics has on the inference we now turn to a more complex model of the formation and chemical enrichment of a Milky Way-type galaxy taken from the IllustrisTNG simulations.
+Note, that by now also the NIHAO simulations \citep{Wang2015, Buck2020,Buck2020c} have implemented \texttt{CHEMPY} supported yield tables including the TNG yield set \citep{Buck2021} and hence would make for a nice dataset for our analysis.
+However, we have decided to use the exact same galaxy as in \citet{Philcox_2019} for better comparison of our results.
+
+In detail, we use a single galaxy from the $z = 0$ snapshot of the highest-resolution TNG100-1 simulation.
+We choose a subhalo (index 523071) with mass close to $10^{12}\,\mathrm{M}_\odot$ to select a Milky Way-like galaxy.
+From this, we extract a set of 1,000 random `stellar particles' from a total of $\sim 40,000$.
+Each star particle has a mass of $\sim 1.4\times 10^6 \mathrm{M}_\odot$ \citep{2019ComAC...6....2N}.
+These act as proxies for stellar environments, giving the elemental mass fractions, $\{d_i^j\}$, and cosmological scale factor, $a_i$, at the time of stellar birth.
+Mass fractions are converted to [X/Fe] abundance ratios using \citet{2009ARA&A..47..481A} solar abundances as in \textit{Chempy}, with the scale-factor ($a_i$) to birth-time ($T_i$) conversion performed using \texttt{astropy} \citep{astropy:2013,astropy:2018},\footnote{\href{http://www.astropy.org}{http://www.astropy.org}} assuming a $\Lambda$CDM cosmology with \citet{planck2015} parameters, as in TNG \citep{2018MNRAS.475..648P}.\footnote{Note, as for the \textit{Chempy} mock data, we exclude any particles with $T_i\notin[2,12.8]\,\mathrm{Gyr}$ to ensure that the true times are well separated from our training age limits, avoiding neural network errors.
+This removes $\sim5\%$ of the stars.} Observational errors are incorporated as above, giving a full data-set that is identical in structure to the \texttt{Chempy} mock data.
+For more details and a plot of the [Mg/Fe] vs.
+[Fe/H] plane for this galaxy see Sec.~6.3 and Fig.~5 of \citet{Philcox_2019}.
+
+We note that this TNG galaxy was deliberately chosen by \citet{Philcox_2019} to have both a high-$\alpha$ and low-$\alpha$ chemical evolution sequence to test their inference on a mock galaxy with Milky Way-like properties.
+There is still some debate on the exact formation of this bimodality but it is generally attributed to gas-rich mergers and different modes of star formation \citep[e.g.][]{2018MNRAS.474.3629G,2018MNRAS.477.5072M,2019MNRAS.484.3476C,Buck2020,Buck2023}.
+Similarly, in chemo-dynamical models, Milky Way-like bimodalities can also be achieved by a combination of radial migration and selection effects without the need for mergers or starbursts \citep[e.g.][]{2009MNRAS.396..203S,2013A&A...558A...9M,2017ApJ...835..224A}.
+
+We show our inference results for the TNG data set in Fig.~\ref{fig:TNG_N_star_analysis} and Fig.~\ref{fig:CHEMPY_TNG_sim_sbi}, respectively.
+Again, we find that inference becomes better with increasing number of stars.
+Despite the drastic difference in chemical enrichment model between training and testing data, our SBI pipeline is impressively capable of inferring the correct posterior values.
+SBI results for $\alpha_{\rm IMF}$ are almost perfect while $\log_{10}(N_{\rm Ia})$ is slightly biased low.
+This is exactly opposite to the results of HMC where inference for $\alpha_{\rm IMF}$ is biased and results for $\log_{10}(N_{\rm Ia})$ are more in agreement.
+Hence, our SBI inference is on par with the HMC results and given their extreme computational advantage they actually supersede HMC.
+
+Looking at the joint posterior in Fig.~\ref{fig:CHEMPY_TNG_sbi} we see that also in this case SBI recovers the true parameters with less than $3\sigma$ bias.
+
 \section{Discussion}
-\label{sec: conclusion}
+\label{sec: discussion}
 
-\textcolor{red}{here goes the assumption that stars are iid since each stars is produced by a single call to chempy.
-However, real physics might correlate stars abundacnes by star formation histories and cumulative enrichment...}
+\textcolor{red}{here goes the assumption that stars are iid since each star is produced by a single call to chempy.
+However, real physics might correlate stars abundances by star formation histories and cumulative enrichment...}
+
+Our study demonstrates that simulation-based inference (SBI) provides a powerful and efficient alternative to conventional methods such as Hamiltonian Monte Carlo (HMC) for inferring global galactic parameters.
+By leveraging neural density estimators and neural network emulators, we achieved remarkable computational efficiency without compromising precision or accuracy.
+
+The key assumption underpinning our methodology — that individual stars are independently sampled from their respective stellar environments — is foundational for tractability.
+However, this assumption warrants closer examination.
+In reality, stellar abundances are correlated due to shared star formation histories, cumulative enrichment, and dynamical interactions.
+Future efforts should explore methodologies capable of incorporating such correlations to further refine the accuracy of SBI, potentially through hierarchical modeling or graph-based methods.
+
+An exciting avenue for future research involves expanding SBI to directly infer empirical nucleosynthetic yields, which remain a major uncertainty in galactic chemical evolution models.
+In the current framework, CHEMPY relies on tabulated yields from theoretical studies, which may not fully represent the complex processes driving stellar enrichment.
+Adapting SBI to simultaneously infer galactic parameters and refine empirical yield tables would require changes to the simulator.
+Specifically, CHEMPY would need to incorporate parameterized yield modifications as part of its input space, allowing for flexible adjustments to enrichment rates during inference.
+This would also necessitate larger training datasets and enhanced validation techniques to ensure convergence.
+Such an approach could provide a unified framework for calibrating galactic models directly against observational data.
+
+Compared to previous HMC-based studies \citep[e.g.][]{Philcox_2019}, our results highlight SBI's resilience to certain types of model misspecification, such as mismatched yield tables.
+The robustness of SBI in these cases stems from its ability to approximate complex posterior distributions efficiently.
+Notably, SBI retained high accuracy in scenarios where HMC struggled, particularly for the high-mass slope of the IMF ($\alpha_{\rm ISM}$).
+This suggests SBI's potential for real-world applications, where the underlying models may deviate from observational data.
+
+The applicability of our method extends to future spectroscopic surveys, such as those planned by the 4MOST or WEAVE consortia, which will provide orders of magnitude more data than current datasets.
+Our findings indicate that SBI can seamlessly scale to such large datasets, offering significant advantages in terms of speed and computational cost.
+
+Nevertheless, limitations remain.
+The simplified physics of the \texttt{CHEMPY} model, while advantageous for computational efficiency, omits the complexities of dynamical gas flows, feedback, and metal mixing present in cosmological simulations.
+While our tests on IllustrisTNG data affirm the robustness of SBI, integrating more sophisticated models into the inference pipeline represents an exciting avenue for future work.
+On this line of reasoning we refer also to the discussion in \citet{Philcox_2019}.
 
@@
 
-Here we have explored innovative SBI methods to constrain important galactic parameters from the multi-dimensional observation of chemical abundances of stars.
-Our results show that SBI provides a super fast method, that beats previous inference algorithms by orders of magnitude, and delivers accurate and precise results.
-Hence, SBI enables faster and more accurate evaluation of observational data after a small initial computational investment of training a NPE ($\sim3$ h).
-Since evaluation of the NPE is only weakly dependent on the dimensionality of the input observations we expect almost perfect scaling to larger chemical abundances observations and even further improvements in the inference results with future high quality data from massive spectroscopic surveys.
+This study introduces simulation-based inference (SBI) as an innovative framework for constraining galactic parameters using stellar chemical abundances.
+By training neural posterior estimators on forward simulations from the \texttt{CHEMPY} model, we achieved precise and accurate inferences for two key parameters: the high-mass slope of the initial mass function ($\alpha_{\rm IMF}$) and the normalization of Type Ia supernova rates ($\log_{10}(N_{\rm Ia})$).
+
+Our results underscore the transformative advantages of SBI over traditional methods like Hamiltonian Monte Carlo (HMC), marking a paradigm shift in galactic parameter inference:
+\begin{itemize}
+    \item Orders-of-magnitude speed-up: SBI dramatically reduces computational requirements, achieving runtime improvements exceeding 1,000-fold compared to HMC.
+For instance, while HMC requires ∼40 hours to infer parameters from just 200 stars, SBI completes inference on thousands of stars in mere minutes.
+This efficiency makes SBI uniquely suited for analyzing the massive datasets expected from next-generation spectroscopic surveys.
+    \item Scalability: SBI’s amortized nature allows it to scale seamlessly with the size of the dataset.
+By training a neural posterior estimator once, the method can be applied repeatedly at virtually no additional computational cost.
+This scalability is essential for leveraging the millions of stars that future surveys like 4MOST and WEAVE will provide, enabling precise population-level inferences.
+    \item Robustness to model misspecifications: Unlike HMC, which shows significant biases when faced with discrepancies between model assumptions and data, SBI demonstrates remarkable robustness.
+Even under conditions of mismatched yield tables or data generated from hydrodynamical simulations, SBI provides accurate and reliable results.
+This robustness ensures SBI’s applicability in real-world scenarios where exact model fidelity cannot be guaranteed.
+\end{itemize}
+
+In addition to its immediate advantages, SBI lays the foundation for future advancements in galactic modeling.
+Its flexibility can enable the direct inference of empirical nucleosynthetic yields and facilitate integration with more complex galaxy formation models.
+These enhancements will further solidify SBI as a cornerstone method in galactic archaeology.
+
+In conclusion, SBI represents a breakthrough in simulation-based analysis, delivering unparalleled speed, precision, and scalability.
+By overcoming the computational limitations of traditional techniques like HMC, SBI paves the way for extracting deeper insights into the chemical and dynamical evolution of galaxies in the era of massive spectroscopic surveys.
 

### 2025-01-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
    \institute{Interdisciplinary Center for Scientific Computing (IWR), University of Heidelberg,
- Im Neuenheimer Feld 205, D-69120 Heidelberg
+ Im Neuenheimer Feld 205, D-69120 Heidelberg, Germany
  \and
@@
    {Our SBI approach achieves a $\sim1,000\times$ ($\sim72,000\times$) speed-up compared to HMC, reducing inference runtime from $\sim40$ hours for 200 stars to mere minutes for thousands of stars.
-Inference on $1,000$ stars yielded precise estimates for the IMF slope ($\alpha_{\rm IMF} = −2.296 \pm 0.009$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = −2.889 \pm 0.010$), deviating by less than 0.04\% from the ground truth.
+Inference on $1,000$ stars yielded precise estimates for the IMF slope ($\alpha_{\rm IMF} = -2.296 \pm 0.009$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = -2.889 \pm 0.010$), deviating by less than 0.04\% from the ground truth.
 SBI also demonstrated greater robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation.}
@@
 
-In order to establish our new method based on simulation based inference we need two ingredients: A simulator (in our case a Galactic chemical evolution model) that simulates observational data from a set of model parameters (in our case the IMF slope and the Type Ia supernovae normalization) and a flexible way of parametrizing the posterior density conditioned on the observation in order to perform our inference.
+In order to establish our new method based on SBI we need two ingredients: A simulator (in our case a GCE model) that simulates observational data from a set of model parameters (in our case the IMF slope and the Type Ia supernovae normalization) and a flexible way of parametrizing the posterior density conditioned on the observation in order to perform our inference.
 In the next subsections we describe both ingredients in detail.
@@
 Our simulator is based on the \texttt{CHEMPY} model \citep{Rybizki_2017}.
-\texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics.
+\texttt{CHEMPY} is a simple GCE model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics.
 We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.
@@
      \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
-As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, which controls the peak of the star formation rate (SFR), and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$.
+As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, the peak of the star formation rate (SFR), $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$.
 We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
+     %
      \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
@@
 \begin{center}
-\caption{Free \texttt{Chempy} parameters for each star, with their prior values and Gaussian widths.
+\caption{Free \texttt{CHEMPY} parameters for each star, with their prior values and Gaussian widths.
 Stellar birth-times are set for each star individually from a Uniform prior, based on realistic age estimates.}
@@
 We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
-To test our method, we aim further at inferring parameters from a sample of stars taken from a hydrodynamical simulation of a MW type galaxy which we take from the IlustrisTNG project \citep{Pillepich2018}.
-To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars.
+To test our method, we aim further at inferring parameters from a sample of stars taken from a hydrodynamical simulation of a MW type galaxy which we take from the IllustrisTNG project \citep{Pillepich2018}.
+To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{CHEMPY}, for enrichment by SN\,Ia, SN\,II and AGB stars.
 The utilized yields are summarized in Tab.\,\ref{tab:chempy_TNG_yields}, matching \citet[Tab.\,2]{2018MNRAS.473.4077P}, and we note that the SN\,II yields are renormalized such that the IMF-weighted yield ratios at each metallicity are equal to those from the \citet{2006ApJ...653.1145K} mass range models alone.
-\texttt{Chempy} uses only net yields, such that they provide only newly synthesized material, with the remainder coming from the initial SSP composition.
+\texttt{CHEMPY} uses only net yields, such that they provide only newly synthesized material, with the remainder coming from the initial SSP composition.
 These tables may not well-represent true stellar chemistry, and the effects of this mismatch are examined in Sec.\,\ref{subsec:mocks_wrong_yield} by performing inference using an alternative set of yields that does not match the yield set of the training data.
-For the analysis of observational data, we would want to use the most up-to-date yields, such as \citet{2016ApJ...825...26K} AGB yields, and carefully chose elements which are known to be well reproduced by our current models (e.g.
+For the analysis of observational data, we would want to use the most up-to-date yields, such as \citet{2016ApJ...825...26K} AGB yields, and carefully choose elements which are known to be well reproduced by our current models (e.g.
 shown by \citet{2019ApJ...874..102W,2019arXiv190806113G}), though this is not appropriate in our context.
@@
 Here $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
-This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}.
+This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{CHEMPY}\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe, as in \citep{2019ApJ...874..102W}, since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}.
 
-With these modifications, \texttt{Chempy} allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
-It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{Chempy} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
+With these modifications, \texttt{CHEMPY} allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
+It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{CHEMPY} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
 However,  such a test is necessary to prepare for an inference on real data.
 
-\subsection{Neural network emulator for \texttt{CHEMPY}:}
+\subsection{Neural network emulator for \texttt{CHEMPY}}
 Despite the simplifications made by the GCE model \texttt{CHEMPY}, the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space incurs some difficulties when sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$.
@@
 \begin{enumerate}
-    \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow generation of training data for SBI.
-With the NN emulator, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{Chempy}.
+    \item \textbf{Speed:} The run-time of the \texttt{CHEMPY} function is $\sim1$\,s per input parameter set, which leads to very slow generation of training data for SBI.
+With the NN emulator, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{CHEMPY}.
     \item \textbf{Differentiability:} The NN is written in pytorch which allows for automatic differentiation.
-Additionally the NN has a simple closed-form analytic structure \citep[described in the appendix of][]{Philcox_2019}, unlike the complex \texttt{Chempy} model.
+Additionally the NN has a simple closed-form analytic structure \citep[described in the appendix of][]{Philcox_2019}, unlike the complex \texttt{CHEMPY} model.
 This allows it to be differentiated, so one can use it to sample via advanced methods such HMC as done in \citep{Philcox_2019}.
@@
 Despite the additional complexity introduced by using multiple stellar data-points, our NN simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
-The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation.
-With the above network parameter choices, the NN predicts abundances with an absolute percantage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
+The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{CHEMPY} to a single matrix computation.
+With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
 Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.
@@
 In order to determine the optimal values of the global galactic parameters ($\Vec{\Lambda}$) one has to sample the posterior of Eq.~\ref{eq:posterior}.
-This is in practices a fairly costly computation and even in the case of advanced techniques such as Hamilton Monte Carlo sampling the posterior can only be evaluated for a small set of stars ($\lesssim200$) using fairly long compute times \citep[$\sim40$ hours][]{Philcox_2019}.
+In practice this is a costly computation, since even with advanced techniques such as HMC sampling the posterior can only be evaluated for a small set of stars ($\lesssim200$) and requires long compute times \citep[$\sim40$ hours][]{Philcox_2019}.
 %
-However, recent advances in implicit-likelihood inference or simulation-based inference \citep{Cranmer2020} offer another very efficient approach to approximate the posterior (see next paragraph for more details).
+However, recent advances in implicit-likelihood inference or SBI \citep{Cranmer2020} offer another very efficient approach to approximate the posterior (see next paragraph for more details).
 These methods train a neural conditional density estimator to represent the conditional posterior, $\mathbb{P}(\vec\Lambda|\{D_i\})$, which can be very efficiently evaluated given observational data $\{D_i\}$.
@@
 \\\nonumber
-&=& \mathbb{P}(\vec\Lambda)\prod_{j=1}^{n_\mathrm{star}}\mathbb{P}(D_j|\vec\Lambda) 
+&=& \mathbb{P}(\vec\Lambda)\prod_{j=1}^{n_\mathrm{star}}\mathbb{P}(D_j|\vec\Lambda) \quad\quad\quad\quad\quad\quad\quad\,\,\,\, \text{(i.i.d.)}
 \\\nonumber
@@
 
-This factorization of the joint posterior into single star posteriors $\mathbb{P}(\vec\Lambda|D_j)$ has the advantage that we do not need to specify the exact number of observational data points before.
+This factorization of the joint posterior into single star posteriors $\mathbb{P}(\vec\Lambda|D_j)$ has the advantage that we do not need to specify the exact number of observational data points beforehand.
 We can simply train our neural density estimator on single star observations and then at evaluation time, we simply combine as many posterior estimates as we have observational data.
@@
 \paragraph{Simulation-based inference}
-In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a Bayesian inference framework -- works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\Vec{\theta}$ (in our case a galactic chemical evolution model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\Vec{\theta}|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\Vec{\theta}$ that reproduce the simulated observations $\Vec{X}$.
+In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a Bayesian inference framework -- works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\Vec{\theta}$ (in our case a GCE model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\Vec{\theta}|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\Vec{\theta}$ that reproduce the simulated observations $\Vec{X}$.
 Once this mapping is trained, we can apply it to observations of stellar abundances $\Vec{X_R}$ to infer $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$) (see Fig.
@@
 
-We train our NPE with $10^5$ data points consisting of $n_{\rm elemets}=8$ simulated with the NN emulator described in the previous section with inputs sampled from the priors shown in Tab.
+We train our NPE with $10^5$ data points consisting of $n_{\rm elements}=8$ simulated with the NN emulator described in the previous section with inputs sampled from the priors shown in Tab.
 \ref{tab:priors}.
-Training takes $\lesssim10$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPUs.
+Training takes $\lesssim10$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPU cores.
 We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator.
@@
 
-Note however, there is one issue here.
-Since we only have access to samples from the posterior and not the posterior itself it is difficult to evaluate equation~\ref{eq:posterior}.
+This does however present one issue, that since we only have access to samples from the posterior and not the posterior itself it is difficult to evaluate equation~\ref{eq:posterior}.
 We circumvent this by approximating each single star posterior by a Gaussian and fit for the parameters of mean and covariance.
 With this it is straight forward to evaluate equation~\ref{eq:posterior} analytically.
-In fact, under our assumption the combined posterior is a product of the prior and the product of multiple Gaussians for the individual star posteriors.
+In fact, under our assumption the combined posterior is a product between the prior and the product of multiple Gaussians for the individual star posteriors.
 The latter product is also a Gaussian with mean $\mathbf{\mu'}$ and variance $\mathbf{\sigma'}$:
@@
 
-Note, given the tiny and simple neural network that represents our NPE the above assumption of Gaussian single star posteriors and the consecutive simplification is a decent one to make.
+Given the tiny and simple neural network that represents our NPE, we note that the above assumption of Gaussianity in each of the single star posteriors not expected to notably increase our pipeline's error.
 We have further empirically verified that single star posteriors are indeed close to Gaussian.
@@
     \item Mock observations drawn from \texttt{CHEMPY} from the same yield set as the training data for the neural network emulator.
-With this we esnure to test our training strategy and the performance without any systematic distribution shifts.
+With this we ensure to test our training strategy and the performance without any systematic distribution shifts.
     \item  \texttt{CHEMPY} mock data using a different yield set to test for potential biases through model missspecification in our SBI setup.
-    \item Simulated data from a large-scale hydrodynamical simulation taken from the IlustrisTNG suite \citep{Pillepich2018} but with the same yield set as our \texttt{CHEMPY} training data to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
+    \item Simulated data from a large-scale hydrodynamical simulation taken from the IllustrisTNG suite \citep{Pillepich2018} but with the same yield set as our \texttt{CHEMPY} training data to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
 \end{enumerate}
 
-Note, the methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.
+Note, the methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{CHEMPY}.
 
 \subsection{\texttt{CHEMPY} Mock Observational Data}
-Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and varying local parameters $\vec{\Theta}_i$ using priors from Tab.
+Our analysis uses mock observations drawn from the neural network emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and varying local parameters $\vec{\Theta}_i$ using priors from Tab.
 \ref{tab:priors}.
@@
 For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainties as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).
-See Sec.~\ref{subsec:chempy_tng}.}
+See Sec.~\ref{subsec:chempy_tng} for a full description.}
     \label{fig:CHEMPY_TNG_N_star_analysis}
@@
     \vspace{-.5cm}
-    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for the mock data created with a different yield set than the training data. See Sec.~\ref{subsec:mocks_wrong_yield}.}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for the mock data created with a different yield set than the training data.
+See Sec.~\ref{subsec:mocks_wrong_yield} for a full description.}
     \label{fig:CHEMPY_alt_N_star_analysis}
@@
     \vspace{-.5cm}
-    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for the mock data taken from an IlustrisTNG Milky Way-like galaxy. See Sec.~\ref{subsec:tng_sim}.}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for the mock data taken from an IllustrisTNG Milky Way-like galaxy. See Sec.~\ref{subsec:tng_sim} for a full description.}
     \label{fig:TNG_N_star_analysis}
@@
 Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
-Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1,000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}.
+Our combined runtime for the NPE training plus sampling is then 610 seconds for 1,000 stars hence our methods making it more than $1,180$ times faster than current HMC methods \citep[cf.][who need $40$h for only 200 stars]{Philcox_2019}.
 But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim72,000$.
@@
 For a single star observation, our NPE has an APE of \variable{output/global_posterior_APE.txt} when looking at all 6 parameters of interest.
-If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}.
+If we restrict ourselves to only the global two parameters, $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$, our NPE achieves an APE of \variable{output/posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}.
 Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high.
@@
     \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1,000$ stars.
-The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
+The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.
+See Sec.~\ref{subsec:chempy_tng} for a full description.}
     \label{fig:CHEMPY_TNG_sbi} 
@@
 But in the limit of large numbers of stars (few hundred) the accuracy and precision is superior compared to HMC.
-In particular, after using a few ten stars, the NPE estimate is less biased than the HMC results.
+In particular, after using a few tens of stars, the NPE estimate is less biased than the HMC results.
 Finally, given our computational advantage we will be able to use orders of magnitude more stars for our inference.
-This is particular important since sample variances play an important role when using small samples of stars as also noted in \citet{Philcox_2019}.
+This is particularly important since sample variances play a large role when using small samples of stars as also noted in \citet{Philcox_2019}.
 %
@@
 
-Note, as shown in Fig.~\ref{fig:posterior_APE} in principle our analysis is also able to infer the local parameters $\Vec{\Theta}_i$ and $T_i$.
+Note that our analysis is in principle also able to infer the local parameters $\Vec{\Theta}_i$ and $T_i$.
 This would allow us to estimate/infer stellar ages as well.
 However, currently our NPE network is not particularly good at estimating ages from abundances alone.
-On avergae our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during mock up of our data.
+On average our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data.
 
 In summary, our SBI pipeline is quite capable of correctly and precisely inferring  global parameters of chemical enrichment models from stellar abundance alone when using the same physical model and yield tables as during training.
-Next, we will check what happens if for some reason the training data was generated with a different yield set.
+Next, we will check what happens if the training data is generated with a different yield set to that use at inference time.
 
@@
     \vspace{-.25cm}
-    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_sbi} but for the mock data created with a different yield set than the training data. See Sec.~\ref{subsec:mocks_wrong_yield}.}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_sbi} but for the mock data created with a different yield set than the training data. See Sec.~\ref{subsec:mocks_wrong_yield} for a full description.}
     \label{fig:CHEMPY_alt_sbi} 
@@
 
-There is quite some discussion in the literature about stellar nucleosynthesis with various different yield sets proposed \citep[see e.g.
+There is an extensive discussion in the literature about stellar nucleosynthesis with various different yield sets proposed \citep[see e.g.
 discussion in][]{Rybizki_2017}.
-In fact, any tabulated yield set will differ from reality and during an application of our inference pipeline it will not be clear which tabulated yield set most closely matches reality and should be used.
+In fact, any tabulated yield set will differ from reality and during an application of our inference pipeline it will not be clear which tabulated yield set most closely matches reality and hence which should be used.
 In order to investigate how sensitive our method is to a model missspecification by using an incorrect yield set, we create another set of mock data using \texttt{CHEMPY} with a different yield set than during training of our NPE.
@@
 In short this yield set uses SN Ia yields from \citet{2003NuPhA.718..139T}, SN II yields from \citet{2013ARA&A..51..457N} and AGB yields from \citet{2016ApJ...825...26K}.
-By chosing this set of yields we have made sure that contributions to all three processes differ by $\mathcal{O}(10\%)$ \citep[see Sec.~6.2 of][for more details]{Philcox_2019}.
+By choosing this set of yields we have made sure that contributions to all three processes differ by $\mathcal{O}(10\%)$ \citep[see Sec.~6.2 of][for more details]{Philcox_2019}.
 
@@
     \vspace{-.25cm}
-    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_sim_sbi} but for the mock data taken from an IlustrisTNG Milky Way-like galaxy. See Sec.~\ref{subsec:tng_sim}.}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_sim_sbi} but for the mock data taken from an IllustrisTNG Milky Way-like galaxy. See Sec.~\ref{subsec:tng_sim} for a full description.}
     \label{fig:CHEMPY_TNG_sim_sbi} 
@@
 Our mock data generation and inference then follows the one of Sec.~\ref{subsec:chempy_tng}.
-Note, this means we apply our NPE trained on data with the TNG yield set to data created with the above mentioned alternative yield set.
+This means we apply our NPE trained on data with the TNG yield set to data created with the above mentioned alternative yield set.
 This effectively probes the effect of model missspecification on the inference results.
@@
 We show our inference results for this setup for varying number of stars (blue) in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis} comparing again against HMC results (red) from \citet{Philcox_2019}.
-Note that the SBI results are less biased than the HMC results especially for $\alpha_{\rm IMF}$.
+We see that the SBI results are less biased than the HMC results especially for $\alpha_{\rm IMF}$.
 In fact with 200 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively.
@@
 Comparing this to the HMC results, we see that SBI is performing much better when the model is misspecified and leads to significantly less biased results.
-This is next to drastically reduced compute times another a key advantage of our SBI method compared to more standard approaches such as HMC.
+This is next to drastically reduced compute times -- another a key advantage of our SBI method compared to more standard approaches such as HMC.
 
@@
 
-As a galactic chemical evolution code, \texttt{CHEMPY} is a one-zone model with simplified ISM physics that only approximately describes star formation and feedback as well as metal mixing in the ISM.
+As a GCE code, \texttt{CHEMPY} is a one-zone model with simplified ISM physics that only approximately describes star formation and feedback as well as metal mixing in the ISM.
 In the parametrization of \texttt{CHEMPY} as used here, we can assign each star to its own ISM environment, but we cannot exchange gas between environments and do not model sudden star formation or infall events.
@@
 These act as proxies for stellar environments, giving the elemental mass fractions, $\{d_i^j\}$, and cosmological scale factor, $a_i$, at the time of stellar birth.
-Mass fractions are converted to [X/Fe] abundance ratios using \citet{2009ARA&A..47..481A} solar abundances as in \textit{Chempy}, with the scale-factor ($a_i$) to birth-time ($T_i$) conversion performed using \texttt{astropy} \citep{astropy:2013,astropy:2018},\footnote{\href{http://www.astropy.org}{http://www.astropy.org}} assuming a $\Lambda$CDM cosmology with \citet{planck2015} parameters, as in TNG \citep{2018MNRAS.475..648P}.\footnote{Note, as for the \textit{Chempy} mock data, we exclude any particles with $T_i\notin[2,12.8]\,\mathrm{Gyr}$ to ensure that the true times are well separated from our training age limits, avoiding neural network errors.
-This removes $\sim5\%$ of the stars.} Observational errors are incorporated as above, giving a full data-set that is identical in structure to the \texttt{Chempy} mock data.
+Mass fractions are converted to [X/Fe] abundance ratios using \citet{2009ARA&A..47..481A} solar abundances as in \texttt{CHEMPY}, with the scale-factor ($a_i$) to birth-time ($T_i$) conversion performed using \texttt{astropy} \citep{astropy:2013,astropy:2018},\footnote{\href{http://www.astropy.org}{http://www.astropy.org}} assuming a $\Lambda$CDM cosmology with \citet{planck2015} parameters, as in TNG \citep{2018MNRAS.475..648P}.\footnote{Note, as for the \texttt{CHEMPY} mock data, we exclude any particles with $T_i\notin[2,12.8]\,\mathrm{Gyr}$ to ensure that the true times are well separated from our training age limits, avoiding neural network errors.
+This removes $\sim5\%$ of the stars.} Observational errors are incorporated as above, giving a full data-set that is identical in structure to the \texttt{CHEMPY} mock data.
 For more details and a plot of the [Mg/Fe] vs.
@@
 
-\textcolor{red}{here goes the assumption that stars are iid since each star is produced by a single call to chempy.
+\textcolor{red}{here goes the assumption that stars are iid since each star is produced by a single call to \texttt{CHEMPY}.
 However, real physics might correlate stars abundances by star formation histories and cumulative enrichment...}
@@
 
-An exciting avenue for future research involves expanding SBI to directly infer empirical nucleosynthetic yields, which remain a major uncertainty in galactic chemical evolution models.
-In the current framework, CHEMPY relies on tabulated yields from theoretical studies, which may not fully represent the complex processes driving stellar enrichment.
+An exciting avenue for future research involves expanding SBI to directly infer empirical nucleosynthetic yields, which remain a major uncertainty in GCE models.
+In the current framework, \texttt{CHEMPY} relies on tabulated yields from theoretical studies, which may not fully represent the complex processes driving stellar enrichment.
 Adapting SBI to simultaneously infer galactic parameters and refine empirical yield tables would require changes to the simulator.
-Specifically, CHEMPY would need to incorporate parameterized yield modifications as part of its input space, allowing for flexible adjustments to enrichment rates during inference.
+Specifically, \texttt{CHEMPY} would need to incorporate parameterized yield modifications as part of its input space, allowing for flexible adjustments to enrichment rates during inference.
 This would also necessitate larger training datasets and enhanced validation techniques to ensure convergence.
@@
     \item Orders-of-magnitude speed-up: SBI dramatically reduces computational requirements, achieving runtime improvements exceeding 1,000-fold compared to HMC.
-For instance, while HMC requires ∼40 hours to infer parameters from just 200 stars, SBI completes inference on thousands of stars in mere minutes.
+For instance, while HMC requires $\sim 40\,\mathrm{hours}$ to infer parameters from just 200 stars, SBI completes inference on thousands of stars in mere minutes.
 This efficiency makes SBI uniquely suited for analyzing the massive datasets expected from next-generation spectroscopic surveys.
@@
 
-% The ... repository, which contains code, additional documentation, and interactive dashboards, is available on GitHub.%\footnote{URL: \url{https://github.com/bGuenes/sbi_chemical_abundances}}
-
-\section{Simulation-based calibration}
-\label{sec:sbc}
-
 \begin{figure*}[]
@@
 \end{figure}
+
+% The ... repository, which contains code, additional documentation, and interactive dashboards, is available on GitHub.%\footnote{URL: \url{https://github.com/bGuenes/sbi_chemical_abundances}}
+
+\section{Simulation-based calibration}
+\label{sec:sbc}
 

### 2025-02-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We design the NN as a simple feed-forward neural network with 2 hidden layers and 100 neurons in the first and 40 neurons in the second layer.
-Training this tiny emulator takes about 50s on the CPU.
+The NN is trained on $\sim100,000$ data points and validated on $\sim50,000$ additional data points created with \texttt{CHEMPY} using a uniform prior over the $5\sigma$-range of the original Gaussian prior stated in table~\ref{tab:priors}.
+The batch size is set to 64 and the learning rate is set to 0.001.
+We train for 20 epochs using a schedule free optimizer \citep{schedulefree}.
+Training this tiny emulator takes about 200s on the CPU.
 
@@
 With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
+In fact, we will add additional observational uncertainty to our mock observational data later during training of the neural posteriro estimater network.
+
 Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.
@@
 \end{figure}
-
-% %The larger volume of data should be able to give tighter statistical constraints on those parameters that are held fixed across the galaxy, but complexity is added since we must allow each star to carry its own set of local ISM parameters.
 
@@
 
-We use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 10 hidden features and 1 transformation layer for the normalizing flow.
+We use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Neural Spline Flow (NSF) \citet{durkan2019neuralsplineflows} containing 20 hidden features and 10 transformation layers for the normalizing flow.
+The NSF uses as invertible transformation smooth and flexible spline functions to allow efficient sampling.
+The expressivity of the spline allows the NDE to be capture complex distributions, while also maintaining computational tractability.
+The final model was selected after extensive hyperparameter tuning, varying: the architecture between NSF, Masked Autoregressive Flow (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity}, and Mask Autoregressive
+Flow with rational-quadratic spline (MAF-RQS), the number of neurons between {10, 20, 50, 100}, and the number of transformations between {1, 5, 10, 30}, optimizing over the test set for both the highest mean log posterior probability and the best calibration as measured by the TARP value.
 

### 2025-02-17 src/scripts/train_torch_chempy.py
[showyourwork] overleaf sync

@@ -11,6 +11,8 @@ from torch.distributions.uniform import Uniform
 from chempy_torch_model import Model_Torch
 
 import time as t
+import os
+import schedulefree
 
 # ---  You want to retrain? ---
 re_train = False
@@ -58,13 +60,15 @@ model = Model_Torch(val_x.shape[1], val_y.shape[1])
 
 # ----- Train the model -------------------------------------------------------------------------------------------------------------------------------------------
 
-optimizer = torch.optim.Adam(model.parameters(), lr=0.001)
+#optimizer = torch.optim.Adam(model.parameters(), lr=0.001)
+optimizer = schedulefree.AdamWScheduleFree(model.parameters(), lr=1e-3)
 loss_fn = torch.nn.MSELoss()
+
 
 # --- Train the neural network ---
 if re_train:
     # --- Load in training data ---
-    path_training = paths.data / 'chempy_data/chempy_TNG_train_data.npz'
+    path_training = paths.data / 'chempy_data/chempy_train_uniform_prior_5sigma.npz' #chempy_TNG_train_data.npz'
     training_data = np.load(path_training, mmap_mode='r')
 
     elements = training_data['elements']
@@ -149,9 +153,10 @@ else:
 
     fig, (ax_box, ax_hist) = plt.subplots(2, sharex=True, gridspec_kw={"height_ratios": (.20, .80)})
 
+    ax_hist.hist(ape.flatten(), bins=100, density=True, cumulative=True, range=(0, 30), color='tomato')
     ax_hist.hist(ape.flatten(), bins=100, density=True, range=(0, 30), color='tomato')
     ax_hist.set_xlabel('Error (%)', fontsize=15)
-    ax_hist.set_ylabel('Density', fontsize=15)
+    ax_hist.set_ylabel('CDF', fontsize=15)
     ax_hist.spines['top'].set_visible(False)
     ax_hist.spines['right'].set_visible(False)
     # percentiles
@@ -167,7 +172,7 @@ else:
     ax_box.spines['right'].set_visible(False)
     ax_box.spines['top'].set_visible(False)
 
-    fig.suptitle('APE of the Neural Network', fontsize=20)
+    #fig.suptitle('APE of the Neural Network', fontsize=20)
     plt.xlim(0, 30)
     fig.tight_layout()
 

### 2025-02-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In order to establish our new method based on SBI we need two ingredients: A simulator (in our case a GCE model) that simulates observational data from a set of model parameters (in our case the IMF slope and the Type Ia supernovae normalization) and a flexible way of parametrizing the posterior density conditioned on the observation in order to perform our inference.
+In order to establish our new method based on SBI we need two ingredients: A simulator (in our case a GCE model) that simulates observational data from a set of model parameters (in our case the IMF slope and the Type Ia supernovae normalization) and a flexible way of parametrizing the posterior density conditioned on the observation in order to perform our inference (see Fig.
+\ref{fig:flowchart} for a schematic visual representation of our method).
 In the next subsections we describe both ingredients in detail.
@@
 
-$\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.75\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\
+$\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.89\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\
 
@@
 shown by \citet{2019ApJ...874..102W,2019arXiv190806113G}), though this is not appropriate in our context.
-To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{portinari} and do not allow for any `hypernovae' \citep[in contrary to][]{2018ApJ...861...40P}).
+To facilitate best comparison with Ilustris TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{portinari} and do not allow for any `hypernovae' \citep[in contrary to][]{2018ApJ...861...40P}).
 
@@
      \vspace{-.5cm}
-     \caption{Absolute percentage error of the NN emulator for the \texttt{CHEMPY} simulator.
-The orange histogram shows the full distribution of percentage errors with the vertical dashed line indicating the median and the vertical dotted lines indicating the first and third quartile.
+     \caption{Cumulative absolute percentage error of the NN emulator for the \texttt{CHEMPY} simulator.
+The orange histogram shows the cumulative distribution of percentage errors with the vertical dashed line indicating the median and the vertical dotted lines indicating the first and third quartile.
 The box plot on the top of the plot extends from the first quartile to the third quartile of the data, with a line at the median.
@@
 
-\paragraph{Bayesian model}
+\subsection{Bayesian model}
 \texttt{CHEMPY} effectively is a Bayesian model for stellar abundances given a set of parameters $\{\Vec{\Lambda},\Vec{\Theta},T\}$ and \citet{Philcox_2019} extended the \texttt{CHEMPY} framework to be able to model multiple stellar data-points.
@@
 
+\paragraph{Posteriors for Galactic parameters}
 As stated in \citet{Philcox_2019} the full posterior for this case is given by 
@@
 
-We use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Neural Spline Flow (NSF) \citet{durkan2019neuralsplineflows} containing 20 hidden features and 10 transformation layers for the normalizing flow.
+We use a NPE \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Neural Spline Flow (NSF) \citet{durkan2019neuralsplineflows} containing 20 hidden features and 10 transformation layers for the normalizing flow.
 The NSF uses as invertible transformation smooth and flexible spline functions to allow efficient sampling.
@@
 The final model was selected after extensive hyperparameter tuning, varying: the architecture between NSF, Masked Autoregressive Flow (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity}, and Mask Autoregressive
-Flow with rational-quadratic spline (MAF-RQS), the number of neurons between {10, 20, 50, 100}, and the number of transformations between {1, 5, 10, 30}, optimizing over the test set for both the highest mean log posterior probability and the best calibration as measured by the TARP value.
+Flow with rational-quadratic spline (MAF-RQS), the number of neurons between {10, 20, 50, 100}, and the number of transformations between {1, 5, 10, 30}, optimizing over the test set for both the highest mean log posterior probability and the best calibration as measured by the TARP value \citep{Lemos2023}.
+For more details on TARP see \ref{sec:sbc} in the appendix.
 
-We train our NPE with $10^5$ data points consisting of $n_{\rm elements}=8$ simulated with the NN emulator described in the previous section with inputs sampled from the priors shown in Tab.
-\ref{tab:priors}.
-Training takes $\lesssim10$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPU cores.
+We train our NPE with $10^5$ data points consisting of $n_{\rm elements}=8$ simulated with the NN emulator described in the previous section with inputs sampled from a uniform prior over the original $5\sigma$-range of the Gaussian priors shown in Tab.
+\ref{tab:priors} to provide better coverage of the parameter space.
+Training takes $\sim70$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPU cores.
 We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator.
@@
 We have made sure that our NPE is well calibrated and posterior distributions are trustable. For more details on simulation-based calibration see appendix \ref{sec:sbc}.
+
+Note, the methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{CHEMPY}.
 
@@
 \paragraph{Multi-star inference}
-Following equation~\ref{eq:posterior} we can calculate the joint posterior for a combined inference using abundance observations of multiple stars.
+Following eq.~\ref{eq:posterior} we can calculate the joint posterior for a combined inference using abundance observations of multiple stars.
 For this, we will condition our NPE individually on single star observations and sample the posterior.
-Then, according to equation~\ref{eq:posterior} we simply have to multiply individual posteriors and account for the inverse weighting by some power of the prior which we take from table~\ref{tab:priors}.
+Then, according to eq.~\ref{eq:posterior} we simply have to multiply individual posteriors and account for the inverse weighting by some power of the prior which we take from table~\ref{tab:priors}.
 
-This does however present one issue, that since we only have access to samples from the posterior and not the posterior itself it is difficult to evaluate equation~\ref{eq:posterior}.
+This does however present one issue, that since we only have access to samples from the posterior and not the posterior itself it is difficult to evaluate eq.~\ref{eq:posterior}.
 We circumvent this by approximating each single star posterior by a Gaussian and fit for the parameters of mean and covariance.
-With this it is straight forward to evaluate equation~\ref{eq:posterior} analytically.
+With this it is straight forward to evaluate eq.~\ref{eq:posterior} analytically.
 In fact, under our assumption the combined posterior is a product between the prior and the product of multiple Gaussians for the individual star posteriors.
@@
 
-Note, the methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{CHEMPY}.
-
 \subsection{\texttt{CHEMPY} Mock Observational Data}
@@
 In order to fully represent observational data, we augment this data with observational uncertainties by adding a Gaussian error of $0.05$ dex for the abundances representative for typical APOGEE data \citep{Majewski2016}.
-All our datasets and accompanying code to reproduce our data generation process and analysis is publicly available\footnote{\url{https://github.com/TobiBu/sbi-chempy}}.
 
@@
 This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
-Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
-Our combined runtime for the NPE training plus sampling is then 610 seconds for 1,000 stars hence our methods making it more than $1,180$ times faster than current HMC methods \citep[cf.][who need $40$h for only 200 stars]{Philcox_2019}.
-But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim72,000$.
-\textcolor{red}{check if those speed-up values are correctly calculated...}
+Our method takes around $50$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
+Our combined runtime for the NPE training plus sampling is then 4250 seconds for 1,000 stars hence our methods making it more than $170$ times faster than current HMC methods \citep[cf.][who need $40$h for only 200 stars]{Philcox_2019}.
+But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim14,400$.
 Hence, shorter computing times make it feasible to use orders of magnitudes more observations.

### 2025-02-17 src/scripts/evaluate_posterior.py

@@ -77,17 +77,17 @@ print(f"Total : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f
 print("")
 
 with open(paths.output / f'global_posterior_APE.txt', 'w') as f:
-    f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")
+    f.write(f"${median:.1f}^{+{u_quantile-median:.1f}}_{-{median-l_quantile:.1f}}\,\%$%")
 
 for i in range(ape.shape[1]):
     l_quantile, median, u_quantile = np.percentile(ape[:,i], [25, 50, 75])
     print(labels_in[i] + f" : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
     if i in [0,1]:
         with open(paths.output / f'posterior_APE.txt', 'a') as f:
-            f.write(f"${median:.1f} + {u_quantile-median:.1f} - {median-l_quantile:.1f}\,\%$%")
+            f.write(f"${median:.1f}^{+{u_quantile-median:.1f}}_{-{median-l_quantile:.1f}}\,\%$%")
 
 # --- Plot calbration using ltu-ili ---
-from metrics import PosteriorCoverage
+from .metrics import PosteriorCoverage
 
 plot_hist = ["coverage", "histogram", "predictions", "tarp"]
 metric = PosteriorCoverage(

### 2025-02-17 src/scripts/plot_functions.py

@@ -386,12 +386,12 @@ def ape_plot(ape, labels_in, save_path):
     colors = ["tomato", "skyblue", "olive", "gold", "teal", "orchid"]
 
     for i in range(ape.shape[1]):
-        ax_hist.hist(ape[:,i], bins=25, density=True, range=(0, 100), label=labels_in[i], color=colors[i], alpha=0.5)
+        ax_hist.hist(ape[:,i], bins=25, density=True, cumulative=True, range=(0, 100), label=labels_in[i], color=colors[i], alpha=0.5)
         median = np.percentile(ape[:,i], 50)
         ax_hist.axvline(median, color=colors[i], linestyle='--')
             
     ax_hist.set_xlabel('Error (%)', fontsize=15)
-    ax_hist.set_ylabel('Density', fontsize=15)
+    ax_hist.set_ylabel('CDF', fontsize=15)
     ax_hist.spines['top'].set_visible(False)
     ax_hist.spines['right'].set_visible(False)
     ax_hist.legend()
@@ -404,7 +404,7 @@ def ape_plot(ape, labels_in, save_path):
     ax_box.spines['right'].set_visible(False)
     ax_box.spines['top'].set_visible(False)
 
-    fig.suptitle('APE of the Posterior', fontsize=20)
+    #fig.suptitle('APE of the Posterior', fontsize=20)
     plt.xlim(0, 100)
     fig.tight_layout()
     plt.savefig(save_path)

### 2025-02-17 src/scripts/train_sbi.py

@@ -16,7 +16,6 @@ from torch.distributions.uniform import Uniform
 
 import time as t
 import pickle
-from tqdm import tqdm
 
 from chempy_torch_model import Model_Torch
 
@@ -32,8 +31,8 @@ labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
 priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
 
 combined_priors = utils.MultipleIndependent(
-    [Normal(p[0]*torch.ones(1), p[1]*torch.ones(1)) for p in priors] +
-    [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
+    [Uniform(p[0]*torch.ones(1)-5*p[1], p[0]*torch.ones(1)+5*p[1]) for p in priors] +
+    [Uniform(torch.tensor([1.0]), torch.tensor([13.8]))],
     validate_args=False)
 
 # --- Set up the model ---
@@ -61,7 +60,7 @@ check_sbi_inputs(simulator, prior)
 
 
 # ----- Train the SBI -------------------------------------------------------------------------------------------------------------------------------------------
-density_estimator_build_fun = posterior_nn(model="maf", hidden_features=10, num_transforms=1, blocks=1)
+density_estimator_build_fun = posterior_nn(model="nsf", hidden_features=20, num_transforms=10, blocks=1)
 inference = NPE_C(prior=prior, density_estimator=density_estimator_build_fun, show_progress_bars=True)
 
 start = t.time()

### 2025-02-17 src/scripts/evaluate_posterior.py
[showyourwork] overleaf sync

@@ -77,14 +77,14 @@ print(f"Total : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f
 print("")
 
 with open(paths.output / f'global_posterior_APE.txt', 'w') as f:
-    f.write(f"${median:.1f}^{+{u_quantile-median:.1f}}_{-{median-l_quantile:.1f}}\,\%$%")
+    f.write(f"${median:.1f}^{+u_quantile-median:.1f}_{-median-l_quantile:.1f}\,\%$%")
 
 for i in range(ape.shape[1]):
     l_quantile, median, u_quantile = np.percentile(ape[:,i], [25, 50, 75])
     print(labels_in[i] + f" : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
     if i in [0,1]:
         with open(paths.output / f'posterior_APE.txt', 'a') as f:
-            f.write(f"${median:.1f}^{+{u_quantile-median:.1f}}_{-{median-l_quantile:.1f}}\,\%$%")
+            f.write(f"${median:.1f}^{+u_quantile-median:.1f}_{-median-l_quantile:.1f}\,\%$%")
 
 # --- Plot calbration using ltu-ili ---
 from .metrics import PosteriorCoverage

### 2025-02-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Our mock data generation and inference then follows the one of Sec.~\ref{subsec:chempy_tng}.
-This means we apply our NPE trained on data with the TNG yield set to data created with the above mentioned alternative yield set.
+This means we apply our NPE that we trained on \texttt{CHEMPY} stellar abundances simulated with the Ilustris TNG yield set to a set of stellar abundances simulated with \texttt{CHEMPY} but using the above mentioned alternative yield set.
 This effectively probes the effect of model missspecification on the inference results.
@@
 We see that the SBI results are less biased than the HMC results especially for $\alpha_{\rm IMF}$.
-In fact with 200 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively.
+In fact with 100 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively.
 Again, the joint posterior is shown in Fig.~\ref{fig:CHEMPY_alt_sbi} and shows that while individually parameter inferences look good, jointly taken the SBI results are on the edge of being $3\sigma$ away from the ground truth.

### 2025-02-17 src/scripts/evaluate_posterior.py

@@ -87,7 +87,7 @@ for i in range(ape.shape[1]):
             f.write(f"${median:.1f}^{+u_quantile-median:.1f}_{-median-l_quantile:.1f}\,\%$%")
 
 # --- Plot calbration using ltu-ili ---
-from .metrics import PosteriorCoverage
+from metrics import PosteriorCoverage
 
 plot_hist = ["coverage", "histogram", "predictions", "tarp"]
 metric = PosteriorCoverage(

### 2025-02-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\begin{figure*}[]
+     \centering
+     \includegraphics[width=\textwidth, trim={6cm, 0cm, 6cm, 0cm}, clip]{figures/ili_coverage.pdf}
+     \includegraphics[width=\textwidth, trim={6cm, 0cm, 6cm, 0cm}, clip]{figures/ili_predictions.pdf}
+     %\vspace{-.5cm}
+     \caption{Top: TARP plot showing the expected coverage probability vs.
+the credibility level $\alpha$ for each of the six individual parameters in our inference.
+The dashed black 1:1-line shows an ideal calibrated posterior and the blue solid line shows the TARP value for our NPE.
+Bottom: }
+     \label{fig:tarp}
+     \script{evaluate_sbi.py}
+\end{figure*}
 % The ... repository, which contains code, additional documentation, and interactive dashboards, is available on GitHub.%\footnote{URL: \url{https://github.com/bGuenes/sbi_chemical_abundances}}
 
-\section{Simulation-based calibration}
+\section{Neural Posterior Calibration}
 \label{sec:sbc}
@@
 
-TARP \citep{Lemos2023} is an alternative calibration check.
+TARP \citep{Lemos2023} is an alternative calibration check for the joint distribution, for which defining a rank is not straightforward.
 Given a test set $(\theta^*,x^*)$ and a set of reference points $\theta_r$, TARP calculates statistics for posterior calibration by - drawing posterior samples $\theta$ given each observation $x^*$ and calculating the distance $r$ between $\theta^*$ and $\theta_r$ counting for how many of the posterior samples the distance to $\theta_r$ is smaller than $r$ \citep[see e.g.
@@
 
+\section{Additional inference results}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/Nstar_comp_different_prior.pdf}
+    \vspace{-.5cm}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for mock data a created with parameters for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$.}
+    \label{fig:N_star_analysis_different_prior}
+    \script{additional_inference.py}
+\end{figure*}
 \end{appendix}

### 2025-02-17 src/scripts/evaluate_posterior.py

@@ -77,14 +77,14 @@ print(f"Total : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f
 print("")
 
 with open(paths.output / f'global_posterior_APE.txt', 'w') as f:
-    f.write(f"${median:.1f}^{+u_quantile-median:.1f}_{-median-l_quantile:.1f}\,\%$%")
+    f.write(f"${median:.1f}^{{+{u_quantile-median:.1f}}}_{{-{median-l_quantile:.1f}}}\,\%$%")
 
 for i in range(ape.shape[1]):
     l_quantile, median, u_quantile = np.percentile(ape[:,i], [25, 50, 75])
     print(labels_in[i] + f" : {median:.1f}% + {u_quantile-median:.1f} - {median-l_quantile:.1f}")
     if i in [0,1]:
         with open(paths.output / f'posterior_APE.txt', 'a') as f:
-            f.write(f"${median:.1f}^{+u_quantile-median:.1f}_{-median-l_quantile:.1f}\,\%$%")
+            f.write(f"${median:.1f}^{{+{u_quantile-median:.1f}}}_{{-{median-l_quantile:.1f}}}\,\%$%")
 
 # --- Plot calbration using ltu-ili ---
 from metrics import PosteriorCoverage

### 2025-02-17 src/scripts/plot_functions.py

@@ -394,7 +394,7 @@ def ape_plot(ape, labels_in, save_path):
     ax_hist.set_ylabel('CDF', fontsize=15)
     ax_hist.spines['top'].set_visible(False)
     ax_hist.spines['right'].set_visible(False)
-    ax_hist.legend()
+    ax_hist.legend(loc='lower right', fontsize=15)
 
     bplot = ax_box.boxplot(ape, vert=False, autorange=False, widths=0.5, patch_artist=True, showfliers=False, boxprops=dict(facecolor='tomato'), medianprops=dict(color='black'))
     for patch, color in zip(bplot['boxes'], colors):

### 2025-02-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
           Berkay Günes\inst{1,2}
+          \and
+          Giuseppe Viterbo\inst{1,2}
+          \and
+          William H. Oliver\inst{1,2}
           \and
@@
 Flow with rational-quadratic spline (MAF-RQS), the number of neurons between {10, 20, 50, 100}, and the number of transformations between {1, 5, 10, 30}, optimizing over the test set for both the highest mean log posterior probability and the best calibration as measured by the TARP value \citep{Lemos2023}.
-For more details on TARP see \ref{sec:sbc} in the appendix.
+For more details on the TARP value see section \ref{sec:sbc} in the appendix.
 
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figures/Nstar_comp_different_prior.pdf}
-    \vspace{-.5cm}
-    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for mock data a created with parameters for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$.}
-    \label{fig:N_star_analysis_different_prior}
-    \script{additional_inference.py}
-\end{figure*}
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\textwidth]{figures/Nstar_comp_different_prior.pdf}
+%     \vspace{-.5cm}
+%     \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for mock data a created with parameters for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$.}
+%     \label{fig:N_star_analysis_different_prior}
+%     \script{additional_inference.py}
+% \end{figure*}
+
 \end{appendix}

### 2025-02-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Training takes $\sim70$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPU cores.
+On an Apple M1 Max training only takes 36 minutes.
 We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator.
@@
 Our method takes around $50$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
-Our combined runtime for the NPE training plus sampling is then 4250 seconds for 1,000 stars hence our methods making it more than $170$ times faster than current HMC methods \citep[cf.][who need $40$h for only 200 stars]{Philcox_2019}.
+Our combined runtime for the NPE training plus sampling is then 4250 (2160 on Apple M!
+Max) seconds for 1,000 stars hence our methods making it more than $170$ ($340$) times faster than current HMC methods \citep[cf.][who need $40$h for only 200 stars]{Philcox_2019}.
 But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim14,400$.
@@
 
-\begin{figure*}[]
-     \centering
-     \includegraphics[width=\textwidth]{figures/sbc_rank_plot_NPE_C.pdf}
-     %\vspace{-.5cm}
-     \caption{SBC ranks of ground truth parameters under the inferred posterior samples for each of the six parameters (red bars).
+% \begin{figure*}[]
+%      \centering
+%      \includegraphics[width=\textwidth]{figures/ili_histogram.pdf}
+%      %\vspace{-.5cm}
+%      \caption{SBC ranks of ground truth parameters under the inferred posterior samples for each of the six parameters (red bars).
 The grey area shows the 99\% confidence interval of a uniform distribution given the number of samples provided.}
-     \label{fig:sbc}
-     \script{evaluate_sbi.py}
-\end{figure*}
+%      \label{fig:sbc}
+%      \script{evaluate_sbi.py}
+% \end{figure*}
 
@@
      \centering
-     \includegraphics[width=\columnwidth]{figures/tarp_plot_NPE_C.pdf}
+     \includegraphics[width=\columnwidth]{figures/ili_tarp.pdf}
      %\vspace{-.5cm}
@@
      \label{fig:tarp}
-     \script{evaluate_sbi.py}
+     \script{plot_sbc.py}
 \end{figure}
@@
      \centering
+     \includegraphics[width=\textwidth, trim={4cm, 0cm, 4cm, 0cm}, clip]{figures/ili_histogram.pdf}
      \includegraphics[width=\textwidth, trim={6cm, 0cm, 6cm, 0cm}, clip]{figures/ili_coverage.pdf}
@@
      %\vspace{-.5cm}
-     \caption{Top: TARP plot showing the expected coverage probability vs.
+     \caption{Posterior calibration diagnostics showing from top to bottom the SBC ranks, the TARP plots and the true vs. predicted parameter plots for each of the six parameters.
+     Top: SBC ranks of ground truth parameters under the inferred posterior samples for each of the six parameters (red bars).
+The grey area shows the 99\% confidence interval of a uniform distribution given the number of samples provided.
+Middle: TARP plot showing the expected coverage probability vs.
 the credibility level $\alpha$ for each of the six individual parameters in our inference.
 The dashed black 1:1-line shows an ideal calibrated posterior and the blue solid line shows the TARP value for our NPE.
-Bottom: }
+Bottom: True vs.
+predicted parameter plots showing the average of the posterior samples and as errorbar the standard deviation of the samples vs their ground truth parameter.}
      \label{fig:tarp}
-     \script{evaluate_sbi.py}
+     \script{plot_sbc.py}
 \end{figure*}

### 2025-02-18 src/scripts/evaluate_posterior.py

@@ -6,8 +6,6 @@ from tqdm import tqdm
 from Chempy.parameter import ModelParameters
 
 from scipy.stats import norm
-from plot_functions import ape_plot
-import matplotlib.pyplot as plt
 
 # ----- Evaluate the posterior -------------------------------------------------------------------------------------------------------------------------------------------
 
@@ -49,8 +47,8 @@ def clean_data(x, y):
 
 val_theta, val_x = clean_data(val_theta, val_x)
 
-val_theta = val_theta[:10000]
-val_x = val_x[:10000]
+val_theta = val_theta[:5000]
+val_x = val_x[:5000]
 
 # convert to torch tensors
 val_theta = torch.tensor(val_theta, dtype=torch.float32)
@@ -88,20 +86,3 @@ for i in range(ape.shape[1]):
     if i in [0,1]:
         with open(paths.output / f'posterior_APE.txt', 'a') as f:
             f.write(f"${median:.1f}^{{+{u_quantile-median:.1f}}}_{{-{median-l_quantile:.1f}}}\,\%$%")
-
-# --- Plot calbration using ltu-ili ---
-from metrics import PosteriorCoverage
-
-plot_hist = ["coverage", "histogram", "predictions", "tarp"]
-metric = PosteriorCoverage(
-    num_samples=1000, sample_method='direct',
-    labels=labels_in,
-    plot_list = plot_hist
-)
-
-fig = metric(
-    posterior=posterior,
-    x=abundances, theta=val_theta)
-
-for i, plot in enumerate(fig):
-    fig[i].savefig(paths.figures / f'ili_{plot_hist[i]}.pdf')

### 2025-02-18 src/scripts/plot_functions.py

@@ -386,7 +386,7 @@ def ape_plot(ape, labels_in, save_path):
     colors = ["tomato", "skyblue", "olive", "gold", "teal", "orchid"]
 
     for i in range(ape.shape[1]):
-        ax_hist.hist(ape[:,i], bins=25, density=True, cumulative=True, range=(0, 100), label=labels_in[i], color=colors[i], alpha=0.5)
+        ax_hist.hist(ape[:,i], bins=25, density=True, cumulative=True, histtype='step', lw=2, range=(0, 100), label=labels_in[i], color=colors[i])#, alpha=0.5)
         median = np.percentile(ape[:,i], 50)
         ax_hist.axvline(median, color=colors[i], linestyle='--')
             

### 2025-02-18 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -388,13 +388,13 @@ def ape_plot(ape, labels_in, save_path):
     for i in range(ape.shape[1]):
         ax_hist.hist(ape[:,i], bins=25, density=True, cumulative=True, histtype='step', lw=2, range=(0, 100), label=labels_in[i], color=colors[i])#, alpha=0.5)
         median = np.percentile(ape[:,i], 50)
-        ax_hist.axvline(median, color=colors[i], linestyle='--')
+        #ax_hist.axvline(median, color=colors[i], linestyle='--')
             
     ax_hist.set_xlabel('Error (%)', fontsize=15)
     ax_hist.set_ylabel('CDF', fontsize=15)
     ax_hist.spines['top'].set_visible(False)
     ax_hist.spines['right'].set_visible(False)
-    ax_hist.legend(loc='lower right', fontsize=15)
+    ax_hist.legend(loc='lower right', fontsize=12)
 
     bplot = ax_box.boxplot(ape, vert=False, autorange=False, widths=0.5, patch_artist=True, showfliers=False, boxprops=dict(facecolor='tomato'), medianprops=dict(color='black'))
     for patch, color in zip(bplot['boxes'], colors):

### 2025-02-18 src/scripts/plot_sbc.py

@@ -1,0 +1,69 @@
+import paths
+import pickle
+import torch
+import numpy as np
+from Chempy.parameter import ModelParameters
+
+
+# ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
+name = "NPE_C"
+
+# --- Define the prior ---
+a = ModelParameters()
+labels_out = a.elements_to_trace
+labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
+
+# ----- load the posterior -------------------------------------------------------------------------------------------------------------------------------------------
+with open(paths.data / f'posterior_{name}.pickle', 'rb') as f:
+    posterior = pickle.load(f)
+
+# --- Load the validation data ---
+# Validation data created with CHEMPY, not with the NN simulator
+print("Evaluating the posterior...")
+path_test = paths.data / 'chempy_data/chempy_TNG_val_data.npz'
+val_data = np.load(path_test, mmap_mode='r')
+
+val_theta = val_data['params']
+val_x = val_data['abundances']
+
+# --- Clean the data ---
+# Chempy sometimes returns zeros or infinite values, which need to removed
+def clean_data(x, y):
+    # Remove all zeros from the training data
+    index = np.where((y == 0).all(axis=1))[0]
+    x = np.delete(x, index, axis=0)
+    y = np.delete(y, index, axis=0)
+
+    # Remove all infinite values from the training data
+    index = np.where(np.isfinite(y).all(axis=1))[0]
+    x = x[index]
+    y = y[index]
+
+    return x, y
+
+val_theta, val_x = clean_data(val_theta, val_x)
+
+val_theta = val_theta[:5000]
+val_x = val_x[:5000]
+
+# convert to torch tensors
+val_theta = torch.tensor(val_theta, dtype=torch.float32)
+val_x = torch.tensor(val_x, dtype=torch.float32)
+abundances =  torch.cat([val_x[:,:2], val_x[:,3:]], dim=1)
+
+# --- Plot calbration using ltu-ili ---
+from metrics import PosteriorCoverage
+
+plot_hist = ["coverage", "histogram", "predictions", "tarp"]
+metric = PosteriorCoverage(
+    num_samples=1000, sample_method='direct',
+    labels=labels_in,
+    plot_list = plot_hist
+)
+
+fig = metric(
+    posterior=posterior,
+    x=abundances, theta=val_theta)
+
[…]

### 2025-02-18 src/scripts/train_sbi.py

@@ -31,7 +31,7 @@ labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
 priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
 
 combined_priors = utils.MultipleIndependent(
-    [Normal(p[0]*torch.ones(1), 3*p[1]*torch.ones(1)) for p in priors] +
+    [Normal(p[0]*torch.ones(1), p[1]*torch.ones(1)) for p in priors] +
     [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
     validate_args=False)
 

### 2025-02-18 src/scripts/train_torch_chempy.py
[showyourwork] overleaf sync

@@ -11,11 +11,10 @@ from torch.distributions.uniform import Uniform
 from chempy_torch_model import Model_Torch
 
 import time as t
-import os
 import schedulefree
 
 # ---  You want to retrain? ---
-re_train = False
+re_train = True #False
 
 # ----- Load the data ---------------------------------------------------------------------------------------------------------------------------------------------
 # ---  Load in the validation data ---
@@ -68,7 +67,7 @@ loss_fn = torch.nn.MSELoss()
 # --- Train the neural network ---
 if re_train:
     # --- Load in training data ---
-    path_training = paths.data / 'chempy_data/chempy_train_uniform_prior_5sigma.npz' #chempy_TNG_train_data.npz'
+    path_training = paths.data / 'chempy_data/chempy_TNG_train_data.npz' #chempy_train_uniform_prior_5sigma.npz'
     training_data = np.load(path_training, mmap_mode='r')
 
     elements = training_data['elements']

### 2025-02-18 src/scripts/train_torch_chempy.py
[showyourwork] overleaf sync

@@ -95,6 +95,7 @@ if re_train:
         train_loss = []
         for i in range(0, train_x.shape[0], batch_size):
             optimizer.zero_grad()
+            optimizer.train()
         
             # Get the batch
             x_batch = train_x[i:i+batch_size].requires_grad_(True)
@@ -112,6 +113,7 @@ if re_train:
             optimizer.step()
         
         # Validation loss
+        optimizer.eval()
         y_pred = model(val_x)
         val_loss = loss_fn(y_pred, val_y)
     

### 2025-02-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\textcolor{red}{to be adapted... especially the zenodo link\\
-Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/TobiBu/sbi-chempy}}}and refer to Appendix \ref{sec:appendix_code_and_data} for an overview of our code and structure.
-All our datasets and network weights are publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}}
+Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/TobiBu/sbi-chempy}}} and refer to Appendix \ref{sec:appendix_code_and_data} for a more extended overview of our code availability.
+All our datasets and network weights are publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/records/14883024}}
 
@@
 
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\textwidth]{figures/Nstar_comp_different_prior.pdf}
-%     \vspace{-.5cm}
-%     \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for mock data a created with parameters for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$.}
-%     \label{fig:N_star_analysis_different_prior}
-%     \script{additional_inference.py}
-% \end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/Nstar_comp_different_prior.pdf}
+    \vspace{-.5cm}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for mock data a created with parameters for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$.}
+    \label{fig:N_star_analysis_different_prior}
+    \script{additional_inference.py}
+\end{figure*}
 

### 2025-02-19 src/scripts/additional_inference.py

@@ -26,7 +26,7 @@ priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_opti
 elements = a.elements_to_trace
 
 # ----- Load posterior -----
-with open(f'data/posterior_{name}.pickle', 'rb') as f:
+with open(paths.data / f'posterior_{name}.pickle', 'rb') as f:
     posterior = pickle.load(f)
 
 

### 2025-02-19 src/scripts/train_sbi.py

@@ -30,17 +30,18 @@ labels_out = a.elements_to_trace
 labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ['time']
 priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
 
+"""
 combined_priors = utils.MultipleIndependent(
     [Normal(p[0]*torch.ones(1), p[1]*torch.ones(1)) for p in priors] +
     [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
     validate_args=False)
-
 """
+
 combined_priors = utils.MultipleIndependent(
-    [Uniform(p[0]*torch.ones(1)-3*p[1], p[0]*torch.ones(1)+3*p[1]) for p in priors] +
+    [Uniform(p[0]*torch.ones(1)-5*p[1], p[0]*torch.ones(1)+5*p[1]) for p in priors] +
     [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
     validate_args=False)
-"""
+
 
 
 # --- Set up the model ---
@@ -68,7 +69,7 @@ check_sbi_inputs(simulator, prior)
 
 
 # ----- Train the SBI -------------------------------------------------------------------------------------------------------------------------------------------
-density_estimator_build_fun = posterior_nn(model="nsf", hidden_features=20, num_transforms=10, blocks=1)
+density_estimator_build_fun = posterior_nn(model="nsf", hidden_features=10, num_transforms=5)#, blocks=1)
 inference = NPE_C(prior=prior, density_estimator=density_estimator_build_fun, show_progress_bars=True)
 
 start = t.time()

### 2025-02-19 src/scripts/train_torch_chempy.py

@@ -14,7 +14,7 @@ import time as t
 import schedulefree
 
 # ---  You want to retrain? ---
-re_train = False
+re_train = True #False
 
 # ----- Load the data ---------------------------------------------------------------------------------------------------------------------------------------------
 # ---  Load in the validation data ---
@@ -67,7 +67,7 @@ loss_fn = torch.nn.MSELoss()
 # --- Train the neural network ---
 if re_train:
     # --- Load in training data ---
-    path_training = paths.data / 'chempy_data/chempy_TNG_train_data.npz' #chempy_train_uniform_prior_5sigma.npz'
+    path_training = paths.data / 'chempy_data/chempy_train_uniform_prior_5sigma.npz' #chempy_TNG_train_data.npz'
     training_data = np.load(path_training, mmap_mode='r')
 
     elements = training_data['elements']

### 2025-02-19 src/scripts/evaluate_emulator.py
[showyourwork] overleaf sync

@@ -1,0 +1,77 @@
+import paths
+import numpy as np
+import matplotlib.pyplot as plt
+
+import torch
+from chempy_torch_model import Model_Torch
+
+# ----- Load the data ---------------------------------------------------------------------------------------------------------------------------------------------
+# ---  Load in the validation data ---
+path_test = paths.data / 'chempy_data/chempy_TNG_val_data.npz'
+val_data = np.load(path_test, mmap_mode='r', allow_pickle=True)
+
+val_x = val_data['params']
+val_y = val_data['abundances']
+
+
+# --- Clean the data ---
+# Chempy sometimes returns zeros or infinite values, which need to be removed
+def clean_data(x, y):
+    # Remove all zeros from the training data
+    index = np.where((y == 0).all(axis=1))[0]
+    x = np.delete(x, index, axis=0)
+    y = np.delete(y, index, axis=0)
+
+    # Remove all infinite values from the training data
+    index = np.where(np.isfinite(y).all(axis=1))[0]
+    x = x[index]
+    y = y[index]
+
+    return x, y
+
+val_x, val_y     = clean_data(val_x, val_y)
+
+# convert to torch tensors
+val_x = torch.tensor(val_x, dtype=torch.float32)
+val_y = torch.tensor(val_y, dtype=torch.float32)
+
+# ----- Define the model ------------------------------------------------------------------------------------------------------------------------------------------
+model = Model_Torch(val_x.shape[1], val_y.shape[1])
+model.load_state_dict(torch.load(paths.data / 'pytorch_state_dict.pt'))
+model.eval()
+print("Model loaded")
+
+# ----- Calculate the Absolute Percantage Error -----
+
+ape = 100 * torch.abs((val_y - model(val_x)) / val_y).detach().numpy()
+
+fig, (ax_box, ax_hist) = plt.subplots(2, sharex=True, gridspec_kw={"height_ratios": (.20, .80)})
+
+ax_hist.hist(ape.flatten(), bins=100, density=True, cumulative=True, range=(0, 30), color='tomato')
+ax_hist.hist(ape.flatten(), bins=100, density=True, range=(0, 30), color='tomato')
+ax_hist.set_xlabel('Error (%)', fontsize=15)
[…]

### 2025-02-19 src/scripts/train_sbi.py

@@ -38,7 +38,7 @@ combined_priors = utils.MultipleIndependent(
 """
 
 combined_priors = utils.MultipleIndependent(
-    [Uniform(p[0]*torch.ones(1)-5*p[1], p[0]*torch.ones(1)+5*p[1]) for p in priors] +
+    [Uniform(p[0]*torch.ones(1)-3*p[1], p[0]*torch.ones(1)+3*p[1]) for p in priors] +
     [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
     validate_args=False)
 
@@ -69,7 +69,7 @@ check_sbi_inputs(simulator, prior)
 
 
 # ----- Train the SBI -------------------------------------------------------------------------------------------------------------------------------------------
-density_estimator_build_fun = posterior_nn(model="nsf", hidden_features=10, num_transforms=5)#, blocks=1)
+density_estimator_build_fun = posterior_nn(model="nsf", hidden_features=20, num_transforms=10)#, blocks=1)
 inference = NPE_C(prior=prior, density_estimator=density_estimator_build_fun, show_progress_bars=True)
 
 start = t.time()

### 2025-02-19 src/scripts/train_torch_chempy.py

@@ -1,7 +1,6 @@
 import paths
 import numpy as np
 import matplotlib.pyplot as plt
-from scipy.stats import norm
 
 from Chempy.parameter import ModelParameters
 
@@ -143,42 +142,3 @@ if re_train:
     plt.tight_layout()
     plt.savefig(paths.figures / "loss_NN_simulator.pdf")
     plt.clf()
-
-
-model.load_state_dict(torch.load(paths.data / 'pytorch_state_dict.pt'))
-print("Model loaded")
-
-# ----- Calculate the Absolute Percantage Error -----
-
-ape = 100 * torch.abs((val_y - model(val_x)) / val_y).detach().numpy()
-
-fig, (ax_box, ax_hist) = plt.subplots(2, sharex=True, gridspec_kw={"height_ratios": (.20, .80)})
-
-ax_hist.hist(ape.flatten(), bins=100, density=True, cumulative=True, range=(0, 30), color='tomato')
-ax_hist.hist(ape.flatten(), bins=100, density=True, range=(0, 30), color='tomato')
-ax_hist.set_xlabel('Error (%)', fontsize=15)
-ax_hist.set_ylabel('CDF', fontsize=15)
-ax_hist.spines['top'].set_visible(False)
-ax_hist.spines['right'].set_visible(False)
-# percentiles
-p1,p2,p3 = np.percentile(ape, [25, 50, 75])
-ax_hist.axvline(p2, color='black', linestyle='--')
-ax_hist.axvline(p1, color='black', linestyle='dotted')
-ax_hist.axvline(p3, color='black', linestyle='dotted')
-ax_hist.text(p2, 0.2, fr'${p2:.1f}^{{+{p3-p2:.1f}}}_{{-{p2-p1:.1f}}}\%$', fontsize=12, verticalalignment='top')
-
-ax_box.boxplot(ape.flatten(), vert=False, autorange=False, widths=0.5, patch_artist=True, showfliers=False, boxprops=dict(facecolor='tomato'), medianprops=dict(color='black'))
-ax_box.set(yticks=[])
-ax_box.spines['left'].set_visible(False)
-ax_box.spines['right'].set_visible(False)
-ax_box.spines['top'].set_visible(False)
-
-#fig.suptitle('APE of the Neural Network', fontsize=20)
-plt.xlim(0, 30)
-fig.tight_layout()
-
[…]

### 2025-02-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
           \and
-          et al.
+          Sven Buder\inst{3,4}
           }
@@
  \and
- Universität Heidelberg, Zentrum für Astronomie, Institut für Theoretische Astrophysik, Albert-Ueberle-Straße 2, D-69120 Heidelberg, Germany\\
+ Universität Heidelberg, Zentrum für Astronomie, Institut für Theoretische Astrophysik, Albert-Ueberle-Straße 2, D-69120 Heidelberg, Germany
+ \and
+ Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia
+ \and
+ ARC Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), Australia\\
  \email{tobias.buck@iwr.uni-heidelberg.de}
@@
 
-Recent advances in stellar spectroscopic surveys, such as APOGEE and GALAH, have produced unprecedented datasets of stellar chemical abundances.
+Recent advances in stellar spectroscopic surveys, such as APOGEE \citep{apogee17} and GALAH \citep{Buder2021,Buder2024}, have produced unprecedented datasets of stellar chemical abundances across a third of the period table.
 These datasets hold the potential to unlock detailed constraints on galactic parameters across diverse environments.
@@
 This study focuses on two critical global galactic parameters: 
-the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF ($\alpha_{\rm IMF}$) and the SN\,Ia normalization ($\log_{10}(N_{\rm Ia})$), the rate of SN\,Ia explosions per unit mass.
+the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF ($\alpha_{\rm IMF}$) and the SN\,Ia normalization, $\log_{10}(N_{\rm Ia})$, the rate of SN\,Ia explosions per unit mass.
 We demonstrate the accuracy, scalability, and robustness of SBI through tests on mock datasets generated by \texttt{CHEMPY} \citep{Rybizki_2017}, as well as on data from hydrodynamical simulations.
@@
 
+\begin{table}[]
+\caption{Alternative nucleosynthetic yield tables used in section \ref{sec: Results} for model misspecification tests.}
+     \centering
+     \begin{tabular}{c|c}
+       Type & Yield Table \\
+        \hline
+         SN\,Ia & \citet{2003NuPhA.718..139T}\\
+         %SN\,II & \citet{2013ARA&A..51..457N}\\
+         AGB & \citet{2016ApJ...825...26K}
+     \end{tabular}
+\label{tab:chempy_ALT_yields}
+\end{table}
+
 \paragraph{Chemical elements}
@@
      \label{fig:ape_NN}
-     \script{train_torch_chempy.py}
+     \script{evaluate_emulator.py}
 \end{figure}
@@
 
-We use our SBI method described in the previous section to infer the global galactic parameters $\alpha_{\rm {IMF}}$ and $\log_{10}(N_{\rm{Ia}})$.
+We use our SBI method described in the previous section to infer the global galactic parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$.
 In order to demonstrate the performance and robustness of our methods we use three mock data-sets:
@@
 With this we ensure to test our training strategy and the performance without any systematic distribution shifts.
-    \item  \texttt{CHEMPY} mock data using a different yield set to test for potential biases through model missspecification in our SBI setup.
+    \item  \texttt{CHEMPY} mock data using a different yield set to test for potential biases through model misspecification in our SBI setup. (see  \ref{tab:chempy_ALT_yields})
     \item Simulated data from a large-scale hydrodynamical simulation taken from the IllustrisTNG suite \citep{Pillepich2018} but with the same yield set as our \texttt{CHEMPY} training data to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
@@
 Our method takes around $50$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
-Our combined runtime for the NPE training plus sampling is then 4250 (2160 on Apple M!
-Max) seconds for 1,000 stars hence our methods making it more than $170$ ($340$) times faster than current HMC methods \citep[cf.][who need $40$h for only 200 stars]{Philcox_2019}.
+Our combined runtime for the NPE training plus sampling is then $2160$ seconds on an Apple M1 Max for 1,000 stars hence our methods making it more than $340$ times faster than current HMC methods which take around $40$h for only $200$ stars \citep{Philcox_2019}.
 But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim14,400$.
@@
 \paragraph{Validation through absolute percentage error}
-We start evaluating our method by using \texttt{CHEMPY} to produce a mock observational dataset to ensure no systematical shft between training and testing data in terms of physics parameters.
+We start evaluating our method by using \texttt{CHEMPY} to produce a mock observational dataset to ensure no systematical shift between training and testing data in terms of physics parameters.
 
@@
 For a single star observation, our NPE has an APE of \variable{output/global_posterior_APE.txt} when looking at all 6 parameters of interest.
-If we restrict ourselves to only the global two parameters, $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$, our NPE achieves an APE of \variable{output/posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}.
-Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high.
-However, for the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ we can boost the accuracy by combining the inference for many stars of the same galaxy.
+If we restrict ourselves to only the two global parameters $\vec\Lambda$, our NPE achieves an APE of \variable{output/posterior_APE.txt} as shown in Fig.~\ref{fig:posterior_APE}.
+There we can also see, that the accuracy for an individual star of the NPE is not particularly high.
+However, for the global parameters $\vec\Lambda$ we can boost the accuracy by combining the inference for many stars of the same galaxy.
 
@@
 
-\textcolor{red}{here goes the assumption that stars are iid since each star is produced by a single call to \texttt{CHEMPY}.
+%\textcolor{red}{here goes the assumption that stars are iid since each star is produced by a single call to \texttt{CHEMPY}.
 However, real physics might correlate stars abundances by star formation histories and cumulative enrichment...}

### 2025-02-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 In addition, the IMF slope may itself be not a constant but rather a function of metallicity, introducing further complexity \citep[e.g.][]{2019MNRAS.482..118G,Martin2019}.
-Similarly, the choice of SN\,Ia delay-time-distribution and normalization plays a crucial role in the enrichment of the ISM \citep[e.g.][]{Buck2021} and is heavily debated \citep{2010ApJ...722.1879M,2012MNRAS.426.3282M,2015ApJ...810..137J}.
+Similarly, the choice of SN\,Ia delay-time-distribution and normalization plays a crucial role in the enrichment of the interstellar medium \citep[ISM; e.g.][]{Buck2021} and is heavily debated \citep{2010ApJ...722.1879M,2012MNRAS.426.3282M,2015ApJ...810..137J}.
+
+\begin{figure*}[]
+     \centering
+     %\vspace{-.25cm}
+     \includegraphics[width=1\linewidth]{figures/sbi_overview.png}
+     \vspace{-.5cm}
+     \caption{SBI flow chart.
+From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} \citep{Rybizki_2017,Philcox_2019} which we use to train a \emph{neural network} emulator to speed up the data generation process.
+Using the \emph{neural network} emulator we produce training data to train the Neural Density Estimator.
+With this we infer the posterior distribution of the model parameters from a single star.
+Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
+     \label{fig:flowchart}
+\end{figure*}
 
@@
 
-\begin{figure*}[]
-     \centering
-     %\vspace{-.25cm}
-     \includegraphics[width=1\linewidth]{figures/sbi_overview.png}
-     \vspace{-.5cm}
-     \caption{SBI flow chart.
-From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} \citep{Rybizki_2017,Philcox_2019} which we use to train a \emph{neural network} emulator to speed up the data generation process.
-Using the \emph{neural network} emulator we produce training data to train the Neural Density Estimator.
-With this we infer the posterior distribution of the model parameters from a single star.
-Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
-     \label{fig:flowchart}
-\end{figure*}
-
 % %--------------------------------------------------------------------
@@
 We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.
-
-In particular, we are using the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model.
-
-\paragraph{\texttt{CHEMPY} parameters}
-In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
-\begin{enumerate}
-     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$.
-We treat these as star-independent and assume them to be constant across galactic environments and cosmic time\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{Chabrier2014,2016MNRAS.462.2832C,2019MNRAS.482..118G,Martin2019}.}.
-     We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}). 
-     %
-     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
-As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, the peak of the star formation rate (SFR), $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$.
-We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
-     %
-     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
-\end{enumerate}
-
-The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
-This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk \citep[which offsets the metallicity for different galactocentric radii, e.g.][for a simulated example]{Buck2020, Wang2024}.
-Similarly \cite{2019ApJ...874..102W} find that ISM parameter variations are deprojected in the [X/Mg] vs [Mg/H] plane (their Fig.\,17) and that abundance tracks in that space are independent of the stellar sample's spatial position within the Galaxy (their Fig.\,3).
-
-Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step.
-This corresponds to the constraint $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)>0.294$.
-For this reason, a truncated Normal prior will be used for the SFR parameter.
-Furthermore, we constrain $T_i$ to the interval $[1,13.8]$\,Gyr (assuming an age of the Universe of 13.8\,Gyr), ignoring any stars formed before $1$\,Gyr, which is justified as these are expected to be rare.
 
@@
 
-\paragraph{Nucleosynthetic yield tables}
+In particular, we are using the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model.
+
+\paragraph{\texttt{CHEMPY} parameters:}
+In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
+\begin{enumerate}
+     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, which effectively sets the number of CC-SNe a SSP generates and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$, which controls the total number of SN Is per SSP.
+We treat these as star-independent and assume them to be constant across galactic environments and cosmic time\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{Chabrier2014,2016MNRAS.462.2832C,2019MNRAS.482..118G,Martin2019}.}.
+     We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}). 
+     %
+     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$.
+As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, which qunatifies the star formation rate per unit gas, the peak of the star formation rate (SFR), $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$.
+We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
+     %
+     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
+\end{enumerate}
+
+The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance.
+This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk \citep[which offsets the metallicity for different galactocentric radii, e.g.][for a simulated example]{Buck2020, Wang2024}.
+Similarly \cite{2019ApJ...874..102W} find that ISM parameter variations are deprojected in the [X/Mg] vs [Mg/H] plane (their Fig.\,17) and that abundance tracks in that space are independent of the stellar sample's spatial position within the Galaxy (their Fig.\,3).
+
+Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$\footnote{\texttt{CHEMPY} parametrizes the SFR with a $\Gamma$ distribution
+\begin{equation*}
+\label{eq:SFR}
+\mathrm{SFR}\left(t,k,\vartheta\right)=\frac{1}{\Gamma(k)\vartheta^k}t^{k-1}\exp\left(\frac{-t}{\vartheta}\right), \mathrm{for}\,k=2 \rightarrow \vartheta = \mathrm{SFR}_\mathrm{peak} 
+\end{equation*}
+where the shape parameter is fixed to $k=2$ such that the scale parameter ($\vartheta$) determines the peak of the SFR}) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step.
+In our formalism, this corresponds to the constraint $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)>0.294$.
+For this reason, a truncated Gaussian prior will be used for the SFR parameter.
+Furthermore, we constrain $T_i$ to the interval $[1,13.8]$\,Gyr (assuming an age of the Universe of 13.8\,Gyr), ignoring any stars formed before $1$\,Gyr, which is justified as these are expected to be rare.
+
+\paragraph{Nucleosynthetic yield tables:}
 We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
@@
 
-\paragraph{Chemical elements}
+\paragraph{Chemical elements:}
 In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si since these are the only elements traced by TNG.
@@
 Here $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}.
-This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{CHEMPY}\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe, as in \citep{2019ApJ...874..102W}, since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}.
+As is customary, we use H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{CHEMPY}\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe, as in \citep{2019ApJ...874..102W}, since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}.
 
-With these modifications, \texttt{CHEMPY} allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters.
-It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{CHEMPY} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
+With these modifications, \texttt{CHEMPY} allows to predict TNG-like chemical abundances for a given set of galactic parameters.
+It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, it is thus not certain \textit{a priori} how useful \texttt{CHEMPY} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}.
 However,  such a test is necessary to prepare for an inference on real data.
@@
 With the NN emulator, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{CHEMPY}.
-    \item \textbf{Differentiability:} The NN is written in pytorch which allows for automatic differentiation.
-Additionally the NN has a simple closed-form analytic structure \citep[described in the appendix of][]{Philcox_2019}, unlike the complex \texttt{CHEMPY} model.
-This allows it to be differentiated, so one can use it to sample via advanced methods such HMC as done in \citep{Philcox_2019}.
+    \item \textbf{Differentiability:} The NN is written in pytorch and has additionally a simple closed-form analytic structure \citep[described in the appendix of][]{Philcox_2019}, unlike the complex \texttt{CHEMPY} model.
+Both aspects allow it to be differentiated either via auto-diff or analytically, so one can use it to sample via advanced methods such HMC as done in \citep{Philcox_2019}.
 \end{enumerate}
 
-Despite the additional complexity introduced by using multiple stellar data-points, our NN simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
+Despite the additional complexity introduced by using multiple stellar data-points, our NN simply needs to predict the birth-time abundances for a single star (with index $i$) from a given set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$.
 The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{CHEMPY} to a single matrix computation.
-With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
+With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of \variable{output/ape_NN_log.txt} dex)  which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
 In fact, we will add additional observational uncertainty to our mock observational data later during training of the neural posteriro estimater network.

### 2025-02-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{table}[]
-\caption{Alternative nucleosynthetic yield tables used in section \ref{sec: Results} for model misspecification tests.}
-     \centering
-     \begin{tabular}{c|c}
-       Type & Yield Table \\
-        \hline
-         SN\,Ia & \citet{2003NuPhA.718..139T}\\
-         %SN\,II & \citet{2013ARA&A..51..457N}\\
-         AGB & \citet{2016ApJ...825...26K}
-     \end{tabular}
-\label{tab:chempy_ALT_yields}
-\end{table}
-
 \paragraph{Chemical elements:}
@@
 The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{CHEMPY} to a single matrix computation.
-With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of \variable{output/ape_NN_log.txt} dex)  which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
-In fact, we will add additional observational uncertainty to our mock observational data later during training of the neural posteriro estimater network.
-
-Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.
+With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of \variable{output/ape_NN_log.txt} dex)  which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN})\footnote{Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.}.
+In fact, we will add additional observational uncertainty to our mock observational data later during training of the neural posterior estimator network.
 
@@
 For better cross-comparison we decided to use the same alternative yield set as presented in Tab.~5 of \citep{Philcox_2019}.
+
+\begin{table}[]
+\caption{Alternative nucleosynthetic yield tables used in section \ref{sec: Results} for model misspecification tests.}
+     \centering
+     \begin{tabular}{c|c}
+       Type & Yield Table \\
+        \hline
+         SN\,Ia & \citet{2003NuPhA.718..139T}\\
+         %SN\,II & \citet{2013ARA&A..51..457N}\\
+         AGB & \citet{2016ApJ...825...26K}
+     \end{tabular}
+\label{tab:chempy_ALT_yields}
+\end{table}
+
 In short this yield set uses SN Ia yields from \citet{2003NuPhA.718..139T}, SN II yields from \citet{2013ARA&A..51..457N} and AGB yields from \citet{2016ApJ...825...26K}.

### 2025-02-19 src/scripts/evaluate_emulator.py

@@ -75,3 +75,6 @@ plt.clf()
 
 with open(paths.output / 'ape_NN.txt', 'w') as f:
     f.write(f'${p2:.1f}^{{+{p3-p2:.1f}}}_{{-{p2-p1:.1f}}}\,\%$%')
+
+with open(paths.output / 'ape_NN_log.txt', 'w') as f:
+    f.write(f'${np.log10(1+p2):.1f}\,\%$%')

### 2025-02-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{CHEMPY} to a single matrix computation.
-With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of \variable{output/ape_NN_log.txt} dex)  which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN})\footnote{Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.}.
+With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of %\variable{output/ape_NN_log.txt} dex)
+which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN})\footnote{Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.}.
 In fact, we will add additional observational uncertainty to our mock observational data later during training of the neural posterior estimator network.
@@
 In fact, any tabulated yield set will differ from reality and during an application of our inference pipeline it will not be clear which tabulated yield set most closely matches reality and hence which should be used.
-In order to investigate how sensitive our method is to a model missspecification by using an incorrect yield set, we create another set of mock data using \texttt{CHEMPY} with a different yield set than during training of our NPE.
-For better cross-comparison we decided to use the same alternative yield set as presented in Tab.~5 of \citep{Philcox_2019}.
+In order to investigate how sensitive our method is to a model misspecification by using an incorrect yield set, we create another set of mock data using \texttt{CHEMPY} with a different yield set than during training of our NPE.
+For better cross-comparison we decided to use the same alternative yield sets as presented in Tab.~5 of \citep{Philcox_2019}.
 
@@
         \hline
-         SN\,Ia & \citet{2003NuPhA.718..139T}\\
-         %SN\,II & \citet{2013ARA&A..51..457N}\\
+         SN\,Ia & \citet{2003NuPhA.718..139T} \\
+         SN\,II & \citet{2013ARA&A..51..457N} \\
          AGB & \citet{2016ApJ...825...26K}
@@
 
-In short this yield set uses SN Ia yields from \citet{2003NuPhA.718..139T}, SN II yields from \citet{2013ARA&A..51..457N} and AGB yields from \citet{2016ApJ...825...26K}.
 By choosing this set of yields we have made sure that contributions to all three processes differ by $\mathcal{O}(10\%)$ \citep[see Sec.~6.2 of][for more details]{Philcox_2019}.

### 2025-02-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \end{eqnarray}
-where $p_V(V_i)$ is the prior on variable $V_i$ (belonging to the set $V$).
+where $p_{\vec\Theta}(\vec\Theta_i)p_{T_i}(T_i)$ are the priors on the variables $\vec\Theta_i$ or $T_i$ belonging to a given set of stars.
 
@@
 
-\paragraph{Simulation-based inference}
+\paragraph{Simulation-based inference:}
 In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a Bayesian inference framework -- works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\Vec{\theta}$ (in our case a GCE model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\Vec{\theta}|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\Vec{\theta}$ that reproduce the simulated observations $\Vec{X}$.
-Once this mapping is trained, we can apply it to observations of stellar abundances $\Vec{X_R}$ to infer $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$) (see Fig.
-\ref{fig:flowchart} for a schematic visual representation of our method).
+Once this mapping is trained, we can apply it to observations of stellar abundances $\Vec{X_R}$ to infer $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$).
+We show a schematic visual representation of our method in Fig.
+\ref{fig:flowchart}.
 Note, we do not need to know anything about $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$), we solely need to be able to sample from it.
 
-We use a NPE \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Neural Spline Flow (NSF) \citet{durkan2019neuralsplineflows} containing 20 hidden features and 10 transformation layers for the normalizing flow.
+We use a NPE \citep{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Neural Spline Flow \citep[NSF;][]{durkan2019neuralsplineflows} containing 20 hidden features and 10 transformation layers for the normalizing flow.
 The NSF uses as invertible transformation smooth and flexible spline functions to allow efficient sampling.
 The expressivity of the spline allows the NDE to be capture complex distributions, while also maintaining computational tractability.
-The final model was selected after extensive hyperparameter tuning, varying: the architecture between NSF, Masked Autoregressive Flow (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity}, and Mask Autoregressive
+The final model was selected after extensive hyperparameter tuning, varying: the architecture between NSF, Masked Autoregressive Flow \citep[MAF;][]{papamakarios2018maskedautoregressiveflowdensity}, and Mask Autoregressive
 Flow with rational-quadratic spline (MAF-RQS), the number of neurons between {10, 20, 50, 100}, and the number of transformations between {1, 5, 10, 30}, optimizing over the test set for both the highest mean log posterior probability and the best calibration as measured by the TARP value \citep{Lemos2023}.
@@
 
-We train our NPE with $10^5$ data points consisting of $n_{\rm elements}=8$ simulated with the NN emulator described in the previous section with inputs sampled from a uniform prior over the original $5\sigma$-range of the Gaussian priors shown in Tab.
+We train our NPE with $10^5$ data points consisting of $n_{\rm elements}=8$ simulated with the NN emulator described in the previous section.
+Inputs are sampled from a uniform prior over the original $5\sigma$-range of the Gaussian priors shown in Tab.
 \ref{tab:priors} to provide better coverage of the parameter space.
@@
 
-Note, the methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{CHEMPY}.
+Note, the methods presented here can naturally be extended to any fast and flexible GCE model \citep[e.g.][and others]{Talbot1971,vice,omega2018}, not just \texttt{CHEMPY}.
 
-\begin{figure}[]
-     \centering
-     \includegraphics[width=\columnwidth]{figures/ape_posterior_NPE_C.pdf}
-     \vspace{-.5cm}
-     \caption{Absolute percentage error of the neural posterior density estimate for a single star.
-Different colored histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines.
-The box plots show again the first and third quantiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box.
-The global parameters of main interest for this work are shown by the light blue and red histogram.}
-     \label{fig:posterior_APE}
-     \script{plot_posterior_APE.py}
-\end{figure}
-
-\paragraph{Multi-star inference}
+\paragraph{Multi-star inference:}
 Following eq.~\ref{eq:posterior} we can calculate the joint posterior for a combined inference using abundance observations of multiple stars.
 For this, we will condition our NPE individually on single star observations and sample the posterior.
-Then, according to eq.~\ref{eq:posterior} we simply have to multiply individual posteriors and account for the inverse weighting by some power of the prior which we take from table~\ref{tab:priors}.
+Then, according to eq.~\ref{eq:posterior} we simply have to multiply individual posteriors and account for the inverse weighting by some power of the prior which we take from Table~\ref{tab:priors}.
 
@@
 In fact, under our assumption the combined posterior is a product between the prior and the product of multiple Gaussians for the individual star posteriors.
-The latter product is also a Gaussian with mean $\mathbf{\mu'}$ and variance $\mathbf{\sigma'}$:
+The latter product is also a Gaussian with mean $\mathbf{\mu'}$ and variance $\mathbf{\sigma'^2}$:
 \begin{align}
@@
 
+\begin{figure}[]
+     \centering
+     \includegraphics[width=\columnwidth]{figures/ape_posterior_NPE_C.pdf}
+     \vspace{-.5cm}
+     \caption{Absolute percentage error of the neural posterior density estimate for a single star.
+Different colored histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines.
+The box plots show again the first and third quantiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box.
+The global parameters of main interest for this work are shown by the light blue and red histogram.}
+     \label{fig:posterior_APE}
+     \script{plot_posterior_APE.py}
+\end{figure}
+
 We use our SBI method described in the previous section to infer the global galactic parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$.
@@
 \subsection{\texttt{CHEMPY} Mock Observational Data}
+\label{subsec:mock obs}
 Our analysis uses mock observations drawn from the neural network emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and varying local parameters $\vec{\Theta}_i$ using priors from Tab.
@@
 \end{figure*}
+
+\begin{figure}
+%\vspace{-1.3cm}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields.png}
+    \vspace{-.25cm}
+    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1,000$ stars.
+The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.
+See Sec.~\ref{subsec:chempy_tng} for a full description.}
+    \label{fig:CHEMPY_TNG_sbi} 
+    \script{chempy_tng_inference.py}
+\end{figure}
 \begin{figure*}
@@
 
-\paragraph{Validation through absolute percentage error}
+\paragraph{Validation through absolute percentage error:}
 We start evaluating our method by using \texttt{CHEMPY} to produce a mock observational dataset to ensure no systematical shift between training and testing data in terms of physics parameters.
@@
 
+However, currently our NPE network is not particularly good at estimating ages from abundances alone.
+On average our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data.
+
 \subsection{Inference on mock data from \texttt{CHEMPY} with TNG yield set}
 \label{subsec:chempy_tng}
-
-\begin{figure}
-%\vspace{-1.3cm}
-    \centering
-    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields.png}
-    \vspace{-.25cm}
-    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1,000$ stars.
-The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.
-See Sec.~\ref{subsec:chempy_tng} for a full description.}
-    \label{fig:CHEMPY_TNG_sbi} 
-    \script{chempy_tng_inference.py}
-\end{figure}
-
 %
 Combining the inference on multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$.
-We therefore perform inference using a range of stars $n\in[1,1000]$ (see Fig.
-\ref{fig:CHEMPY_TNG_N_star_analysis}) to show how inference accuracy increases with number of observations.
-We find that in the limit of few stars (less than $\sim100$) SBI shows a larger uncertainties than the HMC results of \citet{Philcox_2019}.
-But in the limit of large numbers of stars (few hundred) the accuracy and precision is superior compared to HMC.
-In particular, after using a few tens of stars, the NPE estimate is less biased than the HMC results.
+We therefore perform inference using a range of stars $n\in[1,1000]$ to show how inference accuracy increases with number of observations (see Fig.
+\ref{fig:CHEMPY_TNG_N_star_analysis}).
+We find that in the limit of less than $\sim100$ stars SBI shows a larger uncertainties than the HMC results of \citet{Philcox_2019}.
+But when using more than a few hundred stars the accuracy and precision is superior compared to HMC.
+In particular, after using a few tens of stars, the NPE estimate is already less biased than the HMC results.
 Finally, given our computational advantage we will be able to use orders of magnitude more stars for our inference.
@@
 %
-Using a sample of 1,000 stars we infer \variable{output/CHEMPY_TNG_sbi.txt} which deviates less than $\sim0.04\%$ from the correct value.
+Using a sample of 1,000 stars we infer \variable{output/CHEMPY_TNG_sbi.txt} which deviates less than $\sim0.04\%$ from the ground truth value.
 
@@
 This would allow us to estimate/infer stellar ages as well.
-However, currently our NPE network is not particularly good at estimating ages from abundances alone.
-On average our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data.
+But note, as discussed above at the end of sub-section \ref{subsec:mock obs} our NPE currently is not well calibrated to estimate stellar ages accurately enough.
 
@@
 In fact, any tabulated yield set will differ from reality and during an application of our inference pipeline it will not be clear which tabulated yield set most closely matches reality and hence which should be used.
-In order to investigate how sensitive our method is to a model misspecification by using an incorrect yield set, we create another set of mock data using \texttt{CHEMPY} with a different yield set than during training of our NPE.
+In order to investigate how sensitive our method is to model misspecification by using an incorrect yield set, we create another set of mock data using \texttt{CHEMPY} with a different yield set (Tab.
+\ref{tab:chempy_ALT_yields}) than during training of our NPE.
 For better cross-comparison we decided to use the same alternative yield sets as presented in Tab.~5 of \citep{Philcox_2019}.
@@
 \begin{table}[]
-\caption{Alternative nucleosynthetic yield tables used in section \ref{sec: Results} for model misspecification tests.}
+\caption{Alternative nucleosynthetic yield tables used for model misspecification tests.}
      \centering
@@
          SN\,Ia & \citet{2003NuPhA.718..139T} \\
-         SN\,II & \citet{2013ARA&A..51..457N} \\
+         %SN\,II & \citet{2013ARA&A..51..457N} \\
          AGB & \citet{2016ApJ...825...26K}
@@
 Our mock data generation and inference then follows the one of Sec.~\ref{subsec:chempy_tng}.
-This means we apply our NPE that we trained on \texttt{CHEMPY} stellar abundances simulated with the Ilustris TNG yield set to a set of stellar abundances simulated with \texttt{CHEMPY} but using the above mentioned alternative yield set.
-This effectively probes the effect of model missspecification on the inference results.
+This means we apply our NPE that we trained on \texttt{CHEMPY} stellar abundances simulated with the Ilustris TNG yield set to a sets of stellar abundances simulated with \texttt{CHEMPY} but using the alternative yield sets mentioned in Tab.
+\ref{tab:chempy_ALT_yields}.
+This effectively probes the effect of model misspecification on the inference results.
 
-We show our inference results for this setup for varying number of stars (blue) in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis} comparing again against HMC results (red) from \citet{Philcox_2019}.
-We see that the SBI results are less biased than the HMC results especially for $\alpha_{\rm IMF}$.
+This isn't up to date
+\textcolor{blue}{
+In Fig.~\ref{fig:CHEMPY_alt_N_star_analysis} we show our inference results (blue) for this setup for varying number of stars and compare them against HMC results (red) from \citet{Philcox_2019}.
+We see that the SBI results are less biased than the HMC results, especially for $\alpha_{\rm IMF}$.
 In fact with 100 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively.
 Again, the joint posterior is shown in Fig.~\ref{fig:CHEMPY_alt_sbi} and shows that while individually parameter inferences look good, jointly taken the SBI results are on the edge of being $3\sigma$ away from the ground truth.
+}
 
+\textcolor{blue}{
 Nevertheless, the performance of our inference is still very good.
@@
 This is next to drastically reduced compute times -- another a key advantage of our SBI method compared to more standard approaches such as HMC.
+}
 
@@
     \item Robustness to model misspecifications: Unlike HMC, which shows significant biases when faced with discrepancies between model assumptions and data, SBI demonstrates remarkable robustness.
-Even under conditions of mismatched yield tables or data generated from hydrodynamical simulations, SBI provides accurate and reliable results.
+\textcolor{blue}{Even under conditions of mismatched yield tables (should be updated)} or data generated from hydrodynamical simulations, SBI provides accurate and reliable results.
 This robustness ensures SBI’s applicability in real-world scenarios where exact model fidelity cannot be guaranteed.

### 2025-02-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Inference on $1,000$ stars yielded precise estimates for the IMF slope ($\alpha_{\rm IMF} = -2.296 \pm 0.009$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = -2.889 \pm 0.010$), deviating by less than 0.04\% from the ground truth.
-SBI also demonstrated greater robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation.}
+\textcolor{blue}{SBI also demonstrated greater robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation (not with the new model :( )}.}
   % conclusions heading (optional), leave it empty if necessary 
@@
 We design the NN as a simple feed-forward neural network with 2 hidden layers and 100 neurons in the first and 40 neurons in the second layer.
-The NN is trained on $\sim100,000$ data points and validated on $\sim50,000$ additional data points created with \texttt{CHEMPY} using a uniform prior over the $5\sigma$-range of the original Gaussian prior stated in table~\ref{tab:priors}.
+The NN is trained on $\sim700,000$ data points and validated on $\sim50,000$ additional data points created with \texttt{CHEMPY} using a uniform prior over the $5\sigma$-range of the original Gaussian prior stated in table~\ref{tab:priors}.
 The batch size is set to 64 and the learning rate is set to 0.001.
@@
 \ref{tab:priors} to provide better coverage of the parameter space.
-Training takes $\sim70$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPU cores.
-On an Apple M1 Max training only takes 36 minutes.
+Training takes $\sim35$ minutes on an Apple M1 Max.
 We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator.
@@
     \vspace{-.5cm}
-    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line).
+    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars.
+We compare our SBI results (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line) for various test cases as described in Sec.
 For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainties as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).
@@
 There is an extensive discussion in the literature about stellar nucleosynthesis with various different yield sets proposed \citep[see e.g.
-discussion in][]{Rybizki_2017}.
-In fact, any tabulated yield set will differ from reality and during an application of our inference pipeline it will not be clear which tabulated yield set most closely matches reality and hence which should be used.
+the discussion in][]{Rybizki_2017}.
+In fact, all tabulated yield sets currently differ from reality and during an application of our inference pipeline it will not be clear which tabulated yield set most closely matches reality and hence which should be used.
 In order to investigate how sensitive our method is to model misspecification by using an incorrect yield set, we create another set of mock data using \texttt{CHEMPY} with a different yield set (Tab.
@@
 
-By choosing this set of yields we have made sure that contributions to all three processes differ by $\mathcal{O}(10\%)$ \citep[see Sec.~6.2 of][for more details]{Philcox_2019}.
+By choosing this set of yields we have made sure that contributions to all three processes differ by $\mathcal{O}(10\%)$. For more details see also Sec.~6.2 of \citet{Philcox_2019}.
 
@@
 Looking at Fig.~\ref{fig:CHEMPY_alt_sbi} we see that our inference is still consistent with the ground truth within the $3\sigma$ levels.
-Comparing this to the HMC results, we see that SBI is performing much better when the model is misspecified and leads to significantly less biased results.
-This is next to drastically reduced compute times -- another a key advantage of our SBI method compared to more standard approaches such as HMC.
+Comparing this to the HMC results (red band in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis}), we see that SBI is performing much better when the model is misspecified and leads to significantly less biased results.
+This is -- next to drastically reduced compute times -- another a key advantage of our SBI method compared to more standard approaches such as HMC.
 }
@@
 
-\subsection{Inference on mock data from the IllustrisTNG simulation}
+\subsection{Inference on mock data from a IllustrisTNG simulated galaxy}
 \label{subsec:tng_sim}
@@
 In the parametrization of \texttt{CHEMPY} as used here, we can assign each star to its own ISM environment, but we cannot exchange gas between environments and do not model sudden star formation or infall events.
-Hence, this section is dedicated to investigate whether this significantly biases our inference of the SSP parameters (noting that results from \citet{2019ApJ...874..102W} justify the treatment of ISM parameters as latent variables).
+Hence, this section is dedicated to investigating whether this significantly biases our inference of the SSP parameters (noting that results from \citet{2019ApJ...874..102W} justify the treatment of ISM parameters as latent variables).
 
@@
 
-Looking at the joint posterior in Fig.~\ref{fig:CHEMPY_TNG_sbi} we see that also in this case SBI recovers the true parameters with less than $3\sigma$ bias.
+Looking at the joint posteriors in Figs.~\ref{fig:CHEMPY_TNG_sbi}, \ref{fig:CHEMPY_alt_sbi} and \ref{fig:CHEMPY_TNG_sim_sbi} we see that also in this case SBI recovers the true parameters with less than $3\sigma$ bias.
 
@@
 
-The applicability of our method extends to future spectroscopic surveys, such as those planned by the 4MOST or WEAVE consortia, which will provide orders of magnitude more data than current datasets.
+The applicability of our method extends to surveys with more observed elements especially neutron capture elements such as GALAH DR4 \citep{Buder2024} and future spectroscopic surveys, such as those planned by the 4MOST \citep{4most} or WEAVE \citep{weave} consortia, which will provide orders of magnitude more data than current datasets.
 Our findings indicate that SBI can seamlessly scale to such large datasets, offering significant advantages in terms of speed and computational cost.

### 2025-02-20 src/scripts/train_sbi.py
[showyourwork] overleaf sync

@@ -32,7 +32,7 @@ priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_opti
 
 
 combined_priors = utils.MultipleIndependent(
-    [Normal(2*p[0]*torch.ones(1), 2*p[1]*torch.ones(1)) for p in priors] +
+    [Normal(p[0]*torch.ones(1), p[1]*torch.ones(1)) for p in priors] +
     [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
     validate_args=False)
 

### 2025-02-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Constraining these parameters is essential for advancing our understanding of stellar feedback, metal enrichment, and galaxy formation processes.
-However, traditional Bayesian inference techniques, such as Hamiltonian Monte Carlo (HMC), are computationally prohibitive when applied to the large datasets produced by modern stellar surveys.}
+However, traditional Bayesian inference techniques, such as Hamiltonian Monte Carlo (HMC), are computationally prohibitive when applied to large datasets of modern stellar surveys.}
   % aims heading (mandatory)
-   {This study aims to leverage simulation-based inference (SBI) as a scalable, robust, and efficient method for constraining galactic parameters from stellar chemical abundances.
-By focusing on the high-mass IMF slope and SN\,Ia normalization, we aim to demonstrate the advantages of SBI over HMC in terms of speed, scalability, and robustness against model misspecifications.}
+   {We leverage simulation-based-inference (SBI) as a scalable, robust, and efficient method for constraining galactic parameters from stellar chemical abundances and demonstrate its the advantages over HMC in terms of speed, scalability, and robustness against model misspecifications.}
   % methods heading (mandatory)
-   {We combined a Galactic Chemical Evolution (GCE) model, \texttt{CHEMPY}, with a neural network emulator and a Neural Posterior Estimator (NPE) to train an SBI pipeline.
-Mock datasets were generated using \texttt{CHEMPY}, including scenarios with mismatched nucleosynthetic yields, and additional tests were conducted on data from a Milky Way-like galaxy taken from a hydrodynamical simulation.
-SBI results were benchmarked against HMC-based inference, focusing on computational performance, accuracy, and resilience to systematic discrepancies.}
+   {We combine a Galactic Chemical Evolution (GCE) model, \texttt{CHEMPY}, with a neural network emulator and a Neural Posterior Estimator (NPE) to train our SBI pipeline.
+Mock datasets are generated using \texttt{CHEMPY}, including scenarios with mismatched nucleosynthetic yields, with additional tests conducted on data from a simulated Milky Way-like galaxy.
+SBI results are benchmarked against HMC-based inference, focusing on computational performance, accuracy, and resilience to systematic discrepancies.}
   % results heading (mandatory)
-   {Our SBI approach achieves a $\sim1,000\times$ ($\sim72,000\times$) speed-up compared to HMC, reducing inference runtime from $\sim40$ hours for 200 stars to mere minutes for thousands of stars.
-Inference on $1,000$ stars yielded precise estimates for the IMF slope ($\alpha_{\rm IMF} = -2.296 \pm 0.009$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = -2.889 \pm 0.010$), deviating by less than 0.04\% from the ground truth.
-\textcolor{blue}{SBI also demonstrated greater robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation (not with the new model :( )}.}
+   {SBI achieves a $\sim1,000\times$ speed-up compared to HMC, reducing inference runtime from $\gtrsim40$ hours to mere minutes for thousands of stars.
+Inference on $1,000$ stars yields precise estimates for the IMF slope ($\alpha_{\rm IMF} = -2.296 \pm 0.009$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = -2.889 \pm 0.010$), deviating less than 0.04\% from the ground truth.
+SBI also demonstrates greater robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation.}
   % conclusions heading (optional), leave it empty if necessary 
-   {SBI represents a paradigm shift in galactic chemical enrichment studies, enabling efficient and precise analysis of massive stellar datasets.
-By outperforming HMC in speed, scalability, and robustness, SBI is poised to become a cornerstone methodology for future spectroscopic surveys, such as 4MOST and WEAVE, facilitating deeper insights into the chemical and dynamical evolution of galaxies.}
+   {SBI represents a paradigm shift in GCE studies, enabling efficient and precise analysis of massive stellar datasets.
+By outperforming HMC in speed, scalability, and robustness, SBI is poised to become a cornerstone methodology for future spectroscopic surveys facilitating deeper insights into the chemical and dynamical evolution of galaxies.}
 

### 2025-02-20 src/scripts/plot_functions.py
[showyourwork] overleaf sync

@@ -419,8 +419,27 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
     mu_log10N_Ia, sigma_log10N_Ia = mean_std(log10_N_Ia, global_params[0,1], global_params[1,1])
 
     # create a grid of points
-    grid_x = [-2.35,-2.25]
-    grid_y = [-3.0,-2.84]
+    #grid_x = [-2.35,-2.25]
+    #grid_y = [-3.0,-2.84]
+
+    xlim = [-0.05, 0.05]
+    ylim = [-0.05, 0.05]
+
+    if np.abs(global_params[0,0]-mu_alpha) > 0.05:
+        if mu_alpha-global_params[0,0] < 0:
+            xlim[0] = mu_alpha-global_params[0,0]-10*sigma_alpha
+        elif mu_alpha-global_params[0,0] > 0:
+            xlim[1] = mu_alpha-global_params[0,0]+10*sigma_alpha
+
+    if np.abs(global_params[0,1]-mu_log10N_Ia) > 0.05:
+        if mu_log10N_Ia-global_params[0,1] < 0:
+            ylim[0] = mu_log10N_Ia-global_params[0,1]-10*sigma_log10N_Ia
+        elif mu_log10N_Ia-global_params[0,1] > 0:
+            ylim[1] = mu_log10N_Ia-global_params[0,1]+10*sigma_log10N_Ia
+
+    grid_x = [global_params[0,0]+xlim[0], global_params[0,0]+xlim[1]]
+    grid_y = [global_params[0,1]+ylim[0], global_params[0,1]+ylim[1]]
+
     x, y = np.mgrid[grid_x[0]:grid_x[1]:0.001, grid_y[0]:grid_y[1]:0.001]
     pos = np.dstack((x, y))
 

### 2025-02-20 src/scripts/train_torch_chempy.py

@@ -13,7 +13,7 @@ import time as t
 import schedulefree
 
 # ---  You want to retrain? ---
-re_train = True #False
+re_train = False #True
 
 # ----- Load the data ---------------------------------------------------------------------------------------------------------------------------------------------
 # ---  Load in the validation data ---

### 2025-02-20 src/scripts/train_torch_chempy.py
[showyourwork] overleaf sync

@@ -13,7 +13,7 @@ import time as t
 import schedulefree
 
 # ---  You want to retrain? ---
-re_train = False #True
+re_train = True # False
 
 # ----- Load the data ---------------------------------------------------------------------------------------------------------------------------------------------
 # ---  Load in the validation data ---

### 2025-02-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-The key assumption underpinning our methodology — that individual stars are independently sampled from their respective stellar environments — is foundational for tractability.
+The key assumption underpinning our methodology -- that individual stars are independently sampled from their respective stellar environments -- is foundational for tractability.
 However, this assumption warrants closer examination.
@@
 This efficiency makes SBI uniquely suited for analyzing the massive datasets expected from next-generation spectroscopic surveys.
-    \item Scalability: SBI’s amortized nature allows it to scale seamlessly with the size of the dataset.
+    \item Scalability: SBI's amortized nature allows it to scale seamlessly with the size of the dataset.
 By training a neural posterior estimator once, the method can be applied repeatedly at virtually no additional computational cost.
-This scalability is essential for leveraging the millions of stars that future surveys like 4MOST and WEAVE will provide, enabling precise population-level inferences.
+This scalability is essential for leveraging the millions of stars that future surveys like 4MOST \citep{4most} and WEAVE \citep{weave} will provide, enabling precise population-level inferences.
     \item Robustness to model misspecifications: Unlike HMC, which shows significant biases when faced with discrepancies between model assumptions and data, SBI demonstrates remarkable robustness.
 \textcolor{blue}{Even under conditions of mismatched yield tables (should be updated)} or data generated from hydrodynamical simulations, SBI provides accurate and reliable results.
-This robustness ensures SBI’s applicability in real-world scenarios where exact model fidelity cannot be guaranteed.
+This robustness ensures SBI's applicability in real-world scenarios where exact model fidelity cannot be guaranteed.
 \end{itemize}

### 2025-02-23 src/scripts/train_sbi.py

@@ -70,7 +70,7 @@ check_sbi_inputs(simulator, prior)
 
 
 # ----- Train the SBI -------------------------------------------------------------------------------------------------------------------------------------------
-density_estimator_build_fun = posterior_nn(model="maf", hidden_features=5, num_transforms=2)#, blocks=1)
+density_estimator_build_fun = posterior_nn(model="maf", hidden_features=6, num_transforms=3)#, blocks=1)
 inference = NPE_C(prior=prior, density_estimator=density_estimator_build_fun, show_progress_bars=True)
 
 start = t.time()

### 2025-02-25 src/tex/ms.tex
[showyourwork] overleaf sync

@@
   % results heading (mandatory)
-   {SBI achieves a $\sim1,000\times$ speed-up compared to HMC, reducing inference runtime from $\gtrsim40$ hours to mere minutes for thousands of stars.
-Inference on $1,000$ stars yields precise estimates for the IMF slope ($\alpha_{\rm IMF} = -2.296 \pm 0.009$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = -2.889 \pm 0.010$), deviating less than 0.04\% from the ground truth.
-SBI also demonstrates greater robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation.}
+   {SBI achieves a $\sim75,600\times$ speed-up compared to HMC, reducing inference runtime from $\gtrsim42$ hours to mere seconds for thousands of stars.
+Inference on $1,000$ stars yields precise estimates for the IMF slope ($\alpha_{\rm IMF} = -2.298 \pm 0.002$) and SN\,Ia normalization ($\log_{10}(N_{\rm Ia}) = -2.885 \pm 0.003$), deviating less than 0.05\% from the ground truth.
+SBI also demonstrates similar robustness to model misspecification than HMC, recovering accurate parameters even with alternate yield tables or data from a cosmological simulation.}
   % conclusions heading (optional), leave it empty if necessary 
@@
 With the NN emulator, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{CHEMPY}.
+Having access to a fast simulator opens the possibility for testing Sequential Neural Posterior Estimate (SNPE) as an alternative, but we discuss this possibility in Section \ref{sec: discussion}.
     \item \textbf{Differentiability:} The NN is written in pytorch and has additionally a simple closed-form analytic structure \citep[described in the appendix of][]{Philcox_2019}, unlike the complex \texttt{CHEMPY} model.
@@
 With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of %\variable{output/ape_NN_log.txt} dex)
-which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN})\footnote{Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.}.
+which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN})\footnote{Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14925307}{https://zenodo.org/records/14925307}.}.
 In fact, we will add additional observational uncertainty to our mock observational data later during training of the neural posterior estimator network.
@@
 In order to determine the optimal values of the global galactic parameters ($\Vec{\Lambda}$) one has to sample the posterior of Eq.~\ref{eq:posterior}.
-In practice this is a costly computation, since even with advanced techniques such as HMC sampling the posterior can only be evaluated for a small set of stars ($\lesssim200$) and requires long compute times \citep[$\sim40$ hours][]{Philcox_2019}.
+In practice this is a costly computation, since even with advanced techniques such as HMC sampling the posterior can only be evaluated for a small set of stars ($\lesssim200$) and requires long compute times \citep[$\sim42$ hours][]{Philcox_2019}.
 %
@@
 
-We use a NPE \citep{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Neural Spline Flow \citep[NSF;][]{durkan2019neuralsplineflows} containing 20 hidden features and 10 transformation layers for the normalizing flow.
-The NSF uses as invertible transformation smooth and flexible spline functions to allow efficient sampling.
-The expressivity of the spline allows the NDE to be capture complex distributions, while also maintaining computational tractability.
-The final model was selected after extensive hyperparameter tuning, varying: the architecture between NSF, Masked Autoregressive Flow \citep[MAF;][]{papamakarios2018maskedautoregressiveflowdensity}, and Mask Autoregressive
-Flow with rational-quadratic spline (MAF-RQS), the number of neurons between {10, 20, 50, 100}, and the number of transformations between {1, 5, 10, 30}, optimizing over the test set for both the highest mean log posterior probability and the best calibration as measured by the TARP value \citep{Lemos2023}.
+We use a NPE \citep{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow 
+\citep[MAF;][]{papamakarios2018maskedautoregressiveflowdensity} 
+containing 8 hidden features and 4 transformation layers for the normalizing flow. 
+%The NSF uses as invertible transformation smooth and flexible spline functions to allow efficient sampling. 
+The expressivity of the MAF allows the NDE to be capture complex distributions, while also maintaining computational tractability.
+The final model was selected after extensive hyperparameter tuning, varying: the architecture between Neural Spline Flow \citep[NSF;][]{durkan2019neuralsplineflows}, MAF, and MAF with rational-quadratic spline (MAF-RQS), the number of neurons between {10, 20, 50, 100}, and the number of transformations between {1, 5, 10, 30}, optimizing over the test set for both the highest mean log posterior probability and the best calibration as measured by the TARP value \citep{Lemos2023}.
 For more details on the TARP value see section \ref{sec:sbc} in the appendix.
@@
 We train our NPE with $10^5$ data points consisting of $n_{\rm elements}=8$ simulated with the NN emulator described in the previous section.
-Inputs are sampled from a uniform prior over the original $5\sigma$-range of the Gaussian priors shown in Tab.
-\ref{tab:priors} to provide better coverage of the parameter space.
-Training takes $\sim35$ minutes on an Apple M1 Max.
-We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator.
+Inputs are sampled from the Gaussian priors shown in Tab.
+\ref{tab:priors}.
+Training takes $\sim10$ minutes on an Apple M1 Max.
+We evaluate the accuracy of the NPE using $5\times10^3$ validation data points from the original \texttt{CHEMPY} simulator.
 In order to mimic observational noise, we add 5\% observational uncertainties to the abundances simulated with \texttt{CHEMPY} before feeding them to our NPE.
@@
 This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars.
-Our method takes around $50$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
-Our combined runtime for the NPE training plus sampling is then $2160$ seconds on an Apple M1 Max for 1,000 stars hence our methods making it more than $340$ times faster than current HMC methods which take around $40$h for only $200$ stars \citep{Philcox_2019}.
-But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim14,400$.
+Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1,000 stars (each time sampling the single star posterior with 1,000 points and fitting for the Gaussian parameters).
+Our combined runtime for the NPE training plus sampling is then $610$ seconds on an Apple M1 Max for 1,000 stars hence our methods making it more than $1240$ times faster than current HMC methods which take around $42$h for only $200$ stars \citep{Philcox_2019}.
+Note though, that since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim75,600$.
 Hence, shorter computing times make it feasible to use orders of magnitudes more observations.
@@
 However, currently our NPE network is not particularly good at estimating ages from abundances alone.
-On average our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data.
+On average our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data but well in agreement with current uncertainties of stellar age inference.
 
@@
 Using a sample of 1,000 stars we infer \variable{output/CHEMPY_TNG_sbi.txt} which deviates less than $\sim0.04\%$ from the ground truth value.
+We have further checked the accuracy of our inference for a vastly different mock observational dataset created with shifted parameters of $\alpha_{IMF}=-2.1$ and $\log_{10}(N_{Ia})=-3$ and found that also in this case our model is well able to lead to correct inferences (see sec.~\ref{sec: additional inference} in the appendix).
 
@@
          SN\,Ia & \citet{2003NuPhA.718..139T} \\
-         %SN\,II & \citet{2013ARA&A..51..457N} \\
+         SN\,II & \citet{Nomoto2013} \\
          AGB & \citet{2016ApJ...825...26K}
@@
 
-This isn't up to date
-\textcolor{blue}{
 In Fig.~\ref{fig:CHEMPY_alt_N_star_analysis} we show our inference results (blue) for this setup for varying number of stars and compare them against HMC results (red) from \citet{Philcox_2019}.
-We see that the SBI results are less biased than the HMC results, especially for $\alpha_{\rm IMF}$.
-In fact with 100 stars HMC inferences are about \textcolor{red}{calculate value} $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively.
+We see that the SBI results are similar biased as the HMC results.
+In fact with 100 stars HMC inferences are about 4 and 3 $\sigma$ away from the ground truth value for $\alpha_{\rm IMF}$ and $\log_{10}(N_{\rm Ia})$, respectively.
 Again, the joint posterior is shown in Fig.~\ref{fig:CHEMPY_alt_sbi} and shows that while individually parameter inferences look good, jointly taken the SBI results are on the edge of being $3\sigma$ away from the ground truth.
-}
 
-\textcolor{blue}{
 Nevertheless, the performance of our inference is still very good.
@@
 However, we also note that our inference is now slightly biased as the observational data does not match the training data.
-Looking at Fig.~\ref{fig:CHEMPY_alt_sbi} we see that our inference is still consistent with the ground truth within the $3\sigma$ levels.
-Comparing this to the HMC results (red band in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis}), we see that SBI is performing much better when the model is misspecified and leads to significantly less biased results.
-This is -- next to drastically reduced compute times -- another a key advantage of our SBI method compared to more standard approaches such as HMC.
-}
+Looking at Fig.~\ref{fig:CHEMPY_alt_sbi} we see that our inference is inconsistent with the ground truth within several $\sigma$ levels.
+Comparing this to the HMC results (red band in Fig.~\ref{fig:CHEMPY_alt_N_star_analysis}), we see that SBI is performing similarly bad as HMC when the model is misspecified.
+Part of the problem here is that through our assumption of a factorized posterior, we decrease the inference uncertainty as we increase the number of observations.
+In future we will improve upon this through different model architectures.
+In summary, the drastically reduced compute times offer a key advantage of our SBI method compared to more standard approaches such as HMC.
+Nevertheless, in the limit of large star counts our current approach becomes over-confident and model uncertainities are underestimated.
 
@@
 Despite the drastic difference in chemical enrichment model between training and testing data, our SBI pipeline is impressively capable of inferring the correct posterior values.
-SBI results for $\alpha_{\rm IMF}$ are almost perfect while $\log_{10}(N_{\rm Ia})$ is slightly biased low.
-This is exactly opposite to the results of HMC where inference for $\alpha_{\rm IMF}$ is biased and results for $\log_{10}(N_{\rm Ia})$ are more in agreement.
+SBI results for  $\log_{10}(N_{\rm Ia})$ are almost perfect while $\alpha_{\rm IMF}$ is slightly biased high.
+This is very similar to the results of HMC where inference for $\alpha_{\rm IMF}$ is biased and results for $\log_{10}(N_{\rm Ia})$ are more in agreement.
 Hence, our SBI inference is on par with the HMC results and given their extreme computational advantage they actually supersede HMC.
 
-Looking at the joint posteriors in Figs.~\ref{fig:CHEMPY_TNG_sbi}, \ref{fig:CHEMPY_alt_sbi} and \ref{fig:CHEMPY_TNG_sim_sbi} we see that also in this case SBI recovers the true parameters with less than $3\sigma$ bias.
+Looking at the joint posteriors in Figs.~\ref{fig:CHEMPY_TNG_sbi}, \ref{fig:CHEMPY_alt_sbi} and \ref{fig:CHEMPY_TNG_sim_sbi} we see that also in this case SBI recovers the true parameters only with a few $\sigma$ bias similar to HMC.
+Again, part of this problem is our assumpotion of a factorized posterior that underestimated model uncertainties as for the alternative yield case.
 
@@
 Future efforts should explore methodologies capable of incorporating such correlations to further refine the accuracy of SBI, potentially through hierarchical modeling or graph-based methods.
+In particular this assumption leads to over-confident models in the limit of large observational samples that drastically underestimate model uncertainties.
+One very promising avenue to circumvent this probelm will be compositional score models \citep{compositional_score} which we will explore in future work.
 
-An exciting avenue for future research involves expanding SBI to directly infer empirical nucleosynthetic yields, which remain a major uncertainty in GCE models.
+An exciting avenue for future research involves expanding the SBI method presented here to be able to cope with missing data points by switching to a transformer model architecture.
+Also, with slight modifications, SBI could be used 
+to directly infer empirical nucleosynthetic yields, which remain a major uncertainty in GCE models.
 In the current framework, \texttt{CHEMPY} relies on tabulated yields from theoretical studies, which may not fully represent the complex processes driving stellar enrichment.
@@
 Such an approach could provide a unified framework for calibrating galactic models directly against observational data.
+Research in this direction is currently performed and will be part of a future paper.
+
+In the meantime, SBI can also be used to perform model comparison \citep[e.g.][]{model_comp,zhou2024} which can be used to determine which of the tabulated yield sets best match observational constraints.
 
@@
 Our findings indicate that SBI can seamlessly scale to such large datasets, offering significant advantages in terms of speed and computational cost.
+
+As anticipated in Section \ref{sec:methods}, the speed of our NN emulator would allow us to perform SNPE, giving up the amortized property in favor of a stronger constraining power, since the Sequential techniques have been shown empirically to outperform the respective amortized version \citep{Ho2024}.
+This method trains the NDE on a fraction of the initial simulations budget, retrieving a first estimate of the posterior, and at inference time simulate new observations to train on in regions of high posterior density, obtained from the first estimate.
+In this way we can obtain posteriors that are deliberately optimized for a singular data point.
+We have decided to leave this approach for future work because the results are quite promising and the amortized property can be crucial for the scalability.
 
@@
 \begin{itemize}
-    \item Orders-of-magnitude speed-up: SBI dramatically reduces computational requirements, achieving runtime improvements exceeding 1,000-fold compared to HMC.
-For instance, while HMC requires $\sim 40\,\mathrm{hours}$ to infer parameters from just 200 stars, SBI completes inference on thousands of stars in mere minutes.
+    \item Orders-of-magnitude speed-up: SBI dramatically reduces computational requirements, achieving runtime improvements exceeding 75,000-fold compared to HMC.
+For instance, while HMC requires $\sim 42\,\mathrm{hours}$ to infer parameters from just 200 stars, SBI completes inference on thousands of stars in mere minutes.
 This efficiency makes SBI uniquely suited for analyzing the massive datasets expected from next-generation spectroscopic surveys.
@@
     \item Robustness to model misspecifications: Unlike HMC, which shows significant biases when faced with discrepancies between model assumptions and data, SBI demonstrates remarkable robustness.
-\textcolor{blue}{Even under conditions of mismatched yield tables (should be updated)} or data generated from hydrodynamical simulations, SBI provides accurate and reliable results.
+Even under conditions of mismatched yield tables or data generated from hydrodynamical simulations, SBI provides accurate and reliable results.
 This robustness ensures SBI's applicability in real-world scenarios where exact model fidelity cannot be guaranteed.
@@
 These enhancements will further solidify SBI as a cornerstone method in galactic archaeology.
+
+We note though, that the SBI pipeline presented here is not perfect and suffers from simplified assumptions that have been made in order to make the problem tracktable.
+In particular, the assumption that the  posterior factorizes is a strong assumption that leads to over-confidence of the models at large observational samples.
+This is certainly something that needs to be improved in future iterations and methodological work in this direction is currently undergoing.
 
@@
 \section{Additional inference results}
+\label{sec: additional inference}
 
@@
     \vspace{-.5cm}
-    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for mock data a created with parameters for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$.}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_N_star_analysis} but for mock data created with different parameters for $\alpha_{IMF}=-2.1$ and $\log_{10}(N_{Ia})=-3.0$.}
     \label{fig:N_star_analysis_different_prior}
@@
 \end{figure*}
+
+\begin{figure}
+%\vspace{-1.3cm}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/different_prior.png}
+    \vspace{-.25cm}
+    \caption{Same as Fig.~\ref{fig:CHEMPY_TNG_sim_sbi} but for the mock data taken created with a different parameter combination for $\alpha_{IMF}=-2.1$ and $\log_{10}(N_{Ia})=-3.0$.}
+    \label{fig:CHEMPY_different_prior_sbi} 
+    \script{additional_inference.py}
+\end{figure}
 

### 2025-02-25 src/scripts/chempy_tng_inference.py

@@ -41,7 +41,6 @@ gaussian_posterior_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHE
 #stars = np.arange(1,1000)
 #n_stars_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, "CHEMPY_TNG_yields_N_star", stars)
 
-
 # --- Compare to HMC ---
 # Philcox&Rybizki 2019 Table 3
 philcox = {}
@@ -49,6 +48,8 @@ philcox["n_stars"] = np.array([1,10,100])
 philcox["med"] = np.array([[-2.29,-2.87],[-2.31,-2.90],[-2.31,-2.90]])
 philcox["up"] = np.array([[-2.21,-2.76],[-2.29,-2.87],[-2.30,-2.89]])
 philcox["lo"] = np.array([[-2.37,-2.98],[-2.33,-2.93],[-2.32,-2.91]])
+
+gaussian_posterior_plot_n_stars(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields_philcox", philcox, n_stars=100)
 
 stars = np.arange(1,1000)
 n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_Nstar_comp", stars)

### 2025-02-25 src/scripts/plot_functions.py

@@ -364,10 +364,12 @@ def n_stars_plot_comp2(x1, x2, x_true, philcox, save_name, no_stars = np.array([
         ax.tick_params(labelsize=30, size=10, width=3)
         ax.tick_params(which='minor', size=5, width=2)
 
-        print(philcox['med'][:,i])
+        two_sigma_h = philcox['med'][:,i] + 2*(philcox['up'][:,i] - philcox['med'][:,i])
+        two_sigma_l = philcox['med'][:,i] - 2*(philcox['med'][:,i] - philcox['lo'][:,i])
         # Add Philcox
         ax.plot(philcox['n_stars'],philcox['med'][:,i],c='r', label="HMC")
         ax.fill_between(philcox['n_stars'],philcox['lo'][:,i],philcox['up'][:,i],alpha=0.2,color='r')
+        ax.fill_between(philcox['n_stars'],two_sigma_l,two_sigma_h,alpha=0.2,color='r')
 
     for i, name in enumerate([r'$\alpha_{\rm IMF}$', r'$\log_{10} N_{\rm Ia}$']):
         plot(fit[:,i], err[:,i], x_true[0,i], ax[i], name)
@@ -443,6 +445,115 @@ def gaussian_posterior_plot(alpha_IMF, log10_N_Ia, global_params, title):
     x, y = np.mgrid[grid_x[0]:grid_x[1]:0.001, grid_y[0]:grid_y[1]:0.001]
     pos = np.dstack((x, y))
 
+    # create a multivariate normal
+    posterior = multivariate_normal(mean=[mu_alpha,mu_log10N_Ia], cov=[[sigma_alpha**2,0],[0,sigma_log10N_Ia**2]])
+    samples = posterior.rvs(size=100_000_000)
+
+    # create a figure
+    plt.figure(figsize=(15,15))
+    
+    plt.hist2d(samples[:,0], samples[:,1], bins=250, range=[grid_x, grid_y], cmin=1, norm="log")
+
+    # labels
+    label_gt = r'Ground Truth' + f"\n" + r"$\alpha_{\rm IMF} = $" + f'${global_params[0,0]:.2f}$' + f"\n" + r"$\log_{10} N_{\rm Ia} = $" + f'${global_params[0,1]:.2f}$'
[…]

### 2025-02-25 src/scripts/chempy_tng_inference.py
[showyourwork] overleaf sync

@@ -49,7 +49,7 @@ philcox["med"] = np.array([[-2.29,-2.87],[-2.31,-2.90],[-2.31,-2.90]])
 philcox["up"] = np.array([[-2.21,-2.76],[-2.29,-2.87],[-2.30,-2.89]])
 philcox["lo"] = np.array([[-2.37,-2.98],[-2.33,-2.93],[-2.32,-2.91]])
 
-gaussian_posterior_plot_n_stars(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields_philcox", philcox, n_stars=100)
+gaussian_posterior_plot_n_stars(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields_philcox", philcox=philcox, n_stars=100)
 
 stars = np.arange(1,1000)
 n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_Nstar_comp", stars)

### 2025-02-25 src/scripts/plot_functions.py

@@ -350,6 +350,13 @@ def n_stars_plot_comp2(x1, x2, x_true, philcox, save_name, no_stars = np.array([
     fig , ax = plt.subplots(nrows=1, ncols=2, figsize=(26,6))
 
     def plot(fit, err, x_true, ax, name):
+        two_sigma_h = philcox['med'][:,i] + 2*(philcox['up'][:,i] - philcox['med'][:,i])
+        two_sigma_l = philcox['med'][:,i] - 2*(philcox['med'][:,i] - philcox['lo'][:,i])
+        # Add Philcox
+        ax.plot(philcox['n_stars'],philcox['med'][:,i],c='r', label="HMC")
+        ax.fill_between(philcox['n_stars'],philcox['lo'][:,i],philcox['up'][:,i],alpha=0.1,color='r')
+        ax.fill_between(philcox['n_stars'],two_sigma_l,two_sigma_h,alpha=0.1,color='r')
+        
         ax.plot(no_stars, fit, color="b", label="Fit")
         ax.fill_between(no_stars, fit-err, fit+err, alpha=0.1, color="b", label=r"1 & 2 $\sigma$")
         ax.fill_between(no_stars, fit-2*err, fit+2*err, alpha=0.1, color="b")
@@ -363,13 +370,6 @@ def n_stars_plot_comp2(x1, x2, x_true, philcox, save_name, no_stars = np.array([
         ax.set_xlim([1,1000])
         ax.tick_params(labelsize=30, size=10, width=3)
         ax.tick_params(which='minor', size=5, width=2)
-
-        two_sigma_h = philcox['med'][:,i] + 2*(philcox['up'][:,i] - philcox['med'][:,i])
-        two_sigma_l = philcox['med'][:,i] - 2*(philcox['med'][:,i] - philcox['lo'][:,i])
-        # Add Philcox
-        ax.plot(philcox['n_stars'],philcox['med'][:,i],c='r', label="HMC")
-        ax.fill_between(philcox['n_stars'],philcox['lo'][:,i],philcox['up'][:,i],alpha=0.2,color='r')
-        ax.fill_between(philcox['n_stars'],two_sigma_l,two_sigma_h,alpha=0.2,color='r')
 
     for i, name in enumerate([r'$\alpha_{\rm IMF}$', r'$\log_{10} N_{\rm Ia}$']):
         plot(fit[:,i], err[:,i], x_true[0,i], ax[i], name)

### 2025-02-26 src/scripts/additional_inference.py
[showyourwork] overleaf sync

@@ -96,7 +96,7 @@ log10_N_Ia_obs = np.array(log10_N_Ia_obs)
 gaussian_posterior_plot_n_stars(alpha_IMF_obs, log10_N_Ia_obs, global_params_plot, title="different_prior_philcox", philcox=None, no_stars=100)
 
 # --- plot the data ---
-gaussian_posterior_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params_plot, title="different_prior")
+#gaussian_posterior_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params_plot, title="different_prior")
 
 stars = np.arange(1,1000)
 n_stars_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params_plot, "Nstar_comp_different_prior", stars)

### 2025-02-26 src/scripts/chempy_alt_yield_inference.py

@@ -93,4 +93,4 @@ gaussian_posterior_plot_n_stars(alpha_IMF_alt_obs, log10_N_Ia_alt_obs, global_pa
 stars = np.arange(1,len(alt_y_obs))
 n_stars_plot_comp2(alpha_IMF_alt_obs, log10_N_Ia_alt_obs, global_params, philcox, "CHEMPY_alternative_yields_Nstar_comp", stars)
 
-gaussian_posterior_plot(alpha_IMF_alt_obs, log10_N_Ia_alt_obs, global_params, title="CHEMPY_alternative_yields")
+#gaussian_posterior_plot(alpha_IMF_alt_obs, log10_N_Ia_alt_obs, global_params, title="CHEMPY_alternative_yields")

### 2025-02-26 src/scripts/chempy_tng_inference.py

@@ -37,7 +37,7 @@ print(f"log10_N_Ia = {mu_N_Ia_combined:.3f} +/- {sigma_N_Ia_combined:.3f}")
 
 
 # --- Plot the data ---
-gaussian_posterior_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields")
+#gaussian_posterior_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields")
 #stars = np.arange(1,1000)
 #n_stars_plot(alpha_IMF_obs, log10_N_Ia_obs, global_params, "CHEMPY_TNG_yields_N_star", stars)
 

### 2025-02-26 src/scripts/plot_functions.py

@@ -566,7 +566,7 @@ def gaussian_posterior_plot_n_stars(alpha_IMF, log10_N_Ia, global_params, title,
     if philcox is not None:
         # add Philcox ellipses
 
-        legend_philcox = plt.scatter(philcox['med'][:,0][-1], philcox['med'][:,1][-1], marker='s', color='gray', label=label_gt, s=50)
+        legend_philcox = plt.scatter(philcox['med'][:,0][-1], philcox['med'][:,1][-1], marker='s', color='gray', label="HMC", s=50)
 
         sigma_h = (philcox['up'][:,0] - philcox['med'][:,0])
         sigma_l = (philcox['med'][:,0] - philcox['lo'][:,0])
@@ -595,6 +595,7 @@ def gaussian_posterior_plot_n_stars(alpha_IMF, log10_N_Ia, global_params, title,
         #    i = np.abs(np.array(levels) - float(t._text)).argmin()
         #    s = int(sigma[i])
         #    t.set(text=f'{s} $\\sigma$')
+        legend_philcox = plt.legend(handles=[legend_philcox], fontsize=30, shadow=True, fancybox=True, loc=2, bbox_to_anchor=(0, 0.975))
 
     plt.xlabel(r'$\alpha_{\rm IMF}$', fontsize=40)
     plt.ylabel(r'$\log_{10} N_{\rm Ia}$', fontsize=40)
@@ -602,6 +603,7 @@ def gaussian_posterior_plot_n_stars(alpha_IMF, log10_N_Ia, global_params, title,
 
     plt.gca().add_artist(legend_fit)
     plt.gca().add_artist(legend_true)
+    plt.gca().add_artist(legend_philcox)
     
     plt.title(title, fontsize=40)
 

### 2025-02-26 src/scripts/tng_inference.py

@@ -85,4 +85,4 @@ gaussian_posterior_plot_n_stars(alpha_IMF_tng_obs, log10_N_Ia_tng_obs, global_pa
 stars = np.arange(1,len(tng_y_obs))
 n_stars_plot_comp2(alpha_IMF_tng_obs, log10_N_Ia_tng_obs, global_params, philcox, "TNG_sim_Nstar_comp", stars)
 
-gaussian_posterior_plot(alpha_IMF_tng_obs, log10_N_Ia_tng_obs, global_params, title="TNG_simulation")
+#gaussian_posterior_plot(alpha_IMF_tng_obs, log10_N_Ia_tng_obs, global_params, title="TNG_simulation")

### 2025-02-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \vspace{-.25cm}
-    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1,000$ stars.
-The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.
+    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $100$ stars.
+The ground truth value is shown by the red star, the posterior mean of the SBI inference is shown with the black dot and white ellipses show the $1-3\sigma$ contours of our inference.
+The gray ellipses show the results for HMC inference performed by \citet{Philcox_2019}.
 See Sec.~\ref{subsec:chempy_tng} for a full description.}

### 2025-02-26 src/scripts/chempy_alt_yield_inference.py

@@ -88,7 +88,7 @@ philcox["med"] = np.array([[-2.25,-3.01],[-2.21,-2.96],[-2.22,-2.96]])
 philcox["up"] = np.array([[-2.14,-2.86],[-2.17,-2.88],[-2.20,-2.93]])
 philcox["lo"] = np.array([[-2.34,-3.16],[-2.25,-3.04],[-2.24,-2.98]])
 
-gaussian_posterior_plot_n_stars(alpha_IMF_alt_obs, log10_N_Ia_alt_obs, global_params, title="CHEMPY_alternative_yields_philcox", philcox=philcox, no_stars=100)
+gaussian_posterior_plot_n_stars(alpha_IMF_alt_obs, log10_N_Ia_alt_obs, global_params, title="CHEMPY_alternative_yields", philcox=philcox, no_stars=100)
 
 stars = np.arange(1,len(alt_y_obs))
 n_stars_plot_comp2(alpha_IMF_alt_obs, log10_N_Ia_alt_obs, global_params, philcox, "CHEMPY_alternative_yields_Nstar_comp", stars)

### 2025-02-26 src/scripts/chempy_tng_inference.py

@@ -49,7 +49,7 @@ philcox["med"] = np.array([[-2.29,-2.87],[-2.31,-2.90],[-2.31,-2.90]])
 philcox["up"] = np.array([[-2.21,-2.76],[-2.29,-2.87],[-2.30,-2.89]])
 philcox["lo"] = np.array([[-2.37,-2.98],[-2.33,-2.93],[-2.32,-2.91]])
 
-gaussian_posterior_plot_n_stars(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields_philcox", philcox=philcox, no_stars=100)
+gaussian_posterior_plot_n_stars(alpha_IMF_obs, log10_N_Ia_obs, global_params, title="CHEMPY_TNG_yields", philcox=philcox, no_stars=100)
 
 stars = np.arange(1,1000)
 n_stars_plot_comp2(alpha_IMF_obs, log10_N_Ia_obs, global_params, philcox, "CHEMPY_TNG_yields_Nstar_comp", stars)

### 2025-02-26 src/scripts/tng_inference.py

@@ -80,7 +80,7 @@ philcox["med"] = np.array([[-2.27,-2.86],[-2.27,-2.87],[-2.28,-2.89]])
 philcox["up"] = np.array([[-2.19,-2.75],[-2.24,-2.84],[-2.27,-2.88]])
 philcox["lo"] = np.array([[-2.35,-2.97],[-2.3,-2.91],[-2.29,-2.90]])
 
-gaussian_posterior_plot_n_stars(alpha_IMF_tng_obs, log10_N_Ia_tng_obs, global_params, title="TNG_simulation_philcox", philcox=philcox, no_stars=100)
+gaussian_posterior_plot_n_stars(alpha_IMF_tng_obs, log10_N_Ia_tng_obs, global_params, title="TNG_simulation", philcox=philcox, no_stars=100)
 
 stars = np.arange(1,len(tng_y_obs))
 n_stars_plot_comp2(alpha_IMF_tng_obs, log10_N_Ia_tng_obs, global_params, philcox, "TNG_sim_Nstar_comp", stars)

### 2025-02-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 However, currently our NPE network is not particularly good at estimating ages from abundances alone.
-On average our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data but well in agreement with current uncertainties of stellar age inference.
+On average our inference for ages results in an APE of $\sim21\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data but well in agreement with current uncertainties of stellar age inference that range from 15\% to 30\% for turn-off stars and giants respectively.
 

### 2025-02-26 src/scripts/plot_sbc.py

@@ -113,3 +113,27 @@ for i in tqdm(range(len(abundances))):
 
 percentile = np.percentile(correlation, [0, 99])
 percentile_mask = (correlation>percentile[0])&(correlation<percentile[1])
+
+fig = plt.figure(figsize=(35, 5))
+ax = fig.add_subplot(131)
+ax.hist(correlation[percentile_mask], histtype='step', color='black')
+ax.set_xlabel(r'$C_{ \alpha_{\text{IMF}, \log_{10}{N_{\text{Ia}} }} }$', fontsize=32 )  
+ax.set_ylabel('Counts', fontsize=32 )
+
+ax = fig.add_subplot(132)
+ax.scatter(distance[percentile_mask], correlation[percentile_mask], s=1, color='black')
+ax.set_xscale('log')
+ax.set_yscale('log')
+ax.set_xlabel(r'$|| \theta_\text{true} - \theta_{\text{predicted}} ||$', fontsize=32)
+ax.set_ylabel(r'$C_{ \alpha_{\text{IMF}, \log_{10}{N_{\text{Ia}} }} }$', fontsize=32)  
+ax.grid(True, color='gray', linestyle='--', linewidth=0.5)
+
+ax = fig.add_subplot(133)
+ax.scatter(mahalanobis_distance[percentile_mask], correlation[percentile_mask], s=1, color='black')
+ax.set_xscale('log')
+ax.set_yscale('log')
+ax.set_xlabel(r'$|| \theta_\text{true} - \theta_{\text{predicted}} ||_{\text{Mahalanobis}}$', fontsize=32)
+ax.set_ylabel(r'$C_{ \alpha_{\text{IMF}, \log_{10}{N_{\text{Ia}} }} }$', fontsize=32)  
+ax.grid(True, color='gray', linestyle='--', linewidth=0.5)
+
+fig.savefig(paths.figures / 'correlation_high_mass_slope_SN_delay.pdf', bbox_inches='tight')

### 2025-02-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \centering
+    \includegraphics[width=\textwidth]{figures/corner_plot_singlestar.pdf}
+    \vspace{-.5cm}
+    \caption{Corner plot of the posteriors for all six parameters for a single star from the validation set.
+The gray contours show a kde-estimate of the posterior from our SBI inference and the red dot and lines show the ground truth parameter values.
+Gray histograms on the diagonal show a kde estimate of the marginals.}
+    \label{fig:corner_plot}
+    \script{plot_sbc.py}
+\end{figure*}
+
+\begin{figure*}
+    \centering
     \includegraphics[width=\textwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
@@
 There we can also see, that the accuracy for an individual star of the NPE is not particularly high.
-However, for the global parameters $\vec\Lambda$ we can boost the accuracy by combining the inference for many stars of the same galaxy.
 
@@
 On average our inference for ages results in an APE of $\sim21\%$ which is slightly larger than the observational noise of $20\%$ that we add during the mock up of our data but well in agreement with current uncertainties of stellar age inference that range from 15\% to 30\% for turn-off stars and giants respectively.
+
+We accompany the APE analysis by showing the full posterior inference results for a single star in Fig.~\ref{fig:corner_plot}.
+This figure shows that the SBI approach is well able to infer correct parameters and their cross-correlations.
+In partciular, there is a strong correlation between the two global parameters as seen in the top left corner as well as for time (or stellar age) and all other five parameters as visible from the bottom row.
+
+Finally, as already mentioned in the method section for the global parameters $\vec\Lambda$ we can boost the accuracy by combining the inference for many stars of the same galaxy.
 
@@
 
+\section{}
+\label{sec: correlations}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/correlation_high_mass_slope_SN_delay.pdf}
+    \vspace{-.5cm}
+    \caption{.}
+    \label{fig:correlation}
+    \script{plot_sbc.py}
+\end{figure*}
+
 \section{Additional inference results}

### 2025-02-26 src/scripts/plot_sbc.py

@@ -114,6 +114,16 @@ for i in tqdm(range(len(abundances))):
 percentile = np.percentile(correlation, [0, 99])
 percentile_mask = (correlation>percentile[0])&(correlation<percentile[1])
 
+plt.rcParams.update({
+    'font.size': 20,
+    'axes.titlesize': 20,
+    'axes.labelsize': 20,
+    'xtick.labelsize': 13,
+    'ytick.labelsize': 13,
+    'legend.fontsize': 15,
+    'figure.titlesize': 20
+})
+
 fig = plt.figure(figsize=(35, 5))
 ax = fig.add_subplot(131)
 ax.hist(correlation[percentile_mask], histtype='step', color='black')

### 2025-02-26 src/scripts/plot_sbc.py
[showyourwork] overleaf sync

@@ -118,8 +118,8 @@ plt.rcParams.update({
     'font.size': 20,
     'axes.titlesize': 20,
     'axes.labelsize': 20,
-    'xtick.labelsize': 13,
-    'ytick.labelsize': 13,
+    'xtick.labelsize': 20,
+    'ytick.labelsize': 20,
     'legend.fontsize': 15,
     'figure.titlesize': 20
 })

### 2025-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\section{}
+\section{Correlation analysis of inference results}
 \label{sec: correlations}
+In order to characterize the relation between the global parameters $\alpha_{\text{IMF}}$ and $\log_{10} N_{\text{Ia}}$, we have decided to study the correlation obtained from the inference results on the whole validation set.
+Figure \ref{fig:corner_plot} display a possible positive correlation, and in order to obtain a population level statics of this results, we calculate for each star in the validation set the 2x2 covariance matrix\footnote{We have used only the global parameters components of the sample to calculate the covariance matrix.} of the posterior samples, extracting the off diagonal value.
+In Figure \ref{fig:correlation}, the histogram of the correlation value is shown, and all the inference results suggest a positive correlation.
+We kept only the left sided 99 percentile for graphical reasons, but also those removed outliers are in agreement with the conclusion.
+We checked the independence of this results from the accuracy of the inference with the central and right plots of Fig. \ref{fig:correlation}.
 
@@
     \vspace{-.5cm}
-    \caption{.}
+    \caption{The left Figure shows the distribution of the correlation between the global parameters $\alpha_{\text{IMF}}$ and $\log_{10} N_{\text{Ia}}$, obtained by the covariance matrix of the sample of each star in the validation set.
+The central and right Figures show the correlation as a function of the Euclidean and Mahalanobis distance of the true value $\theta_{\text{true}}$ and the sample average $\theta_{\text{predicted}}$.
+This results shows that the correlation is independent on the accuracy of the inference.}
+
     \label{fig:correlation}

### 2025-02-27 src/scripts/plot_sbc.py

@@ -15,7 +15,15 @@ name = "NPE_C"
 # --- Define the prior ---
 a = ModelParameters()
 labels_out = a.elements_to_trace
-labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ["time"]
+# labels_in = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ["time"]
+labels_in = [
+    r"$\alpha_{\text{IMF}}$",
+    r"$\log_{10}{\text{N}_{\text{Ia}}}$",
+    r"$\log_{10}{\text{SFE}}$",
+    r"$\log_{10}{\text{SFR}_{\text{peak}}}$",
+    r"$x_{\text{out}}$",
+    r"$\text{Time}$",
+]
 
 # ----- load the posterior -------------------------------------------------------------------------------------------------------------------------------------------
 with open(paths.data / f"posterior_{name}.pickle", "rb") as f:
@@ -127,7 +135,7 @@ plt.rcParams.update(
     }
 )
 
-fig = plt.figure(figsize=(35, 5))
+fig = plt.figure(figsize=(37, 5))
 ax = fig.add_subplot(131)
 ax.hist(correlation[percentile_mask], histtype="step", color="black")
 ax.set_xlabel(r"$C_{ \alpha_{\text{IMF}, \log_{10}{N_{\text{Ia}} }} }$", fontsize=32)

### 2025-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/TobiBu/sbi-chempy}}} and refer to Appendix \ref{sec:appendix_code_and_data} for a more extended overview of our code availability.
-All our datasets and network weights are publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/records/14883024}}
+All our datasets and network weights are publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/records/14925307}}
 
@@
 The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{CHEMPY} to a single matrix computation.
-With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of %\variable{output/ape_NN_log.txt} dex)
+With the above network parameter choices, the NN predicts abundances with an absolute percentage error of \variable{output/ape_NN.txt} (which translates into a logarithmic error of \variable{output/ape_NN_log.txt} dex)
 which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN})\footnote{Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14925307}{https://zenodo.org/records/14925307}.}.
@@
 
-Note, the methods presented here can naturally be extended to any fast and flexible GCE model \citep[e.g.][and others]{Talbot1971,vice,omega2018}, not just \texttt{CHEMPY}.
+Note, the methods presented here can naturally be extended to any fast and flexible GCE model \citep[e.g.][and others]{Talbot1971,vice,omega2018}, not just \texttt{CHEMPY} and can even be used to infer vastly different galactic parameters such as accretion events from abundance distributions of stars \citep[e.g.][]{Viterbo2024}.
 
@@
 v0.4.1 allows anyone to re-build the entire manuscript including rerunning all analysis.
-The datasets and neural network weights are stored at \url{https://doi.org/10.5281/zenodo.7343715} \textcolor{red}{change to final zenodo dataset!}.
+The datasets and neural network weights are stored at \url{https://zenodo.org/records/14925307}.
+The training and validation data can be found on zenodo as well under this link: \url{https://zenodo.org/records/14507221}.
 

### 2025-02-27 src/scripts/evaluate_emulator.py

@@ -102,4 +102,4 @@ with open(paths.output / "ape_NN.txt", "w") as f:
     f.write(f"${p2:.1f}^{{+{p3-p2:.1f}}}_{{-{p2-p1:.1f}}}\,\%$%")
 
 with open(paths.output / "ape_NN_log.txt", "w") as f:
-    f.write(f"${np.log10(1+p2):.1f}\,\%$%")
+    f.write(f"${np.log10(1+p2/100.):.1f}\,$%")

### 2025-02-27 src/scripts/evaluate_emulator.py
[showyourwork] overleaf sync

@@ -102,4 +102,4 @@ with open(paths.output / "ape_NN.txt", "w") as f:
     f.write(f"${p2:.1f}^{{+{p3-p2:.1f}}}_{{-{p2-p1:.1f}}}\,\%$%")
 
 with open(paths.output / "ape_NN_log.txt", "w") as f:
-    f.write(f"${np.log10(1+p2/100.):.1f}\,$%")
+    f.write(f"${np.log10(1+p2/100.):.3f}\,$%")

### 2025-02-28 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In the left panel of Fig.~\ref{fig:sbc} we show the distribution of ranks (depicted in red) in each dimension.
-Highlighted in grey, you see the 99\% confidence interval of a uniform distribution given the number of samples provided.
-In plain english: for a uniform distribution, we would expect 1 out of 100 (red) bars to lie outside the grey area.
+In the top panel of Fig.~\ref{fig:tarp} we show the distribution of ranks (depicted in red) in each dimension.
+Highlighted with black lines, you see the 99\% confidence interval of a uniform distribution given the number of samples provided.
+In plain english: for a uniform distribution, we would expect 1 out of 100 (blue) bars to lie outside the grey area.
 This figure shows that overall our posteriors are decently calibrated.
-Only for the parameter $\log_{10}\left(\mathrm{SFE}\right)$ we see a single bar outside the 99\% confidence interval.
+Only for the parameter $\log_{10}\left(\mathrm{SFE}\right)$ and Time we see a slight bmiss-calibration.
 But most importantly for the parameters of interest here, $\log_{10}\left(N_\mathrm{Ia}\right)$ and $\alpha_\mathrm{IMF}$ we have a well calibrated posterior.
@@
 However, in contrast to SBC, TARP provides a necessary and sufficient condition for posterior accuracy, i.e., it can also detect inaccurate posterior estimators.
-In Fig.~\ref{fig:tarp} we show the result for our NPE in blue in comparison to the ideal line shown in black dashed style.
+In the middle row of Fig.~\ref{fig:tarp} we show the result for our NPE in blue in comparison to the ideal line shown in black dashed style.
 This figure clearly shows that our NPE is well calibrated.
@@
 Thus, in general, it is recommended using and interpreting TARP like SBC and complementing coverage checks with posterior predictive checks.
+
+Finally, the bottom row of Fig.~\ref{fig:tarp} shows the predicted vs.
+true parameter plots where blue dots show the average of the posterior samples and errorbars show the standard deviation of the samples vs their ground truth parameter.
+We find that for the global parameters we recover the 1:1 relation well while for the other parameters the agreement is tilted.
 

### 2025-05-28 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\U}[1]{\textcolor{black}{#1}}
+\newcommand{\resub}[1]{\textbf{\textcolor{blue}{#1}}}
 \usepackage{subfigure}
@@
 \subsection{Bayesian model}
+\label{subsec:bayesian-model}
 \texttt{CHEMPY} effectively is a Bayesian model for stellar abundances given a set of parameters $\{\Vec{\Lambda},\Vec{\Theta},T\}$ and \citet{Philcox_2019} extended the \texttt{CHEMPY} framework to be able to model multiple stellar data-points.
@@
 We evaluate the accuracy of the NPE using $5\times10^3$ validation data points from the original \texttt{CHEMPY} simulator.
-In order to mimic observational noise, we add 5\% observational uncertainties to the abundances simulated with \texttt{CHEMPY} before feeding them to our NPE.
+In order to mimic observational noise, \resub{we add 5\%  Gaussian observational uncertainties ($=0.05$ dex) to the abundances simulated with \texttt{CHEMPY}, representative of typical APOGEE measurements \citep{apogee17}, to simulate observational noise.
+While this uniform treatment simplifies the modeling, we acknowledge that in real data, uncertainties can vary by stellar type, element, or SNR.
+Future extensions of this work will incorporate more realistic, element-specific or SNR-dependent noise models.}
+
 We have made sure that our NPE is well calibrated and posterior distributions are trustable. For more details on simulation-based calibration see appendix \ref{sec:sbc}.
@@
 This does however present one issue, that since we only have access to samples from the posterior and not the posterior itself it is difficult to evaluate eq.~\ref{eq:posterior}.
-We circumvent this by approximating each single star posterior by a Gaussian and fit for the parameters of mean and covariance.
+\resub{We circumvent this by approximating each single-star posterior by a Gaussian and fitting for the parameters of mean and covariance.
+This choice enables an analytical combination of multiple posteriors (see Eq.~\ref{eq:posterior}), providing computational tractability and allowing efficient inference over large datasets.
+While this approximation may limit the flexibility of the posterior, it significantly reduces computational complexity.
+Alternative strategies, such as using a Mixture Density Network (MDN) or normalizing flows with parametric forms, would provide richer posterior representations.
+However, these would complicate the composition of posteriors across stars.
+We acknowledge this as a simplification and plan to explore more expressive architectures in future work.}
+
 With this it is straight forward to evaluate eq.~\ref{eq:posterior} analytically.
@@
 
-We accompany the APE analysis by showing the full posterior inference results for a single star in Fig.~\ref{fig:corner_plot}.
+We accompany the APE analysis by showing the full \resub{joint posterior over global and local parameters for a single star in Fig.~\ref{fig:corner_plot}.
+When performing inference across multiple stars, we marginalize over local parameters and use the factorized form in Eq.
+5 to approximate the global posterior.
+This distinction between single-star and population-level inference will be emphasized in the revised captions and text for clarity.}
 This figure shows that the SBI approach is well able to infer correct parameters and their cross-correlations.
@@
 We have further checked the accuracy of our inference for a vastly different mock observational dataset created with shifted parameters of $\alpha_{IMF}=-2.1$ and $\log_{10}(N_{Ia})=-3$ and found that also in this case our model is well able to lead to correct inferences (see sec.~\ref{sec: additional inference} in the appendix).
+
+\resub{It is important to note that our SBI results are derived using 1,000 stars, whereas the HMC results from \citet{Philcox_2019} are based on only 200 stars and involve slightly different HMC setups with additional parameters reflecting the individual element abundance weights or model errors contributing to the inference \citep[see section 4 of][]{Philcox_2019}.
+This contributes to the observed differences in posterior contours.
+Additionally, our method approximates posteriors using Gaussian fits, which may underestimate uncertainty at large sample sizes (see also Sec.~\ref{sec: discussion}).
+To ensure a fairer comparison, in appendix~\ref{sec:hmc-sbi} we have run a comparison between our SBI model and a custom HMC model reflecting the exact same setup as for the SBI.
+Figure~\ref{fig:} in the appendix shows that ...}
 
@@
 An exciting avenue for future research involves expanding the SBI method presented here to be able to cope with missing data points by switching to a transformer model architecture.
+\resub{A natural question arises regarding the use of these alternative architectures -- such as transformer-based models -- which could process variable numbers of stellar inputs directly, bypassing the need for single-star posterior approximations.
+These models could potentially capture correlations across stars and remove the need for assuming a factorized posterior.
+However, incorporating such models would require a new approach to network training and posterior inference.
+We are currently investigating these strategies and will report results in a forthcoming paper.}
+
 Also, with slight modifications, SBI could be used 
@@
 Our findings indicate that SBI can seamlessly scale to such large datasets, offering significant advantages in terms of speed and computational cost.
+\resub{While the present study focuses on mock datasets to validate the methodology under controlled conditions, we recognize the importance of testing our method on real survey data.
+Ongoing work is aimed at applying our pipeline to GALAH DR4 and APOGEE datasets.
+These surveys offer diverse chemical tracers and provide a valuable testbed for evaluating the robustness and calibration of our approach on real observations.
+This will be presented in future work.}
 
@@
 
+\resub{The calibration tests shown here are performed using posterior samples from our NPE trained on CHEMPY mock data.
+Validation is carried out on a hold-out set of simulated stars, with Gaussian noise added (0.05 dex), following the same protocol as described in Sec.~\ref{subsec:bayesian-model}.
+These tests evaluate the quality of posterior estimates for the global parameters  using single-star inference.}
+
 In the top panel of Fig.~\ref{fig:tarp} we show the distribution of ranks (depicted in red) in each dimension.
@@
 
+\section{\resub{Comparison between HMC and SBI}}
+\label{sec:hmc-sbi}
+\resub{put a new figure and some text...}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/hmc_corner_plot.pdf}
+    \vspace{-.5cm}
+    \caption{Caption}
+    \label{fig:hmc}
+    \script{hmc.py}
+\end{figure}
+
 \section{Additional inference results}

### 2025-05-28 src/scripts/hmc.py
[showyourwork] overleaf sync

@@ -1,0 +1,86 @@
+import corner
+import matplotlib.pyplot as plt
+import numpy as np
+import numpyro
+import numpyro.distributions as dist
+import torch
+from numpyro.infer import MCMC, NUTS
+
+import paths
+from chempy_torch_model import Model_Torch
+
+# ---- Load your emulator weights (pending file upload) ----
+model = Model_Torch(x_shape=6, y_shape=8)
+model.load_state_dict(torch.load(paths.data / "pytorch_state_dict.pt"))
+model.eval()
+
+# ---- Load observational data from validation file ----
+val_data = np.load(paths.data / "chempy_TNG_val_data.npz", mmap_mode="r")
+val_theta = val_data["params"]
+val_x = val_data["abundances"]
+
+
+def clean_data(x, y):
+    index = np.where((y == 0).all(axis=1))[0]
+    x = np.delete(x, index, axis=0)
+    y = np.delete(y, index, axis=0)
+    index = np.where(np.isfinite(y).all(axis=1))[0]
+    return x[index], y[index]
+
+
+val_theta, val_x = clean_data(val_theta, val_x)
+obs_abundances = val_x[0]
+obs_errors = np.ones_like(obs_abundances) * 0.05  # fixed Gaussian noise
+
+
+# ---- NumPyro model using PyTorch emulator ----
+def numpyro_model(obs_abundances, obs_errors):
+    alpha_imf = numpyro.sample("alpha_imf", dist.Normal(-2.3, 0.3))
+    log10_n_ia = numpyro.sample("log10_n_ia", dist.Normal(-2.89, 0.3))
+    log10_sfe = numpyro.sample("log10_sfe", dist.Normal(-0.3, 0.3))
+    log10_sfr_peak = numpyro.sample("log10_sfr_peak", dist.Normal(0.55, 0.1))
+    xout = numpyro.sample("xout", dist.Normal(0.5, 0.1))
+    birth_time = numpyro.sample("birth_time", dist.Uniform(1.0, 13.8))
+
+    input_tensor = torch.tensor(
+        [alpha_imf, log10_n_ia, log10_sfe, log10_sfr_peak, xout, birth_time],
+        dtype=torch.float32,
+    )
+    predicted_abundances = model(input_tensor).detach().numpy()
+
+    numpyro.sample(
+        "obs", dist.Normal(predicted_abundances, obs_errors), obs=obs_abundances
+    )
+
+
[…]

### 2025-06-02 src/scripts/emcee_sampling.py
[showyourwork] overleaf sync

@@ -1,0 +1,284 @@
+import pickle
+import time as t
+
+import corner
+import emcee
+import matplotlib.pyplot as plt
+import numpy as np
+import sbi.utils as utils
+import torch
+from Chempy.parameter import ModelParameters
+from scipy.stats import norm
+from torch.distributions.normal import Normal
+from torch.distributions.uniform import Uniform
+from tqdm import tqdm
+
+import paths
+from chempy_torch_model import Model_Torch
+
+# ----- Config -------------------------------------------------------------------------------------------------------------------------------------------
+name = "NPE_C"
+
+# ----- Load the data -----
+a = ModelParameters()
+labels_out = a.elements_to_trace
+labels = [a.to_optimize[i] for i in range(len(a.to_optimize))] + ["time"]
+priors = torch.tensor([[a.priors[opt][0], a.priors[opt][1]] for opt in a.to_optimize])
+
+# --- Define the priors ---
+local_GP = utils.MultipleIndependent(
+    [Normal(p[0] * torch.ones(1), p[1] * torch.ones(1)) for p in priors[2:]]
+    + [Uniform(torch.tensor([2.0]), torch.tensor([12.8]))],
+    validate_args=False,
+)
+
+global_GP = utils.MultipleIndependent(
+    [Normal(p[0] * torch.ones(1), p[1] * torch.ones(1)) for p in priors[:2]],
+    validate_args=False,
+)
+
+# ----- load the posterior -------------------------------------------------------------------------------------------------------------------------------------------
+with open(paths.data / f"posterior_{name}.pickle", "rb") as f:
+    posterior = pickle.load(f)
+
+# ---- Load your emulator weights ----
+model = Model_Torch(len(labels), len(labels_out))
+model.load_state_dict(torch.load(paths.data / "pytorch_state_dict.pt"))
+model.eval()
+
+# --- Define the simulator ---
+N_stars = 200
+N_samples = 1000
+
+stars = local_GP.sample((N_stars,))
+global_params = torch.tensor([[-2.3, -2.89]])
+
+stars = torch.cat((global_params.repeat(N_stars, 1), stars), dim=1)
+
+# ----- Simulate abundances -----
+start = t.time()
+abundances = model(stars)
[…]

### 2025-06-03 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-This factorization of the joint posterior into single star posteriors $\mathbb{P}(\vec\Lambda|D_j)$ has the advantage that we do not need to specify the exact number of observational data points beforehand.
+This factorization of the \resub{marginal} joint posterior \resub{of the global parameters} into single star posteriors $\mathbb{P}(\vec\Lambda|D_j)$ has the advantage that we do not need to specify the exact number of observational data points beforehand.
 We can simply train our neural density estimator on single star observations and then at evaluation time, we simply combine as many posterior estimates as we have observational data.
@@
 With this it is straight forward to evaluate eq.~\ref{eq:posterior} analytically.
-In fact, under our assumption the combined posterior is a product between the prior and the product of multiple Gaussians for the individual star posteriors.
-The latter product is also a Gaussian with mean $\mathbf{\mu'}$ and variance $\mathbf{\sigma'^2}$:
+In fact, under our assumption the combined \resub{factorized posterior from eq.~\ref{eq:posterior} is again a Gaussian whose mean and covariance can be calculated from the means and covariances of the individual prior/posterior Gaussians as follows:}
+\resub{
 \begin{align}
-    \mathbf{\mu'} &= \frac{\sum_{i=1}^{N_{stars}} \frac{\mu_i}{\sigma_i^2}}{\sum_{i=1}^{N_{stars}} \frac1{\sigma_i^2}} \\
-\mathbf{\sigma'}^2 &= \frac1 {\sum_{i=1}^{N_{stars}} \frac1{\sigma_i^2}}
+    \mathbf{\Sigma}_{\rm Posterior}^{-1} &= (N-1) \mathbf{\Sigma}_{\rm Prior}^{-1} + \sum_i^N \mathbf{\Sigma}_i^{-1} \\
+\mathbf{\mu}_{\rm Posterior} &= \mathbf{\Sigma}_{\rm Posterior} \Big( (N-1) \mathbf{\Sigma}_{\rm Prior}^{-1} \mathbf{\mu}_{\rm Prior} + \sum_i^N \mathbf{\Sigma}_i^{-1} \mathbf{\mu}_i \Big)
 \end{align}
+}
 
-Further, in our case the prior for the galactic parameters $\Lambda$ is Gaussian as well.
-Therefore the resulting factorized posterior from eq.~\ref{eq:posterior} is again a Gaussian and can be expressed with mean $\mathbf{\mu}$ and variance $\mathbf{\sigma}$ as:
-\begin{align}
-\mathbf{\mu} &= \frac{\frac{\mu'}{\sigma'^2}-\frac{(1-N)\mu_ {prior}}{\sigma_ {prior}^2}}{\frac1{\sigma'^2}-\frac{(1-N)}{\sigma_ {prior}^2}} \\
-\mathbf{\sigma}^2 &= \frac1 {\frac1{\sigma'^2}-\frac{(1-N)}{\sigma_ {prior}^2}}
-\end{align}
+\resub{where we have assumed that the prior for the global parameters has no  covariance such that $\Sigma_\rm{prior}$ is diagonal.}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/corner_plot_singlestar.pdf}
+    \includegraphics[width=\textwidth]{figures/corner_plot_comparison_singlestar_17.pdf}
     \vspace{-.5cm}
     \caption{Corner plot of the posteriors for all six parameters for a single star from the validation set.
-The gray contours show a kde-estimate of the posterior from our SBI inference and the red dot and lines show the ground truth parameter values.
-Gray histograms on the diagonal show a kde estimate of the marginals.}
+The blue contours show a kde-estimate of the posterior from our SBI inference and the red dot and lines show the ground truth parameter values.
+The orange contours show results from a MCMC run on the same data.
+Blue/orange histograms on the diagonal show a kde estimate of the marginals.}
     \label{fig:corner_plot}
-    \script{plot_sbc.py}
+    \script{emcee_sampling.py}
 \end{figure*}
@@
     \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $100$ stars.
-The ground truth value is shown by the red star, the posterior mean of the SBI inference is shown with the black dot and white ellipses show the $1-3\sigma$ contours of our inference.
-The gray ellipses show the results for HMC inference performed by \citet{Philcox_2019}.
+The ground truth value is shown by the red star, the posterior mean of the SBI inference is shown with the blue dot and blue ellipses show the $1-3\sigma$ contours of our inference.
 See Sec.~\ref{subsec:chempy_tng} for a full description.}
@@
 
-\section{\resub{Comparison between HMC and SBI}}
+\section{\resub{Comparison between MCMC and SBI}}
 \label{sec:hmc-sbi}
-\resub{put a new figure and some text...}
+\resub{Figure~\ref{fig:hmc} shows a comparison of the marginal joint posterior of the global parameters for 200 stars between MCMC (orange) and our SBI inference (blue).
+As already shown for a single star in Fig.~\ref{fig:corner_plot} our SBI results converge nicely to the MCMC results also when combining individual inferences over 200 stars.}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/hmc_corner_plot.pdf}
+    \includegraphics[width=\columnwidth]{figures/mcmc_results.pdf}
     \vspace{-.5cm}
-    \caption{Caption}
+    \caption{Marginal joint posterior of the global parameters for both SBI (blue) and MCMC (orange) for 200 stars. Our SBI inference converges nicely to the result of the MCMC inference.}
     \label{fig:hmc}
-    \script{hmc.py}
+    \script{emcee_sampling.py}
 \end{figure}

### 2025-06-03 src/scripts/emcee_sampling.py

@@ -234,20 +234,20 @@ plotter = PlotSinglePosterior(
 
 sbi_samples = []
 
-for k in range(len(abundances)):
-    fig = plotter(
-        posterior=posterior,
-        x=abundances[k].detach().numpy(),
-        theta=stars[k].detach().numpy(),
-        plot_kws=dict(fill=True),
-        mh_samples=mh_samples[k]["samples"],
-        plot_kws_per_model={
-            "SBI": dict(levels=[0.05, 0.32, 1], color=color_sbi, fill=True, alpha=0.6),
-            "MCMC": dict(levels=[0.05, 0.32, 1], color=color_mh, fill=True, alpha=0.4),
-        },
-        signature=f"MCMC_{k}_",
-    )
-    fig.savefig(paths.figures / f"corner_plot_comparison_singlestar_{k}.pdf")
+k = 17  # Choose a specific star index for the plot
+fig = plotter(
+    posterior=posterior,
+    x=abundances[k].detach().numpy(),
+    theta=stars[k].detach().numpy(),
+    plot_kws=dict(fill=True),
+    mh_samples=mh_samples[k]["samples"],
+    plot_kws_per_model={
+        "SBI": dict(levels=[0.05, 0.32, 1], color=color_sbi, fill=True, alpha=0.6),
+        "MCMC": dict(levels=[0.05, 0.32, 1], color=color_mh, fill=True, alpha=0.4),
+    },
+    signature=f"MCMC_{k}_",
+)
+fig.savefig(paths.figures / f"corner_plot_comparison_singlestar_{k}.pdf")
 
 for k in range(len(abundances)):
     tmp_samples = np.load(paths.data / f"MCMC_{k}_single_samples.npy")

### 2025-06-03 src/scripts/plot_functions.py

@@ -543,10 +543,10 @@ def gaussian_posterior_plot_n_stars(
     legend_true = plt.scatter(
         global_params[0, 0],
         global_params[0, 1],
-        marker="o",
+        marker="*",
         color="red",
         label=label_gt,
-        s=150,
+        s=200,
     )
     plt.axhline(global_params[0, 1], color="r", linestyle="dashed", linewidth=2)
     plt.axvline(global_params[0, 0], color="r", linestyle="dashed", linewidth=2)
@@ -582,22 +582,13 @@ def gaussian_posterior_plot_n_stars(
         s = int(sigma[i])
         t.set(text=f"{s} $\\sigma$", color="gray")
 
-    legend_true = plt.scatter(
-        global_params[0, 0],
-        global_params[0, 1],
-        color="k",
-        label=label_gt,
-        s=400,
-        marker="x",
-        linewidths=5,
-    )
     legend_fit = plt.scatter(
         mu_alpha, mu_log10N_Ia, color="blue", label=label_fit, s=100
     )
 
     legend_fit = plt.legend(
         handles=[legend_fit],
-        fontsize=15,
+        fontsize=35,
         shadow=True,
         fancybox=True,
         loc=2,
@@ -605,7 +596,7 @@ def gaussian_posterior_plot_n_stars(
     )
     legend_true = plt.legend(
         handles=[legend_true],
-        fontsize=15,
+        fontsize=35,
         shadow=True,
         fancybox=True,
         loc=2,
@@ -619,7 +610,7 @@ def gaussian_posterior_plot_n_stars(
     plt.gca().add_artist(legend_fit)
     plt.gca().add_artist(legend_true)
 
-    # plt.title(title, fontsize=60)
+    plt.title(title, fontsize=40)
 
     plt.tight_layout()
     plt.savefig(paths.figures / f"{title}.png")

### 2025-06-03 src/scripts/emcee_sampling.py
[showyourwork] overleaf sync

@@ -283,7 +283,7 @@ weights = np.ones(len(all_samples)) / len(all_samples)
 corner.corner(
     all_samples,  # shape: [N_samples, D]
     weights=weights,
-    labels=param_names,
+    labels=labels_in[:2],  # only plot the first two parameters
     fig=fig,  # reuse the same figure
     color="C1",
     bins=30,

### 2025-06-03 src/scripts/plot_functions.py

@@ -284,7 +284,7 @@ def gaussian_posterior_plot(
     grid_x = [global_params[0, 0] + xlim[0], global_params[0, 0] + xlim[1]]
     grid_y = [global_params[0, 1] + ylim[0], global_params[0, 1] + ylim[1]]
 
-    x, y = np.mgrid[grid_x[0] : grid_x[1] : 0.001, grid_y[0] : grid_y[1] : 0.001]
+    x, y = np.mgrid[grid_x[0] : grid_x[1] : 0.0001, grid_y[0] : grid_y[1] : 0.0001]
     pos = np.dstack((x, y))
 
     # create a multivariate normal
@@ -449,7 +449,7 @@ def gaussian_posterior_plot_n_stars(
     grid_x = [global_params[0, 0] + xlim[0], global_params[0, 0] + xlim[1]]
     grid_y = [global_params[0, 1] + ylim[0], global_params[0, 1] + ylim[1]]
 
-    x, y = np.mgrid[grid_x[0] : grid_x[1] : 0.001, grid_y[0] : grid_y[1] : 0.001]
+    x, y = np.mgrid[grid_x[0] : grid_x[1] : 0.0001, grid_y[0] : grid_y[1] : 0.0001]
     pos = np.dstack((x, y))
 
     # create a multivariate normal
@@ -547,9 +547,14 @@ def gaussian_posterior_plot_n_stars(
         color="red",
         label=label_gt,
         s=200,
+        zorder=5,
     )
-    plt.axhline(global_params[0, 1], color="r", linestyle="dashed", linewidth=2)
-    plt.axvline(global_params[0, 0], color="r", linestyle="dashed", linewidth=2)
+    plt.axhline(
+        global_params[0, 1], color="r", linestyle="dashed", linewidth=2, zorder=5
+    )
+    plt.axvline(
+        global_params[0, 0], color="r", linestyle="dashed", linewidth=2, zorder=5
+    )
 
     label_fit = (
         r"SBI"
@@ -588,15 +593,15 @@ def gaussian_posterior_plot_n_stars(
 
     legend_fit = plt.legend(
         handles=[legend_fit],
-        fontsize=35,
+        fontsize=30,
         shadow=True,
         fancybox=True,
         loc=2,
-        bbox_to_anchor=(0, 0.9),
+        bbox_to_anchor=(0, 0.85),
     )
     legend_true = plt.legend(
         handles=[legend_true],
-        fontsize=35,
[…]
